\documentclass[11pt, a4paper, reqno, captions=tableheading,bibliography=totoc]{scrartcl}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{lmodern}
\usepackage[english]{babel}
\usepackage{booktabs}
\usepackage{float}
\usepackage{enumitem}
\usepackage{tikz,pgf}
%\usepackage[utopia]{mathdesign}
\usepackage{palatino}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}
\usepackage[nameinlink]{cleveref}
\usepackage{xcolor}

\theoremstyle{plain}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{prop}[lemma]{Proposition}
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{conjecture}[lemma]{Conjecture}
\newtheorem{fact}[lemma]{Fact}
\newtheorem{assumption}[lemma]{Assumption}
\newtheorem*{reduction}{Reduction}
\theoremstyle{definition}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{es}[lemma]{Example}
\newtheorem*{notation}{Notation}
\newtheorem{rmk}[lemma]{Remark}


\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\p}{\mathbb{P}}
\newcommand{\sP}{\mathcal{P}}
\newcommand{\sL}{\mathcal{L}}
\newcommand{\de}{\partial}
\newcommand{\codim}{\mathrm{codim}}

\newcommand{\oo}{\mathcal{O}}
\newcommand{\Bl}{\mathrm{Bl}}

\newcommand{\iso}{\mathcal{C}_{\mathrm{iso}}}

\newcommand{\imunit}{i}

\newcommand{\SO}{\operatorname{SO}}
\newcommand{\Eig}[1]{\operatorname{Eig}\left( {#1} \right)}
\newcommand{\polq}{{\rm Pol}_Q}
\newcommand{\comment}[1]{}

\newcommand{\scl}[2]{\left\langle {#1}, {#2} \right\rangle}

\newcommand\scalemath[2]{\scalebox{#1}{\mbox{\ensuremath{\displaystyle #2}}}}

\newcommand{\iii}{\textbf{i}}
\definecolor{MyDarkGreen}{cmyk}{0.7,0,1,0}
\newcommand{\blue}[1]{{\color{blue}  [#1]}}
\newcommand{\cbc}{\ensuremath{Cbc}}

\title{Eigenpoint collinearities of plane cubics}
\author{}
\date{}

\linespread{1.15}
\setlength{\parindent}{0pt}
\setlength{\parskip}{.25em}

\begin{document}

\maketitle

\section{Introduction}
\label{introduction}

However, there are several examples of cubic polynomials, whose eigenscheme contains one or more triples of aligned points, as for instance the Fermat cubic polynomial.

The goal of this paper is to classify all the situations when we have one or more triples of aligned points inside a zero-dimensional reduced eigenscheme of a cubic plane curve.

\Cref{aligned} clarifies the reasons of interest in the situation of triples of aligned eigenpoints.


\section{Aligned eigenpoints of ternary cubic form}
\label{aligned}

We recall that, given a homogeneous form $f \in \C[x,y,z]_d$ of degree~$d$, the eigenscheme~$\Eig{f}$ of~$f$ is the determinantal scheme defined by the $2 \times 2$ minors of the matrix
%
\begin{equation}
\label{eq:def_matrix}
    \begin{pmatrix}
    x & y & z \\
    \de_x f & \de_y f & \de_z f
    \end{pmatrix}.
\end{equation}
%
Since the eigenschemes of two proportional homogenous forms are the same,
if $C$ is the curve defined by these forms,
we can write $\Eig{C}$ for such eigenscheme and hence talk about the eigenscheme of a plane curve.

The eigenscheme of a general ternary cubic form has no aligned triples of points. This is a consequence of the geometric properties of the classical Geiser map associated with seven points in the plane, and has been proved in \cite[Proposition~4.5]{BGV}.

Moreover, whenever the eigenscheme of a ternary form is zero-dimensional, it never cointains $4$ or more aligned points.
To see this, we recall a property of eigenschemes of ternary forms, namely,
that when zero-dimensional they are somehow ``general'' with respect to conics.

\begin{lemma}
\label{lemma:no_six_conic}
Let $f \in \C[x,y,z]_d$ be a homogeneous form of degree~$d$.
If $E(f)$ is zero-dimensional,
then no degree six subscheme of~$E(f)$ lies on a conic.
\end{lemma}
\begin{proof}
See \cite[Lemma~9.1]{OS1} for the reduced case.
The proof works also in the non-reduced case.
\end{proof}

\begin{corollary}
\label{corollary:general_no_triple}
As a consequence of \Cref{lemma:no_six_conic}, a zero-dimensional eigenscheme~$\Eig{f}$ for $f \in \C[x,y,z]_d$ never contains $4$ or more aligned points.
Moreover, if $\Eig{f}$ contains two triples of aligned points, those must share a point.
\end{corollary}

Therefore, we investigate the situations when we have one or more triples of aligned eigenpoints in a zero-dimensional, reduced eigenscheme of a ternary cubic form.

\section{Invariance under the action of orthogonal matrices}
\label{invariance}

\begin{definition}
 We define $\mathrm{SO}_3(\mathbb{C})$ to be the complexification of the group of special orthogonal real matrices, namely
 %
 \[
  \mathrm{SO}_3(\mathbb{C}) :=
  \bigl\{
   M \in \mathrm{GL}_3(\C) \, \mid \,
   M M^t = I_3 \  \text{and} \  \det(M) = 1
  \bigr\} \,.
 \]
 %
 The group $\mathrm{SO}_3(\mathbb{C})$ acts on $\C^3$ by matrix multiplication:
 %
 \[
  \begin{array}{ccc}
   \mathrm{SO}_3(\mathbb{C}) \times \C^3 & \rightarrow & \C^3 \\
   (M, v) & \mapsto & Mv
  \end{array}
 \]
 %
 Since all the elements of $\mathrm{SO}_3(\mathbb{C})$ are invertible, the latter action descends to an action on $\p^2(\C)$.

 Moreover, the group~$\mathrm{SO}_3(\mathbb{C})$ acts also on ternary forms via
 \[
  M \cdot f (x,y,z) = f(M^{-1} \cdot \prescript{t} {}( x \ y \ z )  ).
 \]
\end{definition}

\begin{prop}
\label{two_orbits}
 The action of $\mathrm{SO}_3(\mathbb{C})$ on $\p^2(\C)$ has two orbits:
 %
 \begin{align*}
  \mathcal{O}_1 &:=
  \bigl\{
   P \in \p^2(\C) \, | \,
   P = (a:b:c) \  \text{with} \  a^2 + b^2 + c^2 = 0
  \bigr\} \\
  \mathcal{O}_2 &:= \p^2(\C) \setminus \mathcal{O}_1
 \end{align*}
 %
 A representative for $\mathcal{O}_1$ is $(1:\iii:0)$ and a
representative for $\mathcal{O}_2$ is $(1:0:0)$.
\end{prop}
\begin{proof}
 Suppose that $P \in \p^2(\C)$ and $P = (a:b:c)$ with $a^2 + b^2 + c^2 = 0$.
 We produce a matrix $M \in \mathrm{SO}_3(\C)$ such that $M \left(\begin{smallmatrix} 1 \\ \iii \\ 0 \end{smallmatrix}\right)$ and $\left(\begin{smallmatrix} a \\ b \\ c \end{smallmatrix}\right)$ are proportional.
 Up to relabeling the coordinates, we can suppose that $a \neq 0$.
 Hence, by rescaling the coordinates of $P$, we have $P = (1: b: c)$ with $b^2 + c^2 = -1$.
 One can check that the matrix
 %
 \[
  M :=
  \begin{pmatrix}
   -1 & 0 & 0 \\
   0 & \iii b & -\iii c \\
   0 & \iii c & \iii b
  \end{pmatrix}
 \]
 %
 satisfies the requirements.

 Now suppose that $P \in \p^2(\C)$ and $P = (a:b:c)$ with $a^2 + b^2 + c^2 \neq 0$.
 Up to rescaling, we can suppose that $a^2 + b^2 + c^2 = 1$.
 Again, we produce a matrix $M \in \mathrm{SO}_3(\C)$ such that $M \left(\begin{smallmatrix} 1 \\ 0 \\ 0 \end{smallmatrix}\right)$ and $\left(\begin{smallmatrix} a \\ b \\ c \end{smallmatrix}\right)$ are proportional.
 First of all, suppose that $b^2 + c^2 \neq 0$ and let $\omega$ be a root of the polynomial $t^2 - (b^2 + c^2)$ in $\C[t]$.
 Then, the matrix
 %
 \[
   M :=
   \begin{pmatrix}
     a & \omega & 0 \\
     b & -\frac{ab}{\omega} & \frac{c}{\omega} \\
     c & -\frac{ac}{\omega} & -\frac{b}{\omega}
   \end{pmatrix}
 \]
 %
 satisfies the requirements.
 With the same technique, if $a^2 + c^2 \neq 0$, we can produce a matrix $M \in \mathrm{SO}_3(\C)$ that maps $\left(\begin{smallmatrix} 0 \\ 1 \\ 0 \end{smallmatrix}\right)$ to $\left(\begin{smallmatrix} a \\ b \\ c \end{smallmatrix}\right)$; similarly, when $a^2 + b^2 \neq 0$, we can map $\left(\begin{smallmatrix} 0 \\ 0 \\ 1 \end{smallmatrix}\right)$ to $\left(\begin{smallmatrix} a \\ b \\ c \end{smallmatrix}\right)$.
 Since $\left(\begin{smallmatrix} 1 \\ 0 \\ 0 \end{smallmatrix}\right)$, $\left(\begin{smallmatrix} 0 \\ 1 \\ 0 \end{smallmatrix}\right)$, and $\left(\begin{smallmatrix} 0 \\ 0 \\ 1 \end{smallmatrix}\right)$ are all $\mathrm{SO}_3(\C)$-equivalent, the only case to consider is when
 %
 \[
  b^2 + c^2 = a^2 + c^2 = a^2 + b^2 = 0 \,,
 \]
 %
 which, however, can never occur.
\end{proof}

The following result is well known; we recall it for the sake of completeness.

\begin{prop}
 Let $M \in \mathrm{GL}_3(\C)$ and let $f$ be a ternary cubic.
 Let $P = (A: B: C)$ be a point in~$\p^2$.
 Then we have
 \[
  P \in \Eig{f} \iff M \cdot \prescript{t} {}(A \ B \ C) \in \Eig{M \cdot f}.
 \]
\end{prop}

\begin{proof}
In the present proof,
for convenience, we shall consider the transpose of the defining matrix of an eigenscheme.

A point $P = (A: B: C)$ is an eigenpoint for~$f$ if and only if
\begin{equation*}
  \text  {rk}  \begin{pmatrix}
    A & \de_x f(P) \\
    B & \de_y f(P)  \\
    C & \de_z f(P)
    \end{pmatrix}=1,
\end{equation*}
which is equivalent to
\begin{equation}
\label{eq:def_matrix_M}
    \text  {rk} \quad  M  \cdot \begin{pmatrix}
    A & \de_x f(P) \\
    B & \de_y f(P)  \\
    C & \de_z f(P)
    \end{pmatrix}
    =1.
\end{equation}
By setting $\prescript{t} {}(A' \ B' \ C' )= M \cdot \prescript{t} {}(A \ B \ C) $ and $Q=(A':B':C')$, we have that \Cref{eq:def_matrix_M}
is equivalent to
%
\begin{equation}
\label{eq:transformed}
  \text{rk}
  \begin{pmatrix}
    A' &  \\
    B' & M \cdot \nabla f (P) \\
    C' & \\
  \end{pmatrix}=1.
\end{equation}
%
Now we consider the polynomial $M \cdot f$ and we observe that the chain rule gives
%
\begin{gather*}
\partial_x (M\cdot f) = \partial_x \bigl( f(M^{-1}  \ \prescript{t} {} (x \ y \ z)) \bigr) = \prescript{t} {}(M^{-1})^{(1)}(\nabla f) \bigl( M^{-1}\   \prescript{t} {} (x \ y \ z) \bigr) \,, \\
\partial_y (M\cdot f) = \partial_y \bigl( f(M^{-1}  \ \prescript{t} {} (x \ y \ z)) \bigr) = \prescript{t} {}(M^{-1})^{(2)}(\nabla f) \bigl( M^{-1}\   \prescript{t} {} (x \ y \ z) \bigr) \,, \\
\partial_z (M\cdot f) = \partial_z  \bigl( f(M^{-1}  \ \prescript{t} {} (x \ y \ z)) \bigr) = \prescript{t} {}(M^{-1})^{(3)}(\nabla f) \bigl( M^{-1}\   \prescript{t} {} (x \ y \ z) \bigr) \,,
\end{gather*}
%
where $(M^{-1})^{(j)}$ denotes the $j$-th column of the matrix $M^{-1}$. Hence
%
\[
\nabla (M \cdot f) = \prescript{t} {} M^{-1} \cdot (\nabla f) (M^{-1}\   \prescript{t} {} (x \ y \ z)),
\]
%
so we have
%
\[
\nabla (M \cdot f)(Q)=\nabla (M \cdot f)(M \cdot P)=
\prescript{t} {} M^{-1} \cdot (\nabla f) (M^{-1}\   M \cdot P)=\prescript{t} {} M^{-1} \cdot (\nabla f)(P).
\]
%
Finally, if we choose $M \in \mathrm{SO}_3(\mathbb{C})$, we have
$\prescript{t} {} M^{-1}=M$. We deduce that
\Cref{eq:transformed} holds if and only if $Q \in E(M\cdot f)$, so the statement is proved.
\end{proof}

\section{Conditions imposed by aligned eigenpoints}
\label{conditions}

Imposing a cubic ternary form to have one or more aligned triples of eigenschemes implies conditions both on the points and on the cubics.
We begin to explore these conditions by introducing, for each point
$P \in \p^2$,
a $3 \times 10$ matrix encoding the condition that $P$ is an eigenpoint of a ternary cubic.

\begin{definition}
\label{definition:matrix_conditions}
 Consider $\p^9 = \p(\C[x,y,z]_3)$, the space of all ternary cubics.
 Throught this paper, we fix the following vector basis for $\C[x,y,z]_3$:
 \begin{eqnarray}
  \mathcal{B} = (x^3, x^2 y, x y^2, y^3, x^2 z, x y z, y^2 z, x z^2, y z^2, z^3)
  \label{vector_basis}
 \end{eqnarray}
 For $f \in \C[x,y,z]_3$, denote by $[f]$ the corresponding point in~$\p^9$; we denote by $w_f$ the (column) vector of coordinates of~$f$ with respect to the basis above; then $w_f$ is also a vector of projective coordinates of~$[f]$.
 For a point $P \in \p^2$ with coordinates $(A: B: C)$, the condition on elements~$[f]$ of~$\p^9$ that $P$ is an eigenpoint of the ternary cubic form~$f$ can be expressed in the form
 %
 \[
  \Phi(P) \cdot w_f
  = 0 \,,
 \]
 %
 where $\Phi(P)$ is a $3 \times 10$ matrix with entries depending on $A, B, C$.
 The matrix $\Phi(P)$ is called the \emph{matrix of conditions} imposed by~$P$.
We denote by $\phi_1(P)$, $\phi_2(P)$, and~$\phi_3(P)$ the rows of~$\Phi(P)$.
Written as vectors, they are
%
\begin{equation}
\label{equation:matrix_conditions_rows}
\begin{gathered}
\scalemath{0.9}{(-3A^2B, A(A^2 - 2B^2), B(2A^2 - B^2), 3AB^2,
 -2ABC, C(A^2 - B^2), 2 ABC,
 -B  C^2, A  C^2, 0)} \,, \\
\scalemath{0.9}{(-3A^2 C,
-2ABC,
-CB^2,
0,
A(A^2-2C^2),
B(A^2 - C^2),
AB^2,
C(2A^2-C^2),
2ABC,
3AC^2)} \,,\\
\scalemath{0.9}{(0,
-A^2C,
-2ABC,
-3CB^2,
A^2 B,
A(B^2 - C^2),
B(B^2-2C^2),
2ABC,
C(2B^2-C^2),
3BC^2)} \,.
\end{gathered}
\end{equation}
%
Moreover, if $P_1, \dotsc, P_n$ are points in the plane, we denote by $\Phi(P_1, \dotsc, P_n)$ the matrix
%
\[
 \left(
 \begin{array}{c}
  \Phi(P_1) \\
  \vdots \\
  \Phi(P_n)
 \end{array}
 \right)
\]
%
namely, the $3n \times 10$ matrix obtained by vertically stacking the matrices of conditions of~$P_1, \dotsc, P_n$.
\end{definition}

\begin{rmk} @@si puo' dire meglio, usando le derivate parziali@@)
By analyzing the entries of \Cref{equation:matrix_conditions_rows}, it is not difficult to check that the rank of
$\Phi(P)$ is never $\leq 1$.
Moreover, it holds:
\begin{equation}
  C \, \phi_1(P) - B \, \phi_2(P) + A \, \phi_3(P) = 0,
  \label{eq:base}
\end{equation}
therefore,
among the three vectors, at most two of them are linearly independent.
Hence, the matrix~$\Phi(P)$ has rank~$2$.
\end{rmk}

As a consequence, if $P_1, \dots, P_n$ are $n$ points of the plane, we have:
\begin{equation}
\label{bound_rank}
\text{rk} \  \Phi(P_1, \dots, P_n) \leq \min \left\{2n, 10 \right\}
\end{equation}

In what follows we want to study the possible values of the rank of the matrix
$\Phi(P_1, \dots, P_n)$ for several configurations of points $P_1, \dots, P_n$
(and several values of $n$).
In particular, we will study the ideal~$J_k$ of order $k$ minors of the
involved matrix and we will deduce some bounds about the rank from the possible
decompositions of the ideal~$J_k$. Most of these computations will be done
with the aid of a computer algebra system. Nevertheless, in many cases,
the result cannot be reached just by brute force, but it is necessary to
make some preprocessing on the ideal~$J_k$. In particular, it turns out that
it is often convenient to first saturate the ideal~$J_k$ with respect to
the conditions that the
points are distinct or that three of them are not aligned (when this is the
case). Another important simplification that we adopt sometimes, makes use
of the action of $\SO_3(\C)$: thanks to it we can assume that one of
the point is either $(1: 0: 0)$ or $(1: \iii: 0)$; see \Cref{two_orbits}.

Finally, several times the following lemma will be extremely useful
to speed up the computations.

\begin{lemma}
\label{lemma:minors}
Let $l_1 < \cdots <l_n$ be $n$ indices (where $3 \leq n \leq 10$) and let $P = (A: B: C)$ be a point of the plane.
Construct three $1 \times n$ matrices $w_1$, $w_2$, $w_3$ by extracting the entries of position $l_1, \dotsc, l_n$ from $\phi_1(P)$, $\phi_2(P)$, and~$\phi_3(P)$, respectively. If $L$ is a $(n-2) \times n$ matrix, set:
  \[
  L_1 := \left(\begin{array}{c}w_1 \\ w_2 \\ L\end{array}  \right), \quad
  L_2 := \left(\begin{array}{c}w_1 \\ w_3 \\ L\end{array}  \right), \quad
  L_3 := \left(\begin{array}{c}w_2 \\ w_3 \\ L\end{array}  \right)
  \]
  Then
  \[
  B \det(L_1) = A \det(L_2), \quad
  C \det(L_1) = A \det(L_3), \quad
  C \det(L_2) = B \det(L_3)
  \]
  hence $(A: B: C) = \bigl( \det(L_1): \det(L_2): \det(L_3) \bigr)$.
\end{lemma}
\begin{proof}
  The thesis easily follows from the equality $C w_1 - B w_2 + A w_3 = 0$, which is a direct consequence of \Cref{eq:base}.
\end{proof}

\begin{definition}
 \label{definition:reduced_matrix_conditions}
 For $n$ points $P_1, \dotsc, P_n$ in the plane, the \emph{reduced matrix of conditions} of $P_1, \dotsc, P_n$ is the submatrix of~$\Phi(P_1, \dotsc, P_n)$ whose rows are $\phi_1(P_1), \phi_2(P_1), \dotsc, \phi_1(P_n), \phi_2(P_n)$.
\end{definition}


% \begin{lemma}
%   \label{lemma_for_minors2}
%   Let $P(A: B: C)$, $Q(A': B': C')$ be two points of $\mathbb{P}^2_K$
%   and set $R = \alpha P+\beta Q$ (where $\alpha, \beta \in K$).
%   Suppose $l_1 < \cdots < l_n$ are as above, fix $j \in \{1, 2, 3\}$
%   and   let $w_1$ be the vector obtained
%   from the entries of $\phi_j(P)$ of position $l_1, \dots, l_n$
%   and $w_2$ be the vector analogously obtained from $R$. If
%   $L$ is a square matrix which contains the
%   rows $w_1$ and $w_2$, then $\alpha\beta$ divides $\det(L)$.
% \end{lemma}
% \begin{proof}
%   We consider $\det(L)$ as a polynomial in $\alpha$.
%   If we set $\alpha = 0$,
%   we have that $w_1$ and $w_2$ are proportional, so, if $\alpha = 0$
%   then $\det(L)$ is $0$. Hence $\alpha$ divides $\det(L)$. Same
%   argument for $\beta$.
% \end{proof}


First of all, we point out a property of the lines, which are tangent to~$\iso$.
Let $P_1, P_2$ be two distinct points in the plane and define
%
\[
  \sigma(P_1, P_2) := \scl{P_1}{P_1} \scl{P_2}{P_2} - \scl{P_1}{P_2}^2 \,.
\]
%
The quantity~$\sigma$ is the discriminant of the intersection between the line~$r$ through~$P_1$ and~$P_2$ and~$\iso$.

\begin{prop}
\label{proposition:sigma_tangency}
  Let $P_1$, $P_2$ be two distinct points in the plane and let $r$ be the line joining them.
  Then the following are equivalent:
  \begin{enumerate}
  \item $\sigma(P_1, P_2) = 0$;
  \item $\sigma(Q_1, Q_2) = 0$ for all pairs of distinct points $Q_1, Q_2 \in r$;
  \item the line $r$ is tangent to~$\iso$ at some point;
  \item there exists a point $T \in r$ such that  $T \in \iso$ and $\scl{T}{Q} = 0$ for all $Q \in r$, $Q \neq T$; in this case we say that $T$ is \emph{orthogonal} to $r$.
  \end{enumerate}
\end{prop}
\begin{proof}
  By recalling the characterization of~$\sigma$ as a discriminant, we have that $\sigma(P_1, P_2) = 0$ if and only if $r$ is tangent to~$\iso$ in a point~$T$; this shows that the first three items are equivalent.

  Now, if $\sigma(P_1, P_2) = 0$, then $r$ is tangent to~$\iso$ at a point~$T$, hence $\scl{T}{T} = 0$. Moreover, we also have that $\sigma(T, Q) = 0$ for all $Q \in r$ with $Q \neq T$. Hence, $T$ is orthogonal to~$r$.
  The converse is immediate.
\end{proof}

\begin{prop}
\label{proposition:three_aligned_ranks}
Let $P_1, P_2, P_3$ be three distinct aligned points of the plane and let
$r$ be the line passing through them. Then:
\begin{itemize}
\item $5 \leq \text{rk}\ \Phi(P_1, P_2, P_3) \leq 6$;
\item
$\text{rk}\ \Phi(P_1, P_2, P_3) = 5$ if and only if $r$ is tangent
to~$\iso$ in one of the three points $P_1, P_2$, or $P_3$.
\end{itemize}
\end{prop}
\begin{proof} We denote the coordinates of the points as follows:
\[
P_1, P_2, P_3 = (A_1: B_1: C_1), (A_2: B_2: C_2), u_1P_1+u_2P_2.
\]
We have to consider the ideal generated by the $17640$ order six
minors of the matrix $\Phi(P_1, P_2, P_3)$ and we
claim that we can greately simplify
the computations.

Consider the matrix $M$ constructed with the following six rows:
\[
\phi_{i_1}(P_1), \ \phi_{i_2}(P_1), \ \phi_{j_1}(P_2),\  \phi_{j_2}(P_2),
\phi_{k_1}(P_3), \ \phi_{k_2}(P_3)
\]
where $i_1 \not= i_2 \in \{1, 2, 3\}$; $j_1 \not= j_2 \in \{1, 2, 3\}$;
$k_1 \not= k_2 \in \{1, 2, 3\}$.
First, assume that $i_1=j_1=k_1=1$ and
$i_2=j_2=k_2=2$. Fix six columns
$1\leq l_1 < \cdots l_6 \leq 10$ of $M$ and let $N$ be the order $6$ matrix
obtained from $M$ with these columns. Since two rows of $N$ are obtained from
entries of $\phi_1(P_1)$ and of $\phi_2(P_1)$,
\Cref{lemma:minors} gives that $A_1$ divides $\det(N)$. For a similar
reason, also $A_2$ and $u_1A_1+u_2A_2$ divide $\det(N)$, so
$\det(N) = A_1A_2(u_1A_1+u_2A_2)\cdot D$ for a suitable $D$.
Again by \Cref{lemma:minors}, we also see that if we take different values of
$i_1, i_2, j_1, j_2, k_1, k_2$ in the definition of $M$ above, then
the corresponding $\det(N)$ would be of the form $X_1X_2X_3\cdot D$ where
$X_1$ is a coordinate of $P_1$, $X_2$ is a coordinate of $P_2$ and $X_3$ is
a coordinate of $P_3$ (and $D$ is the same as above). Since each of the three
points have at least one non-zero coordinate, if the order six minor
constructed with the columns $l_1, \dots, l_6$ is zero, then necessarily $D$
must be zero. In this way we see that, in order to have that all the order
six minors of $\Phi(P_1, P_2, P_3)$ are zero, it is enough to compute the
ideal of the order six minors of the matrix $M$ above and divide each
minor by $A_1A_2(u_1A_1+u_2A_2)$. By this procedure, we get an ideal that (after
a saturation w.r.t.\ the condition that $P_1$, $P_2$, and $P_1, P_3$ and
$P_2$, $P_3$ are distinct), has
a very simple primary decomposition, which is given by the following three
ideals:
\[
\left(\scl{P_i}{P_1}, \scl{P_i}{P_2},\scl{P_i}{P_3}\right) \quad
\mbox{for } i = 1, 2, 3 \,.
\]
As a consequence, the line $r$ is tangent to~$\iso$
(in $P_1$ or $P_2$ or $P_3$).
It is not difficult to see via a symbolic computation that it is not possible to have that all the order
$4$-minors of $\Phi(P_1, P_2, P_3)$ are zero.
\end{proof}

\begin{prop}
\label{manca il riferimento su ancillary    non e': condition_rank_aligned}
%%%  conto si trova su file prop47_5ago.sage
Let $P_1, P_2, P_4$ be three distinct points of the plane. Then:
\begin{itemize}
\item $5 \leq \text{rk}\ \Phi(P_1, P_2, P_4) \leq 6$;
\item if
$\text{rk}\ \Phi(P_1, P_2, P_4) = 5$, then $P_1, P_2, P_4$
 are aligned.
\end{itemize}
In the last case, the line $P_1+P_2+P_4$ is tangent to~$\iso$
in one of the three points.
\end{prop}
\begin{proof}
We can split the proof into two parts, considering the case in
which $P_4 = (1: 0: 0)$ and the case in which $P_4 = (1: \iii: 0)$.
In both cases, che computation of the ideal of order five minors of the matrix
$\Phi(P_1, P_2, P_4)$ and the saturation of it w.r.t.\ the condition
that the points $P_1, P_2, P_4$ are distinct gives the whole ring, so
the matrix cannot have rank smaller than~$5$. Meanwhile, the computation
of the ideal of the order six minors of $\Phi(P_1, P_2, P_4)$ and its
saturation w.r.t. the condition that the points are distinct \blue{forse si potrebbe dare un nome a questa condizione}, gives that
$P_1, P_2, P_4$ must be aligned and, according to \Cref{proposition:three_aligned_ranks},
that line is tangent to~$\iso$ in one of the three given points.
\end{proof}


\begin{prop}
\label{prop:condition3+1}
Let $P_1, P_2, P_3, P_4$ be four distinct points of the plane such that
$P_1, P_2, P_3$ are aligned and let $r$
be the line passing joining them. If
$\text{rk}\ \Phi(P_1, P_2, P_3, P_4) \leq 7$ then $r$ is tangent to~$\iso$ in one of the three points $P_1, P_2, P_3$.
\end{prop}
\begin{proof}
Again, we distinguish two cases: $P_1 = (1: 0: 0)$ and
$P_1 = (1: \iii: 0)$. In both cases, the ideal of the order $8$
minors \blue {la comunita' che studia varietà determinantali usa la notazione $I_8(\Phi(P_1, P_2, P_3, P_4))$} of $\Phi(P_1, P_2, P_3, P_4)$ can be computed and saturated
w.r.t.\ the conditions that the points are different and that
$P_1, P_2, P_4$ are not aligned.
The direct inspection of the ideal obtained from these procedures gives the thesis.
\end{proof}

\begin{prop}
\label{proposition:four_aligned}
Let $Q_1, \dotsc, Q_4$ be four distinct aligned points of the plane and
let $r$ be the line through them. Then:
\begin{itemize}
\item $6 \leq \text{rk} \ \Phi(Q_1, \dotsc, Q_4) \leq 7$;
\item $\text{rk} \ \Phi(Q_1, \dotsc, Q_4) = 6$ if and only if $r$ is tangent
to~$\iso$.
\end{itemize}
\end{prop}
\begin{proof}
  A direct symbolic computation shows that all the maximal minors of~$\Phi(Q_1, \dotsc, Q_4)$ are
  zero, so $\mathrm{rk}\,(\Phi(Q_1, \dotsc, Q_4)) \leq 7$.
  Indeed, in view of \Cref{lemma:minors}, we can just check the order~$8$ minors of the reduced condition matrix of $Q_1, \dotsc, Q_4$; it turns out that they are all zero.

  On the other hand, the rank of $\Phi(Q_1, \dotsc, Q_4)$ cannot be $5$:
  if this were the case, for any triple~$T$ extracted from the four points, the matrix of conditions would have rank~$5$, hence by \Cref{proposition:three_aligned_ranks} the line~$r$ would be tangent to~$\iso$ in an element of $T$; this is a contradiction.
  So, the first item of the statement is proven.

  We proceed to prove the second item.
  We take the reduced matrix of conditions of~$Q_1, Q_2, Q_3$, to which we add the row~$\phi_1(Q_4)$.
  We call this matrix~$M_1$.
  Similarly, we construct the matrices $M_2$ and~$M_3$.
  We consider the coordinates of the points as variables and we compute
  the ideal $J_1$, the radical of the ideal of the maximal minors of $M_1$ ; similarly, we obtain $J_2$ and~$J_3$.
  The saturation of the sum $J_1 + J_2 + J_3$ by the conditions that the points are distinct is a principal ideal generated by~$\sigma(r)$.
  Since the four points $Q_1, \dotsc, Q_4$ play symmetric roles, the statement is proven.
\end{proof}

We now define three quantities depending on a triple or on a $5$-tuple of points in the plane.
These quantities are crucial to describe what happens when we have aligned eigenpoints.

\begin{definition}
\label{definition:delta1}
 Let $P$, $Q$ and~$R$ be distinct points in the plane.
 We define the quantity
 %
 \[
  \delta_1(P, Q, R) :=
  \scl{P}{P} \scl{Q}{R} - \scl{P}{Q}\scl{P}{R}
  =
  \scl{P\times Q}{P \times R} \,,
 \]
 %
 where $\times$ denotes the cross product. i.e.,
 %
 \[
  P \times Q = (P_2 Q_3 - P_3 Q_2, \, P_3 Q_1 - P_1 Q_3, \, P_1 Q_2 - P_2 Q_1) \,.
 \]

\end{definition}

\begin{definition}
\label{definition:delta1b}
 Let $P$, $Q$ and~$R$ be distinct aligned points in the plane.
 We define the quantity
 %
 \[
  \overline{\delta}_1(P, Q, R) :=
  \scl{P}{P} \scl{Q}{R} + \scl{P}{Q}\scl{P}{R} \,.
  \]
 %
\end{definition}

\begin{definition}
\label{Vconf}
Let $P_1, P_2, P_3, P_4, P_5$ be five distinct points of the plane
such that $P_1, P_2, P_3$ and $P_1, P_4, P_5$ are aligned.
We call such a configuration a \emph{$V$-configuration}.
\end{definition}


\begin{definition}
 Let $P_1, \dots, P_5$ be a $V$-configuration.
We define the quantity
 %
 \[
  \delta_2(P_1, P_2, P_3, P_4, P_5) :=
  \scl{P_1}{P_2} \scl{P_1}{P_3} \scl{P_4}{P_5} -
  \scl{P_1}{P_4} \scl{P_1}{P_5} \scl{P_2}{P_3} \,.
 \]
 %
\end{definition}

\begin{lemma}
\label{lemma:d1d2}
Let $P_1, \dots, P_5$ be a $V$-configuration. Then
\[
\text{rk}\ \Phi(P_1, \dots, P_5) \leq 9
\quad \mbox{
if and only if} \quad
\delta_1(P_1, P_2, P_4) \cdot \delta_2(P_1, \dots, P_5) = 0.
\]
\end{lemma}
\begin{proof}
We denote the coordinate of the points as follows:
$P_1, P_2, P_4 = (A_1: B_1: C_1), (A_2: B_2: C_2), (A_4: B_4: C_4)$,
then $P_3 = u_1P_1+u_2P_2$ and $P_5 = v_1P_1+v_2P_4$. We proceed as in
the proof of \Cref{proposition:three_aligned_ranks}: we find, via symbolic computation, that the determinant of
the reduced matrix of conditions of $P_1, \dotsc, P_5$ is
%
\begin{gather}
\label{delta1delta2}
A_1A_2A_4(u_1A_1+u_2A_2)(v_1A_1+v_2A_4)u_1^2u_2^2v_1^2v_2^2D^5
\delta_1(P_1,P_2,P_4)\delta_2(P_1,\dots,P_5)
\end{gather}
%
where $D$ is the determinant of the matrix whose rows are $P_1, P_2, P_4$
and is non-zero, as well as are non-zero $u_1, u_2, v_1, v_2$ (since we
assume that $P_1, P_2, P_4$ are not aligned and the points are distinct).
As a consequence of \Cref{lemma:minors}, we have that a different
choice of the rows of $M$ would simply modify in the
determinant from \Cref{delta1delta2} the coordinates
$A_1, A_2, \dots, (v_1A_1+v_2A_4)$ to some other coordinates of the points.
Since every point has at least one non-zero coordinate, we have that
$\text{rk}\ \Phi(P_1, \dots, P_5) \leq 9$ if and only if
\[
\delta_1(P_1, P_2, P_4) \cdot \delta_2(P_1, \dots, P_5) = 0 \,. \qedhere
\]
\end{proof}

\begin{lemma}
\label{lemma:special_case_rank_8}
Let $P_1, \dots, P_5$ be a $V$-configuration of points and assume that
\[
\scl{P_1}{P_2}=0, \quad \scl{P_2}{P_2}=0, \quad \scl{P_1}{P_4}=0,
\quad \scl{P_4}{P_4}=0.
\]
Then the matrix $\Phi(P_1, \dots, P_5)$ has rank $8$.
\end{lemma}
\begin{proof}
By \Cref{proposition:sigma_tangency},
the lines $P_1+P_2$ and $P_1+P_4$ are tangent to~$\iso$ in~$P_2$ and~$P_4$, respectively. The point $P_1$ cannot be on~$\iso$, hence, using the
action of $\mathrm{SO}_3(\mathbb{C})$, we can assume $P_1 = (1: 0: 0)$.
Since every element of $\mathrm{SO}_3(\mathbb{C})$ fixes~$\iso$ (and sends a tangent line to it into another tangent line to it), when we transform the point~$P_1$
into $(1: 0: 0)$, we transform the points~$P_2$ and~$P_4$ into, respectively,
the points $(0: \iii: 1)$ and $(0: -\iii: 1)$ (which are the common points to
$\iso$ and the tangent lines through~$P_1$).
Therefore, it is enough to study the
specific configuration of the points:
%
\begin{gather*}
P_1 = (1: 0: 0), \quad P_2=(0: \iii: 1), \quad P_3=(u_1, \iii u_2, u_2), \\
P_4 = (0: -\iii: 1), \quad P_5 = (v_1, -\iii v_2, v_2),
\end{gather*}
%
where $(u_1: u_2), (v_1: v_2) \in \p^1$.
In the $15\times 10$ matrix $\Phi(P_1, \dots, P_5)$ we can erase the
rows: $\phi_2(P_1)$ (which is a zero row), the row $\phi_1(P_2)$
(since $\phi_1(P_2)=\iii\phi_2(P_2)$) and, for similar reasons, the
rows: $\phi_1(P_3)$, $\phi_1(P_4)$ and $\phi_1(P_5)$.
Moreover, since the rank of the matrix $\Phi(P_1, P_2, P_3)$ is $5$,
we can also erase the row $\phi_3(P_3)$ and, for the same reason, the
row $\phi_3(P_5)$. The remaining matrix $M$ is a $8\times 10$ matrix.
It is not possible that all the order $8$ minors
of $M$ are zero: the ideal they generate, after saturations ensuring
that the points are distinct, is the whole ring.
\end{proof}

\begin{prop}
\label{proposition:rank_V}
Let $P_1, \dots, P_5$ be a $V$-configuration of
points. Then we have:
\begin{enumerate}
\item $8 \leq \text{rk}\ \Phi(P_1, \dots, P_5) \leq 10$\,;
\item $\text{rk}\ \Phi(P_1, \dots, P_5) \leq 9$ if and only if
$\delta_1(P_1, P_2, P_4) \cdot \delta_2(P_1, \dots, P_5) =0$\,;
\item $\text{rk}\ \Phi(P_1, \dots, P_5) = 8$ if and only if, one of
the following two conditions is satisfied:
%
\begin{itemize}
  \item the line $P_1+P_2$ is tangent to~$\iso$ in $P_2$ or $P_3$
and the line $P_1+P_4$ is tangent to~$\iso$ in $P_4$ or $P_5$,
\item $\delta_1(P_1, P_2, P_4) = 0$, \
$\overline{\delta}_1(P_1, P_2, P_3) = 0$,
\ $\overline{\delta}_1(P_1, P_4, P_5) = 0$\,.
\end{itemize}
%
\end{enumerate}
\end{prop}
\begin{proof}
%%% si basa sui file:
%%% rank_8_2_1_ii_0.sage e
%%% rank_8_1.sage
If the rank is $\leq 7$, from
\Cref{prop:condition3+1} applied to $P_1, P_2, P_3, P_4$ and $P_1, P_4, P_5, P_2$,
the lines $P_1+P_2$ and $P_1 + P_4$ are tangent to~$\iso$ (the first in $P_2$ or $P_3$ and the second in $P_4$ or $P_5$).
Then we get a contradiction from \Cref{lemma:special_case_rank_8}.
This shows the first item.

The second item follows from \Cref{lemma:d1d2}.

In the rest of the proof, by \Cref{invariance}, we distinguish two possibilities:
$P_1 = (1:\iii :0)$ and
$P_1 = (1: 0: 0)$.
Let $P_2, P_4 = (A_2: B_2: C_2), (A_4: B_4: C_4)$ and
$P_3 = u_1P_1+P_2$, $P_5 = v_1P_1+v_2P_4$ be the other points.
In the first case, we transform the matrix $M = \Phi(P_1, \dots, P_5)$ by elementary row and column operations into $\mathbb{I}_2 \oplus M'$, where $\mathbb{I}_2$ is the order~$2$ identity matrix; thus, $\mathrm{rk} M \leq 9$ if and only if $\mathrm{rk} M' \leq 7$.
The computations require several other strategies, but at the end we are
able to see that in this case (i.e.\ the case in which $P_1$ is on~$\iso$), the matrix $M$ cannot have rank less then $9$.
If $P_1 = (1: 0: 0)$, then the ideal of the order $9$ minors of
$\Phi(P_1, \dots, P_5)$ (after several manipulations and saturations)
gives that there a few possibilities which can be summarized by the
following conditions: either the line $P_1+P_2$ is tangent to~$\iso$ (in $P_2$ or $P_3$) and the line $P_1+P_4$ is tangent to~$\iso$ (in $P_4$ or $P_5$) or $\delta_1(P_1, P_2, P_4) = 0$,
\ $\overline{\delta}_1(P_1, P_2, P_3) = 0$,
\ $\overline{\delta}_1(P_1, P_4, P_5) = 0$.
To conclude the proof, we use \Cref{lemma:special_case_rank_8} and \Cref{lemma:d1d2}.
\end{proof}


\section{$\delta_1=0$ or $\delta_2=0$}

Questa parte va un po' meditata, per cercare una notazione coerente.@@
Forse la definizione \Cref{definition:matrix_conditions} va rimaneggiata.

Given a linear combination $\Lambda = a_0\lambda_0+\cdots+a_9\lambda_9$, we
denote by $v(\Lambda)$ the $10$-component vector given by
$(\lambda_0, \dots, \lambda_9)$. In particular, if $X=(x: y: z)$ and
$f(X) = a_0x^3+a_1x^2y+\cdots+ a_9z^3$ is a degree $3$ homogeneous polynomial
$v(f(X)) = (x^3, x^2y, \dots, z^3)$, while $v(\de_x(f)(X)) = v(\de_x)(f)(X) =%
(3x^2, 2xy, \dots, 0)$. With this notation if $P=(A: B: C)$, we get:
%
\begin{gather*}
\phi_1(P) = A\cdot v(f_x)(P)-B\cdot v(f_y)(P), \\
\phi_2(P) = A\cdot v(f_z)(P)-C\cdot v(f_y)(P),\\
\phi_3(P) = B\cdot v(f_z)(P)-C\cdot v(f_y)(P) 
\end{gather*}
%
Here we want to consider a $V$-configuration of points $P_1, \dots, P_5$
such that $\delta_1(P_1, P_2, P_4) = 0$ or $\delta_2(P_1, \dots, P_5) = 0$.
Moreover, we assume that the rank of the matrix $M$
of the conditions imposed by
the five points (which, by~\Cref{proposition:rank_V}, is at most~$9$)
is indeed $9$, hence there exists precisely one cubic curve $\cbc$ having
$P_1, \dots, P_5$ among its eigenpoints. Another immediate consequence of
the rank $9$ condition is that the
matrix $\Phi(P_1, \dots, P_5)$ has $9$ linearly independent rows, so
we can extract from it a  $9 \times 10$-submatrix
$\mathcal{H}$ whose order $9$ minors are not all zero. If
$\mathcal{H}_i$ is the order $9$ minor of
$\mathcal{H}$ given by erasing the $i$-th column ($i=1, \dots, 10)$, then,
from Cramer's rule, the polynomial defining $\cbc$ is given by
\[
 F(X) = \sum_{i=1}^{10}(-1)^i\det(\mathcal{H}_i)\cdot v(f)(X)_i
\]
(where $v(f)(X)_i$ is the $i$-th monomial of $v(f)(X)$
and $X=(x: y: z)$). In other words,
\[
F(X) = \det \left( \begin{array}{c} \mathcal{H}\\ v(f)(X)
 \end{array} \right)
\]
From the matrix $\mathcal{H}$ it is moreover possible to construct 
the three minors of (\ref{eq:def_matrix}) (i.e.\ the Geiser polynomials
(si chiamano cosi'?@@)), indeed we have:
\begin{prop}
\label{proposition:geiser1}
If $g_1, g_2, g_3$ are the three minors of the matrix 
\[
\left(
\begin{array}{ccc}
x & y & z \\
\de_xF(X) & \de_yF(X) & \de_z F(X)
\end{array}
\right)
\]
they can be computed by:
\[
\det \left( \begin{array}{c} \mathcal{H}\\ 
\phi_1(X)
\end{array} \right),\quad
\det \left( \begin{array}{c} \mathcal{H}\\ 
\phi_2(X)
\end{array} \right), \quad
\det \left( \begin{array}{c} \mathcal{H}\\ 
\phi_3(X)
\end{array} \right)
\]
\end{prop}
\begin{proof} We have:
\begin{eqnarray*}
g_1 & =  & x \cdot \de_y F(X)- y \cdot \de_xF(X) \\
& = & 
x\cdot \de_y \det \left( \begin{array}{c} \mathcal{H}\\ v(f)(X)
\end{array} \right) -y \cdot
\de_x \det \left( \begin{array}{c} \mathcal{H}\\ v(f)(X)
\end{array} \right) \\
& = & \det \left( \begin{array}{c} \mathcal{H}\\ x \cdot v(f_y)(X)-
y \cdot v(f_x)(X)
\end{array} \right)\\
& = & \left( \begin{array}{c} \mathcal{H}\\ 
\phi_1(X)
\end{array} \right)
\end{eqnarray*}
and similarly for $g_2 = x \cdot \de_z F(X)- z \cdot \de_xF(X)$
and $g_3 = y \cdot \de_z F(X)- z \cdot \de_yF(X)$.
\end{proof}


\begin{es} Consider the following five points in a $V$-configuration:
\begin{gather*}
p_1, \ p_2, \ p_4 = (2: -1: 1), \ (-1: 1: 3), \ (3: 6: -1), \\
p_3, \ p_5 = p_1+p_2,  \ p_1+p_4
\end{gather*}
They satisfy the condition $\delta_2(p_1, \dots, p_5) = 0$.
We consider the $9\times 10$ matrix $\mathcal{H}$ whose rows
(suitable rescaled) are:
$\phi_i(p_j)$ for $i = 1, 2$ and $j = 1, 2, 3, 4$ and $\phi_1(p_5)$:
\[
\mathcal{H} =
\left(\begin{array}{rrrrrrrrrr}
12 & 4 & -7 & 6 & 4 & 3 & -4 & 1 & 2 & 0 \\
-12 & 4 & -1 & 0 & 4 & -3 & 2 & 7 & -4 & 6 \\
-3 & 1 & 1 & -3 & 6 & 0 & -6 & -9 & -9 & 0 \\
-9 & 6 & -3 & 0 & 17 & -8 & -1 & -21 & -6 & -27 \\
0 & 1 & 0 & 0 & 0 & 4 & 0 & 0 & 16 & 0 \\
-12 & 0 & 0 & 0 & -31 & 0 & 0 & -56 & 0 & 48 \\
54& 63& 36& -108& -12& -9& 12& 2& -1& 0\\
27 & 36 & 36 & 0 & 21 & 48 & 108 & -17 & -36 & 9 \\
336& 328& 119& -588& -56& -33& 56& 7& -4& 0
\end{array}
\right)
\]
The cubic with $p_1, \dots, p_5$ as eigenpoints is given by
\[
\det \left(
\begin{array}{c}
\mathcal{H}\\
x^3 \ \ x^2y \ \ xy^2 \ \dots \  z^3
\end{array}
\right)
\]
whose value is (after a suitable rescaling):
\[
766x^3 - 1176x^2y + 417xy^2 - 128y^3 + 1488x^2z - 1722xyz
+ 597y^2z - 207xz^2 + 504yz^2 + 911z^3
\]
and the Geiser (@@) polynomials are:
\begin{eqnarray*}
g_1 & = & 392x^3 + 488x^2y - 656xy^2 + 139y^3 + 574x^2z + 594xyz - 574y^2z - 168xz^2 - 69yz^2\\
g_2 & = & 496x^3 - 574x^2y + 199xy^2 - 904x^2z + 1120xyz - 139y^2z - 81xz^2 + 574yz^2 + 69z^3\\
g_3 & = & 496x^2y - 574xy^2 + 199y^3 + 392x^2z - 416xyz + 464y^2z + 574xz^2 + 513yz^2 - 168z^3
\end{eqnarray*}
and can be computed either using their definition,
or~\Cref{proposition:geiser1}.
The common zeros of $g_1, g_2, g_3$ are the points $p_1, \dots, p_5$
and the points $p_6 = (21: 34: -8)$ and $p_7 = (19: 35: -9)$.
\end{es}

\begin{rmk}
In case the matrix $\Phi(P_1, \dots, P_5)$ has rank $8$, the family of
cubics with $P_1, \dots, P_5$ as eigenpoints is one dimensional, so the
above construction can be modified, by substituting one of the rows of
$\mathcal{H}$ with a random row of elements of the field~$K$.
\end{rmk}

Here we want to consider the following problem:
given five generic points in a $V$-configuration and imposing one
fo the two conditions
$\delta_1(P_1, P_2, P_4) = 0$ or $\delta_2(P_1, \dots, P_5) = 0$, assuming
again that the rank of $\Phi(P_1, \dots, P_5)$ is~$9$, what
can we say about all the eigenpoints of the corresponding cubics?


To construct five points $P_1, \dots, P_5$ which are in a $V$-configuration
and such that $\delta_1(P_1, P_2, P_4)= 0$ is quite easy: $P_1$
and $P_2$ can be taken in an arbitrary way, $P_4$ has to be chosen in such
a way that it satisfies the linear
equation $\scl{P_4}{P_2}\scl{P_1}{P_1}-\scl{P_4}{P_1}\scl{P_1}{P_2} = 0$
and $P_3$ and $P_5$ have to be chosen on the lines $P_1+P_2$ and $P_1+P_4$
respectively. In particular, the corresponding variety of cubic curves
has dimension $7$ (parametrized by an opes subset of
$\mathbb{P}^2\times \mathbb{P}^2 \times
\mathbb{P}^1\times  \mathbb{P}^1$).
The construction of a random example
of five points as above, gives a smooth cubic curve whose $7$ eigenpoints
do not have other collinearities (in addition to those of a
$V$-configuration).

The case in which the $V$-configuration satisfies the condition
$\delta_2(P_1, \dots, P_5) = 0$ is somehow different.

We define:
\begin{equation}
  \begin{split}
    U_1 & =  \langle P_1, P_2\rangle \left(\langle P_1, P_1\rangle
  \langle P_4,P_5\rangle - \langle P_1, P_4\rangle \langle P_1, P_5\rangle
  \right)\\
  U_2 & =  \langle P_1, P_2\rangle^2\langle P_4, P_5\rangle
  -\langle P_1, P_4\rangle \langle P_1, P_5\rangle \langle P_2, P_2\rangle
  \label{sst2}
  \end{split}
\end{equation}
in such a way that
$\delta_2(P_1, \dots, P_5) = U_1u_1+U_2u_2$. In order to
have a $V$-configuration satisfying $\delta_2(P_1, \dots, P_5) = 0$ we can
therefore fix $P_1, P_2, P_4$ arbitrarily, $P_5$ any point 
on the line $P_1+P_4$ and in this
way $P_3$ is determined by $u_1 = U_2$ and $u_2 = -U_1$. Note that,
since we assume the points distinct, $U_1$ and $U_2$ cannot be zero.
Also in this case
the subvariety of $\mathbb{P}^9$ given by the cubics with $5$
eigenpoints satisfying $\delta_2 = 0$ is of dimension $7$
(parametrized by an open subset of
$\mathbb{P}^2\times \mathbb{P}^2 \times \mathbb{P}^2\times \mathbb{P}^1$).
Also in this case, it is easy to construct examples, but now it turns out that
the $7$ eigenpoints of the obtained cubics satisfy the further
condition that also $P_6$ and $P_7$ are aligned with $P_1$. To explain
why, we proceed as follows.

Consider again a matrix $\mathcal{H}$ of rank $9$ extracted from
the martix $\Phi(P_1, P_2, P_3, P_4, P_5)$, where now $P_3 = U_2P_1-U_1P_2$
and the three polynomials $g_1, g_2, g_3$ contructed
in~(\ref{proposition:geiser1}). The condition
$\text{rk}\,(\mathcal{H}) = 9$ ensures that
$g_1, g_2$ and $g_3$ cannot be zero and are polynomials of degree
$3$ in $x, y, z$.
The computation shows that three polynomials have a common factor,
hence are of the form $\Omega G_1$, $\Omega G_2$, $\Omega G_3$, where
($\Omega$ is a polynomial in the variables
$A_1, B_1, C_1, A_2, B_2, C_2, A_4, B_4, C_4, v_1, v_2$ but
does not contain the variables $x, y, z$), hence $G_1, G_2, G_3$ are again
degree three polynomials in $x, y, z$, with coefficients in the variables
$A_1, \dots, C_4, v_1, v_2$. The common zeros are the eigenpoints of the
cubic $\cbc$.
%
\begin{rmk} For computational reasons, it is much more convenient
to postpone the substitution $u_1 = U_2$ and $u_2 = -U_1$ after the 
computation of the three polynomials constructed
in~\Cref{proposition:geiser1}.
\end{rmk}
%
\begin{prop}
\label{proposition:G_split}
 The polynomial
$F := C_1G_1-B_1G_2+A_1G_3$ splits into three factors $r_1$, $r_2$, $r_3$,
linear in $x, y, z$.
\end{prop}
\begin{proof}
%%%%%%%%%%%%%%%%%%%%
@@@@@@@@@@@@@@@@@@@@@@@\\
DA qui: 

 We know that $G_1, G_2, G_3$ admit the syzygy $z G_1 - y G_2 + x G_3 = 0$.
Then, due to the exactness of the Koszul complex of $(x,y,z)$ 
(see \cite[Chapter~17]{Eisenbud1995}),
we know that $G_1, G_2, G_3$ are the $2 \times 2$ minors of a matrix of the form
%
\[
 \left(
 \begin{array}{ccc}
  x & y & z \\
  H_1 & H_2 & H_3
 \end{array}
 \right)
\]
%
Let $E$ be the zero-set of $G_1, G_2, G_3$.
If $E$ is zero-dimensional and reduced,
then \cite{BGV} proved that in the ideal of~$E$ there exists a unique triple of generators of the form
%
\[
 \left(
 \begin{array}{ccc}
  x & y & z \\
  \partial_x f & \partial_x f & \partial_x f
 \end{array}
 \right)
\]
%
where $f$ is a ternary cubic.
We actually know that $E$ is zero-dimensional, because this can be checked by taking particular values in the $G_i$.
Hence we obtain that the $G_i$ are the $2\times 2$ minors of the previous matrix.

@@@@@@@@@@@@@@@@@\\
a qui probabilmente ora e' superfluo
%%%%%%%%%%%%%%%

Consider the Laguerre map
%
\[
 \gamma \colon \p^2 \setminus E \longrightarrow \p^2 =
 \p \bigl( \left\langle G_1, G_2, G_3 \right\rangle^\vee \bigr) \,.
\]
%
By construction, the map $\gamma$ maps a point~$P$ to the pencil of cubics passing through~$E$ and~$P$.
Whenever we have an aligned triple of eigenpoints and we pick $P \in \p^2$ that is aligned with them, then the pencil of cubics is given by the line together with the pencil of conics passing through the remaining four eigenpoints.
This means that if we have an aligned triple of eigenpoints, then any point on such a line is mapped by $\gamma$ to the same element.
Hence $\gamma$ contracts the line passing through the three aligned eigenpoints.
Another description of~$\gamma$ is given by $P \mapsto (G_1(P): G_2(P): G_3(P))$.
This can hence be written as $P \mapsto P \wedge \nabla(f)(P)$.
The image of~$P$ is then the intersection point of the two lines:
%
\[
 \left\{
 \begin{array}{l}
  A x + B y + Cz  = 0\\
  \partial_x f(P) + \partial_y f (P) + \partial_z f(P) = 0
 \end{array}
 \right.
\]
%
Since we know that there is a whole line that is contracted to the same point~$Q$,
we know that the first equation is a factor of the second.
Therefore we have $\left\langle Q, (x,y,z) \right\rangle = 0$, so $\left\langle Q, \nabla f (x,y,z) \right\rangle = 0$.
It then follows that $\left\langle P, (x, y, z) \wedge \nabla f(x,y,z) \right\rangle = 0$, which implies that $A G_1 + B G_2 + C G_3$
\\
@@@@
QUI C'E' un meno da qualche parte? \\
@@@\\
evaluates to zero for each $(x,y,z)$ such that $\left\langle Q, (x,y,z) \right\rangle= 0$, namely for each point in the line $P_1 P_2$.
This means that the line $P_1 P_2$ is contained in $E$.
By repeating this argument in the case $P_1 P_4$ we see that $E$ splits into three distinct lines.
\end{proof}

According to \Cref{proposition:G_split}, the polynomial
$C_1G_1-B_1G_2+A_1G_3$ splits into three factors $r_1$, $r_2$, $r_3$,
linear in $x, y, z$, which
correspond to the line $P_1+P_2$, the line $P_1+P_4$ and the line $P_6+P_7$.
The factorization of $F$ gives that the factor $r_3$ is:
\[
r_3 = 2yv_1A_1B_1^4A_2^2B_2A_4^3-2xv_1B_1^5A_2^2B_2A_4^3+\cdots
-2 yv_1A_1B_1^4A_2^2B_2A_4^3
\]
and is composed by almost $2000$ monomials. The relevant fact is that when we
evaluate $r_3$ on the coordinates of $P_1$, we get zero. Hence we have:
\begin{prop}
If we have five generic points $P_1, \dots, P_5$ which are in a
$V$-configuration and satisfy the condition $\delta_2(P_1, \dots, P_5) = 0$,
then the unique cubic determined by the condition that $P_1, \dots, P_5$
are eigenpoints, has also the two other eigenpoints $P_6$ and $P_7$
aligned with $P_1$.
\end{prop}

\section{Particular cases}
file \verb+contiCasoDegenere2.sage+ and file
\verb+conf_sigma12_sigma14.sage+

\section{The locus of cubic ternary forms with an aligned triple of eigenpoints}
\label{locus_one_alignment}

\begin{definition}
 We set $\mathcal{L} \subseteq \p^9$ to be the closure of the locus of classes of cubic forms $f$ having a reduced zero-dimensional eigenscheme $E(f)$ with an aligned eigentriple.
\end{definition}

\begin{theorem}
The variety~$\mathcal{L}$ is an irreducible hypersurface.
\end{theorem}

\begin{proof}
By the Gallet - Logar construction, for a a general fixed aligned triple of points, the set of cubic forms having such a triple as eigenpoints is a linear system of dimension 3. Since the variety of aligned triples is $5$-dimensional, this proves that $\mathcal{L}$ is a hypersurface.

For the irreducibility, we maybe need the argument on the Geiser map. Is there a simpler one?
\end{proof}

\begin{lemma}
\label{lemma:pencil_one_aligned}
 If $f$ and $g$ are general cubics, then any cubic in the pencil $\lambda f + \mu g$ for $(\lambda: \mu) \in \p^1$ has at most one aligned triple of eigenpoints.
\end{lemma}

This is a corollary of the following result.

\begin{prop}
    The locus of cubics that have at least two aligned triples of eigenpoints has codimension~$2$ in~$\p^9$.
\end{prop}
\begin{proof}
    Conti di Sandro.
\end{proof}

\begin{definition}
 We define $\Delta \subset \mathcal{L}$ to be the closure of the locus of cubics with at least two aligned triples of eigenpoints.
\end{definition}

\begin{prop}
  The variety~$\Delta$ has dimension~$7$ and it is the union of two irreducible components~$\Delta_1$ and~$\Delta_2$.
\end{prop}

In what follows, we shall use the following notation.
Denote by $M_1$, $M_2$ and $M_3$ the three minors of \Cref{eq:def_matrix} relative to a cubic form $f$.
The net of cubics, which base locus is the eigenscheme $E(f)$, will be denoted by $\Lambda_f = \langle M_1, M_2, M_3 \rangle$.
\begin{lemma}
\label{lemma:scroll}
 If $f$ and $g$ are general cubics, then
 %
 \[
   \mathcal{N} := \bigcup_{(\lambda : \mu) \in \p^1} \Lambda_{\lambda f + \mu g} \subset \p^9
 \]
 %
 is an embedding of a rational projective bundle and has degree~$3$.
\end{lemma}
\begin{proof}
Consider the projective bundle given by the family of planes
%
\[
{\mathcal P} := \{ \Lambda_{\lambda f + \mu g} \, : \, (\lambda: \mu)\in \p^1 \} \subset \p^1 \times \p^9
\]
%
Then $\mathcal{N}$ is the projection of~$\mathcal{P}$ on the second factor.
However, the map ${\mathcal P} \to {\mathcal N}$ contracts no subvariety of any plane of ${\mathcal P}$, so either it is an embedding or it contracts some horizontal curve. In the latter case, all the planes of the family should intersect in at least one point. In particular, the two nets $\Lambda_f$ and $\Lambda_g$ should have non-empty intersection.
If we denote by $M_1$, $M_2$ and $M_3$ the $2 \times 2$ minors relative to~$f$, and by $N_1$, $N_2$ and $N_3$ the ones relative to~$g$, the vectorial dimension of the linear span $\left\langle M_1, M_2, M_3, N_1, N_2, N_3 \right\rangle$ should be strictly less than $6$. This can be avoided, since such a condition corresponds to a proper closed subscheme of $\p^9 \times \p^9$.

It follows that if $f$ and $g$ are general enough, then $\mathcal{N}$ is a $3$-dimensional rational normal scroll in $\p(\left\langle M_1, M_2, M_3, N_1, N_2, N_3 \right\rangle) \cong \p^5$.
Being a variety of minimal degree, its degree is $5+1-3 = 3$.
\end{proof}

\begin{theorem}
The degree of $\mathcal L$ is equal to
\[
  \deg \ \mathcal L =  15.
\]
\end{theorem}

\begin{proof}
We start by observing that a reduced $0$-dimensional eigenscheme contains an aligned triple if and only if the net of cubics $\Lambda_f = \langle M_1, M_2, M_3 \rangle$ contains a cubic which splits in three lines, a so called \emph{triangle}. Moreover, if $f$ is general enough, we have exactly one aligned triple and the other $4$ points are in general position; in this case, the net $\Lambda_f$ contains exactly three triangles, namely the unions of the line passing through the aligned triple and the reducible conics through the $4$ points in general position.

To determine the degree of $\mathcal L$ we consider a general pencil of cubic forms $\lambda f + \mu g$, and we will compute the number of elements with associated net $\Lambda_{\lambda f + \mu g}$ containing a triangle.

To this aim, denote by ${\mathcal T} \subset \p^9$ the variety of triangles; it is a classical result that its dimension is $6$ and its degree is $15$,
see for instance \cite[Section 2.2.2]{3264}. We now consider the variety~${\mathcal N}$ from \Cref{lemma:scroll}.
\comment{
given by the union of the nets of cubics of the pencil
$$
{\mathcal N} = \bigcup_{(\lambda : \mu) \in \p^1} \Lambda_{\lambda f + \mu g} \subset \p^9.
$$
Observe that we can assume that ${\mathcal N}$ is an embedding of a rational projective bundle; indeed, it can be seen as an immersion of the $\p^2$-bundle over $\p^1$ given by the family the planes ${\mathcal P}=
\{\Lambda_{\lambda f + \mu g}\ : \ (\lambda:\mu)\in \p^1\} \subset \p^1 \times \p^9$. The map ${\mathcal P} \to {\mathcal N}$ contracts no subvariety of any plane of ${\mathcal P}$, so it is either an embedding or it contracts some horizontal curve. In the latter case, all the planes of the family should intersect in at least one point. In particular, the two nets $\Lambda_f$ and $\Lambda_g$ should have non-empty intersection.
If we denote by $M_1$, $M_2$ and $M_3$ the $2 \times 2$ minors relative to $f$, and by $N_1$, $N_2$ and $N_3$ the ones relative to $g$, the vectorial dimension of the linear span
$\langle M_1,M_2,M_3,N_1,N_2,N_3
\rangle$ should be strictly less than $6$. This can be avoided, since such a condition corresponds to a proper closed subscheme of $\p^9 \times \p^9$.

It follows that if ${\mathcal N}$ is general enough, it is a $3$-dimensional rational normal scroll in $\p(\langle M_1,M_2,M_3,N_1,N_2,N_3
\rangle) \cong \p^5$, and being a variety of minimal degree, its degree is $\deg {\mathcal N}=5+1-3=3$.
}
Note that, since each net containing a triangle, actually contains exactly $3$ of them, the number of nets of ${\mathcal N}$ containing some triangle is given by
%
\[
\frac {{\mathcal T} \cdot {\mathcal N}}{3} =\frac{{15} \cdot {3}}{3}=15 \,.
\]
%
This hence implies that $\deg {\mathcal L} = 15$.
\end{proof}

\begin{es}
The following pencil of cubic forms admits exactly $15$ cubics with an aligned triple of eigenpoints ..
\end{es}

\section{Further alignments}
\label{further_alignments}

\begin{theorem}
\label{theorem:possible_alignments}

\end{theorem}

\begin{table}
\label{table:all_alignments}
\caption{All possible configurations of $7$ points with alignments that can appear as eigenschemes of a ternary cubic form, provided that the eigenscheme is zero-dimensional and reduced.}
\centering
\begin{tabular}{|lll|}\hline
  num & collinear vertices & name\\ \hline
 1& [(1, 2, 3)] & "line" \\
 2& [(1, 2, 3), (1, 4, 5)] & "X shape"\\
 3& [(1, 2, 3), (1, 4, 5), (1, 6, 7)] & "star" \\
  & [(1, 2, 3), (1, 4, 5), (2, 4, 6)] & "triangle" \\
 4& [(1, 2, 3), (1, 4, 5), (1, 6, 7), (2, 4, 6)] & "triangle + altitude" \\
  & [(1, 2, 3), (1, 4, 5), (2, 4, 6), (3, 5, 6)] & "two X shapes" \\
 5& [(1, 2, 3), (1, 4, 5), (1, 6, 7),  & "two stars"\\
  & \phantom{[}(2, 4, 6), (2, 5, 7)] & \\
 6& [(1, 2, 3), (1, 4, 5), (1, 6, 7), & "triangle + three altitudes"\\
  & \phantom{[} (2, 4, 6), (2, 5, 7), (3, 4, 7)] & \\
 7& [(1, 2, 3),
   (1, 4, 5),
   (1, 6, 7),
   (2, 4, 6), & "Fano matroid" \\
  & \phantom{[} (2, 5, 7),
   (3, 4, 7),
   (3, 5, 6)] & \\ \hline
\end{tabular}
\end{table}

It is well-known that case (7) cannot be realized in zero characteristic (see \cite{Whitney1935}), therefore we will not consider it in our analysis.

\begin{description}
 \item[Case (1)]
 \item[Case (2)]
 We get two possible conditions:
 \begin{align}
  d_2 &: \langle P_1, P_1 \rangle\cdot \langle P_2, P_4\rangle -
  \langle P_1, P_2 \rangle \cdot \langle P_1, P_4 \rangle \label{rango9_1} \\
  d_3 &: \langle P_1,P_2 \rangle \cdot \langle P_1,P_3\rangle \cdot
  \langle P_4,P_5\rangle -\langle P_1,P_4\rangle \cdot \langle P_1,P_5\rangle
  \cdot \langle P_2,P_3\rangle  \label{rango9_2}
\end{align}
By bilinearity, one has that $d_2$ is equivalent to any of
\begin{equation}
\begin{aligned}
    & \langle P_1, P_1 \rangle\cdot \langle P_3, P_4\rangle -
  \langle P_1, P_3 \rangle \cdot \langle P_1, P_4 \rangle \,, \\
    & \langle P_1, P_1 \rangle\cdot \langle P_3, P_5\rangle -
  \langle P_1, P_3 \rangle \cdot \langle P_1, P_5 \rangle \,, \\
    & \langle P_1, P_1 \rangle\cdot \langle P_2, P_5\rangle -
  \langle P_1, P_2 \rangle \cdot \langle P_1, P_5 \rangle \,.
\end{aligned}
\end{equation}
From examples, we know that $d_3$ actually implies Case a star configuration as in Case (3).
 \item[Case (3)]

\end{description}

\section{The locus of forms with positive-dimensional eigenscheme}

\begin{prop}
Assume that $\dim \Eig{f} = 1$ and that the one-dimensional component of $\Eig{f}$ is a line~$L$.
Then $L \subseteq \mathrm{Sing}(f)$.
\end{prop}
\begin{proof}
Assume by contradiction that
%
\[
 L \subseteq {\overline{R(f)}}.
\]
%
We observe that if $L \subseteq V(f)$, then $L\subset \iso$, by Proposition ....
This is impossible since $\iso$ is irreducible.

So consider the scheme-theoretic intersection
%
\[
Z:= L \cap \iso,
\]
%
If $Z \subset R(f)$, then $Z \subset V(f)$ by Proposition \ref{pro: relation eigenpoints - tangent space to the isotroquadric}. Since $L \not \subset V(f)$, residually $L$ intersects $V(f)$ in a point. Since such a point is also an eigenpoint and it can't be regular, it is necessarily a singular point of $V(f)$. This gives a contradiction, since every line passing through a singular point has intersection multiplicity $\ge 2$ with $V(f)$.
Finally, assume that $Z = p_1 + p_2$ as a divisor on~$\iso$, with $p_1 \in R(f)$ and $p_2\in {\overline {R(f)}} \setminus R(f)$. The necessarily $p_2 \in \mathrm{Sing}(f)$, so that residually to~$Z$ the line~$L$ intersects~$V(f)$ in~$p_2$. Therefore $V(f)$ has a singular point $p_2$ on $\iso$ and $V(f)$ intersect tangentially~$\iso$ in~$p_1$.
The examples show that this never happens ...
\end{proof}

\begin{prop}\label{p2}
Let $C = V(f) \subset \p^2$ be a cubic curve.
Assume that $\dim \Eig{f} = 1$ and that the $1$-dimensional component is a line~$L$.
Then the residual subscheme $Z := \mathrm{Res}_L \bigl( \Eig{f} \bigr)$ in~$\Eig{f}$ with respect to~$L$ is zero-dimensional of degree~$3$ if $\dim \Lambda = 2$, or of degree~$4$, if $\dim \Lambda = 1$.
\end{prop}
\begin{proof}
Since $G_1$, $G_2$ and $G_3$ have a common linear component~$L$, by writing
\[
G_i=L \ H_i, \quad i=1,2,3
\]
we have that the residual subscheme $Z = \mathrm{Res}_L \bigl( \Eig{f} \bigr)$ is a quasi complete intersection
determined by the polynomials $H_1$, $H_2$ and $H_3$; indeed, we recall that $\mathcal{I}_{Z, \p^2} = \mathcal{I}_{\Eig{f}), \p^2} (-1)$.

Let us assume $\dim \Lambda = 2 = \dim \p \bigl( \left\langle H_1,H_2,H_3 \right\rangle \bigr)$.
So we have an exact sequence
\[
 0 \to \mathcal{G} \to \oo_{\p^2} (-2) \oplus \oo_{\p^2} (-2) \oplus \oo_{\p^2} (-2) \to \mathcal{I}_{Z,\p^2} \to 0 \,,
\]
where $\mathcal{G}$ is a rank two reflexive sheaf by \cite[Proposition 1]{Hartshorne1980}.
But on a smooth surface reflexive implies locally free (see \cite[Example~1.1.6]{Huybrechts2010}),
so $\mathcal{G}$ is a rank two vector bundle.

Next we observe that the two independent
syzygies between the generators $G_i$ give rise to the syzygies:
\[
x_2\ H_1 - x_1\ H_2 + x_0\ H_3=0, \qquad \partial_2 f\ H_1 - \partial_1 f \ H_2 +\partial_0 f \ H_3=0,
\]
which occur in degrees $3$ and $4$. We claim that the two relations are independent, again,
otherwise the $G_i$'s would be identically zero. So we apply \cite[Proposition~12]{Ellia2020}, and we have that $\mathcal {G}$ splits. Observe that as $c_1(\mathcal {I}_{Z,\p^2})=0$, we have $c_1(\mathcal {G})=-6$. Moreover, there is no syzygy in degree $2$, since
otherwise the $H_i$ would be linearly dependent, so the $G_i$ would belong to a pencil, which we excluded.

It follows that the splitting of $\mathcal{G}$ is of the form
%
\[
\mathcal{G} \cong \oo_{\p^2} (-3) \oplus \oo_{\p^2} (-3) \,.
\]
%
We get the free resolution of the ideal sheaf:
%
\[
0\to \oo_{\p^2} (-3)\oplus \oo_{\p^2} (-3) \to 3\oo_{\p^2} (-2)\to \mathcal {I}_{Z,\p^2} \to 0 \,.
\]
%
Moreover
%
\[
c_2 (\mathcal{I}_{Z,\p^2} ) = c_2 \bigl( 3\oo_{\p^2}(-2) \bigr) - c_2 (\mathcal{G}) = 12 - 9 = 3 \,,
\]
%
and $h^0 \bigl( \mathcal{I}_{Z,\p^2}(1) \bigr) = 0$.

Finally, assume $\dim \Lambda =1 =\dim \p \bigl( \left\langle H_1, H_2, H_3 \right\rangle \bigr)$.
Then $Z = \mathrm{Res}_L \bigl( \Eig{f} \bigr)$ is the complete intersection of two conics.
\end{proof}


\begin{es}
Consider
\[
f(x_0,x_1,x_2)=x_0^2 (x_1 - x_2)
\]
(Example~\eqref{item: ternary cubic, no regular eigenpoints} with $t=i$).
We have $L=x_0$,
\[
H_1=x_0^2-2x_1^2+2x_1 x_2, \quad H_2= 2x_2^2-x_0^2-2x_1 x_2, \quad H_3= 2x_1x_2-2x_2^2-x_0x_1 \,.
\]
The two syzygies in degree~$3$ are:
\[
x_2 \, H_1 - x_1 \, H_2 + x_0 \, H_3 = 0, \quad x_0 \, H_2 + 2(x_1+x_2) \, H_3 + x_0 \, H_1 = 0.
\]
Finally, $Z= \{ (0:1:1),(2:1:-1),(-2:1:-1)\}$. Observe that one point is on the singular line $x_0=0$
and that the Jacobian scheme is non reduced, it consists of a line with an embedded point.
\end{es}

\begin{rmk}
Assume $\dim E(f) =1$ and that the $1$-dimensional component is an irreducible conic $\tilde Q$. Let $Q$ be the isotropic quadric. Then we have the following possible cases:
\begin{itemize}
\item $\tilde Q = Q$. It seems that this never happens.
\item set $Z:= \tilde Q \cap Q$. Then $Z\subset R(f)$ , so $V(f)$ is tangent to $Q$ in four points.
\item $Z$ consists of one singular point for $V(f)$ and one or two regular eigenpoints.
\item $Z$ consists of two singular points for $V(f)$.
\item $Z$ consists of a triple point for $V(f)$, with one common tangent with $Q$.
\end{itemize}
\end{rmk}

\bibliographystyle{amsalpha}
\bibliography{ooms}

\end{document}
