\documentclass[a4paper, 11pt, reqno]{amsart}

\setlength{\parindent}{.4 in}
\setlength{\textwidth}{6.5 in}
\setlength{\topmargin} {0.2 in}
\setlength{\evensidemargin}{0 in}
\setlength{\oddsidemargin}{0 in}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathtools}
\usepackage{lmodern}
\usepackage[english]{babel}
\usepackage{booktabs}
\usepackage{float}
\usepackage{enumitem}
\usepackage{tikz,pgf}
\usetikzlibrary{calc}
\usetikzlibrary{through}
\usepackage{tkz-euclide}
\usepackage{palatino}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}
\usepackage[nameinlink]{cleveref}

\theoremstyle{plain}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{prop}[lemma]{Proposition}
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{theoremintro}{Theorem}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{conjecture}[lemma]{Conjecture}
\newtheorem{fact}[lemma]{Fact}
\newtheorem{assumption}[lemma]{Assumption}
\newtheorem*{reduction}{Reduction}
\theoremstyle{definition}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{es}[lemma]{Example}
\newtheorem*{notation}{Notation}
\newtheorem{rmk}[lemma]{Remark}

\tikzstyle{point}=[circle, draw, fill=black, inner sep=0pt, minimum size=4pt]
\tikzstyle{line}=[line width=1.5pt, black!70!white]

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\p}{\mathbb{P}}
\newcommand{\sL}{\mathcal{L}}
\newcommand{\sU}{\mathcal{U}}
\newcommand{\sV}{\mathcal{V}}
\newcommand{\sF}{\mathcal{F}}
\newcommand{\de}{\partial}
\newcommand{\oo}{\mathcal{O}}

\newcommand{\iii}{\textit{i}\,}
\newcommand{\codim}{\mathrm{codim}}
\newcommand{\rk}{\ensuremath{\mathrm{rk}}}

\newcommand{\iso}{\mathcal{Q}_{\mathrm{iso}}}

\newcommand{\SO}{\operatorname{SO}}
\newcommand{\Eig}[1]{\mathcal{E}\!\left( {#1} \right)}
\newcommand{\Sym}{\operatorname{Sym}}

\newcommand{\scl}[2]{\left\langle {#1}, {#2} \right\rangle}

\newcommand\scalemath[2]{\scalebox{#1}{\mbox{\ensuremath{\displaystyle #2}}}}

\newcommand{\blue}[1]{{\color{blue}[#1]}}


\title{Eigenpoint collinearities of plane cubics}
\author[Valentina Beorchia]{Valentina Beorchia$^{\circ}$}
\address[\textsc{Valentina Beorchia}]{University of Trieste,
Department of Mathematics, Informatics and Geosciences,
Via Valerio 12/1, 34127 Trieste, Italy}
\email{beorchia@units.it}
\thanks{$^{\circ}$The researcher is a member of ``Gruppo Nazionale per le Strutture Algebriche, Geometriche e le loro Applicazioni'', INdAM. She is partially supported by MUR funds: PRIN project GEOMETRY OF ALGEBRAIC STRUCTURES: MODULI, INVARIANTS, DEFORMATIONS, PI Ugo Bruzzo, Project code: 2022BTA242.}
\author[Matteo Gallet]{Matteo Gallet$^{\diamond}$}
\address[\textsc{Matteo Gallet}]{University of Trieste,
Department of Mathematics, Informatics and Geosciences,
Via Valerio 12/1, 34127 Trieste, Italy}
\email{matteo.gallet@units.it}
\thanks{$^{\diamond}$The researcher is a member of ``Gruppo Nazionale per le Strutture Algebriche, Geometriche e le loro Applicazioni'', INdAM.}
\author[Alessandro Logar]{Alessandro Logar}
\address[\textsc{Alessandro Logar}]{University of Trieste,
Department of Mathematics, Informatics and Geosciences,
Via Valerio 12/1, 34127 Trieste, Italy}
\email{logar@units.it}
\date{}

\linespread{1.15}
\setlength{\parindent}{0pt}
\setlength{\parskip}{.25em}

\begin{document}

\begin{abstract}
 Given a homogeneous polynomial in $n+1$ variables, the fixed points of the map from $\p^n$ to itself defined by its gradient are called the \emph{eigenpoints} of the polynomial. We focus on cubic polynomials in three variables, whose eigenpoints have already been object of study, and analyze configurations of eigenpoints that admit one or more alignments. We give a classification of all possible configurations with alignments: this is accomplished using both geometric characterizations of these situations and an extensive use of computer algebra.
\end{abstract}

\maketitle


\section{Introduction}
\label{introduction}

Tensors are natural generalizations of matrices in higher dimension.
Important notions related to matrices, as for example those of \emph{rank} and of \emph{eigenvector},
can be generalized to tensors and may provide information about them.
Since roughly $20$ years ago, eigenvectors of tensors have been object of study from various points of view.
Our interest will be focused on a specific class of symmetric tensors, namely cubic ternary forms, and on a specific geometric configuration of their eigenvectors.
Before delving into the more technical aspects of our problem,
we begin with a short resume of the existing literature on eigenvectors of tensors.

\subsection*{Eigenvectors of tensors in the literature}
In 2005, Lim \cite{Lim} and Qi \cite{Qi} independently introduced the notions of eigenvalues and eigenvectors for tensors.
Since then, these two concepts have been of interest in several applications,
as in the study of hypergraphs and dynamical systems;
see \cite[Section~4]{QZ} for an introduction or \cite{GMV} for recent developments.
Eigenvectors of tensors also play an important role in the best rank-one approximation problem,
which is relevant in data analysis and signal processing. 
Indeed, by Lim's Variational Principle \cite{Lim}, given a symmetric tensor~$f$, the critical rank-one symmetric tensors for~$f$ are exactly of the form~$v^d$, where $v$ is an eigenvector of~$f$.
This has applications in low-rank approximation of tensors (see \cite{OttSod}) as well as maximum likelihood estimation in algebraic statistics.
For another application in optimization, consider the problem of maximizing a polynomial function~$f$ over the unit sphere in~$\R^{n+1}$: 
the eigenvectors of the symmetric tensor~$f$ are critical points of this optimization problem.


The geometry of eigenvectors is intimately related with the Waring decomposition of a polynomial, 
corresponding to the symmetric Tucker decomposition of the associated symmetric tensor, 
as clarified in~\cite{DOT} and in~\cite{Ott}. 
Indeed, any best rank~$k$ approximation of a symmetric tensor, when it exists, lies in the linear space, 
called \emph{critical space}, spanned by the rank~$1$ tensors of the type $v_i^{\otimes d}$, where $v_i$ varies among all the eigenvectors. 
Therefore, a deep comprehension of the geometry of eigenvectors will
go towards an improvement on the insight of low rank approximation problems.
A first evidence in this direction is given by the so called ODECO (''orthogonally decomposable``) tensors, 
that is, symmetric tensors~$T$ that admit a decomposition
$T = \sum _{i=0}^n v_i ^{\otimes d}$ with $v_0, \dotsc, v_n$ an orthogonal family (see \cite{Rob, BDHE} for further details);
the $v_i$'s turn out to be eigenvectors of~$T$.
Finally, in \cite{OO} Oeding and Ottaviani employ eigenvectors of tensors in an algorithm to compute Waring decompositions of homogeneous polynomials.
Ottaviani points out in \cite[Section~8]{Ottaviani24} the close connection between eigenvectors of specific symmetric tensors and the so-called L\"uroth quartics. 

\subsection*{Generalities on eigenschemes}
When dealing with eigenvectors, as observed in \cite{ASS},
it is not restrictive to focus on partially symmetric tensors, that is elements of~$\Sym^{d-1}\C^{n+1} \otimes \C^{n+1}$.
If we choose a basis for~$\C^{n+1}$, then we can identify a partially symmetric tensor $T \in \Sym^{d-1}\C^{n+1} \otimes \C^{n+1}$ with a tuple of homogeneous polynomials $(g_0, \dotsc, g_n)$. An \emph{eigenvector} of~$T$ is a nonzero vector $v \in \C^{n+1}$ such that
%
\begin{equation}
\label{eq:eigenvector_definition}
  \bigl( g_0(v), \dotsc, g_n(v) \bigr) = \lambda v
\end{equation}
%
for some constant~$\lambda \in \C$.
Since the property in \Cref{eq:eigenvector_definition} is preserved under scalar multiplication, 
it is natural to consider the rational map $T \colon \p^n \dashrightarrow \p^n$ defined by $T(P) = \bigl(g_0(P): \ldots: g_n(P)\bigr)$ 
and regard eigenvectors as points in~$\p^n$; hence the name \emph{eigenpoints} instead of eigenvectors.
When a tensor is symmetric, it can be identified with a homogeneous polynomial, also called a homogeneous form.
As a partially symmetric tensor, a homogeneous form~$f$ is identified by its gradient $\nabla f = (\de_0 f, \dotsc, \de_n f)$ and so its eigenpoints are the fixed points of the polar map $\nabla f \colon \p^n \dasharrow \p^n$.

The condition in \Cref{eq:eigenvector_definition} can be rephrased as
%
\begin{equation}
\label{eq:def_matrix_general}
  \rk
  \begin{pmatrix}
    x_0 & x_1 & \dots & x_n \\
    g_0 & g_1 & \dots & g_n
  \end{pmatrix} \le 1 \,,
\end{equation}
%
so that the set of eigenpoints can be given a natural structure of determinantal scheme, which is commonly called \emph{eigenscheme}, denoted $\Eig{T}$.

The dimension and degree of a general eigenscheme are settled by the following result (see \cite[Theorem 2.1]{CartSturm}, \cite{ASS}, \cite{OO}, and \cite[Equation~5.2]{Abo}):

\begin{theoremintro}
\label{thm:nonempty}
Let $d \ge 2$ and let $T \in (\C^{n+1})^{\otimes d}$.
If $\dim \Eig{T}=0$, then 
%
\[
  \deg \Eig{T} =
  \frac{(d-1)^{n+1}-1}{d-2} =
  \sum_{i=0}^{n} (d-1)^i \,.
\]
%
In the matrix case, that is $d = 2$, we use the formula on the right, which evaluates to $n+1$.

Moreover, if $T$ is general, then its eigenscheme is $0$-dimensional and reduced.
\end{theoremintro}

\subsection*{Eigenschemes of cubic ternary forms}
% \textbf{@@ una frase cosi' che vi sembra?}
% %%%
% %%%%
% The problem we consider in this paper can be simplified in the following way.
% Consider a degree three homogeneous polynomial 
% $f \in \mathbb{C}[x, y, z]$ 
% (hence a cubic curve $C$ in $\mathbb{P}^2$).
% We can associate to $f$ an ideal $J_e$ which is given by the
% order two minors of the matrix $((f_x, f_y, f_z), (x, y, z))$. The variety
% $V(J_e)$ is the locus of the eigenpoints of $f$. In general, $V(J_e)$ is
% given by $7$ points, but there are cases in which $V(J_e)$ is a
% one-dimensional variety.
% In the present paper we characterize the cubics for which $V(J_e)$ is one-dimensional
% (and in these cases we give a descritpion of $V(J_e)$) and, when
% $V(J_e)$ is given by $7$ distinct points, we classify all the possible
% configurations that these points can assume w.r.t.\ their collinearities. 

In the present paper, we focus on the geometry of eigenschemes of $3\times 3 \times 3$ symmetric tensors, 
that is the case $n=2$ and~$d=3$ in \Cref{thm:nonempty}.
In other words, as we have clarified, we consider a homogeneous polynomial $f \in \C[x, y, z]$ of degree~$3$ and associate to it the scheme~$\Eig{f}$ determined by the ideal of $2 \times 2$ minors of the matrix
%
\[
\begin{pmatrix}
    x & y & z \\
    \partial f / \partial x & \partial f / \partial y & \partial f / \partial z
\end{pmatrix} \,.
\]
%
By \Cref{thm:nonempty}, a general $f$ satisfies $\deg \Eig{f}=7$.
Moreover, in general $\Eig{f}$ is constituted by $7$ reduced points, but there are cases when it is one-dimensional, or non-reduced. 

A geometric characterization of eigenschemes of ternary forms is given by the following results.

\begin{theoremintro}(\cite[Theorem 5.1]{ASS})
\label{thm:no_six_conics}
A configuration of seven points in~$\p^2$ is the eigenconfiguration
of a $3 \times 3 \times 3$-tensor if and only if no six of the seven points lie on a conic.
\end{theoremintro}

\begin{theoremintro}(\cite[Theorem 5.7]{BGV})
\label{thm:general_no_d_collinear}
Let $d \ge 3$. The eigenscheme~$\Eig{f}$ of a general homogeneous polynomial $f \in \C[x, y, z]_d$ contains no $d$ collinear points.
\end{theoremintro}

However, there are several examples of cubic polynomials, whose eigenscheme contains one or more triples of aligned points, as for instance the Fermat cubic polynomial.

\subsection*{Our results}
The goal of this paper is to classify all the situations when we have one or more triples of aligned points inside a $0$-dimensional reduced eigenscheme of a ternary cubic form.
More precisely, we
%
\begin{itemize}
    \item determine conditions imposed on the eigenpoints and on ternary cubic forms by the presence of aligned eigenpoints;
    \item compute the degree of the closure of the locus in~$\p^9$ of ternary cubic forms that admit an aligned triple of eigenpoints;
    \item characterize the situations yielding a one-dimensional eigenscheme;
    \item classify all possible combinatorial configurations of aligned eigenpoints when the eigenscheme is $0$-dimensional and reduced.
\end{itemize}
%
All the conditions on the eigenpoints will be expressed in terms of their homogeneous coordinates, and the equations that we will find have to be though as equations in $(\p^2)^7=\p^2 \times \p^2 \times \dotsb \times \p^2$.

Finally, we would like to point out that our approach gives some hints towards the problem of finding equations for the locus of $7$-uple of points, which form an eigenscheme of some symmetric tensor; this is an open question, and it has been posed in \cite{ASS} and \cite{Ottaviani24}.

\subsection*{Content of the paper}
Here is a description of the content of each section.
\Cref{invariance} describes the action of special orthogonal matrices on ternary forms, which we repeatedly use in our analyses since it allows us to "pin down" one of the eigenpoints to two specific cases, thus greatly speeding up the symbolic computations we rely on.
\Cref{conditions} begin our analysis by examining what conditions on the space of ternary cubics are imposed by the existence of aligned eigenpoints. In particular, three kinds of conditions, named~$\delta_1$, $\bar{\delta}_1$, and~$\delta_2$, emerge and play a prominent role in describing configurations of two triples of aligned eigenpoints that share a point, called $V$-configurations.
\Cref{rank_9} and \Cref{rank_8} set the spotlight on $V$-configurations, exhausting all the possible special cases that may appear.
\Cref{locus_one_alignment} introduces the locus of ternary cubics with a single alignment in the eigenscheme, proves its irreducibility and computes its degree.
\Cref{positive_dim} classifies positive-dimensional eigenschemes of ternary cubics.
\Cref{further_alignments} examines all the possible configurations of alignments in a $0$-dimensional reduced eigenscheme and establishes which may arise and which not.

\subsection{Notation and preliminaries}
@@ Inserire notazioni e discussione dell'omogeneità delle formule.

We will work in $\p^2_\C$, endowed with the canonical projective coordinate system.

We shall use vectorial notions for projective points. 
For instance, for $P_1=(A_1:B_1:C_1)$ and $P_2=(A_2:B_2:C_2)$ we shall denote by
%
\[
  s_{12} = \scl{P_1}{P_2} = A_1 A_2 + B_1 B_2 + C_1 C_2 \,.
\]
%
Such an expression has to be intended as a bihomogeneous polynomial in the coordinates of $P_1$ and $P_2$. If we think of the $A_i$, $B_j$ and $C_k$ as indeterminates, the zero set of~$s_{12}$ is well defined in $\p^2 \times \p^2$

Similarly, by $P_1 \times P_2$ we shall denote the {\it cross product} of the representing vectors
%
\[
  P_1 \times P_2 = 
  (B_1 C_2 - C_1 B_2, \, C_1 A_2 - A_1 C_2, \, A_1 B_2 - B_1 A_2) \,.
\]
%
The resulting projective point is well defined. Such a notation will be used, for instance, in \Cref{definition:delta1}.

In a similar fashion, whenever we .... @@ Inserire frase sul fatto che i punti sono ben definiti.

We recall that, given a homogeneous form $f \in \C[x,y,z]_d$ of degree~$d$, the \emph{eigenscheme}~$\Eig{f} \subset \p^2$ of~$f$ is the determinantal scheme defined by the $2 \times 2$ minors of the matrix
%
\begin{equation}
\label{eq:def_matrix}
  \begin{pmatrix}
    x & y & z \\
    \de_x f & \de_y f & \de_z f
  \end{pmatrix}.
\end{equation}
%
Since the eigenschemes of two proportional homogeneous forms are the same,
if $C \subset \p^2$ is a curve, then we can define with no ambiguity the eigenscheme of~$C$, denoted~$\Eig{C}$, as the eigenscheme of any homogeneous form defining~$C$.

Whenever the eigenscheme of a ternary form is $0$-dimensional, it never contains $4$ or more aligned points.
To see this, we recall that when an eigenscheme of a ternary form is  $0$-dimensional, it is somehow ``general'' with respect to conics.

\begin{corollary}
\label{corollary:general_no_triple}
As a consequence of \Cref{thm:no_six_conics},
no $0$-dimensional eigenscheme~$\Eig{f}$ for $f \in \C[x,y,z]_d$ contains $4$ or more aligned points.
Moreover, if $\Eig{f}$ contains two triples of aligned points, those must share a point.
\end{corollary}

By \Cref{thm:general_no_d_collinear}, the eigenscheme of a general ternary cubic form has no aligned triples of points. 
Therefore, it can be expected that ternary cubic forms with one or more aligned triples of eigenpoints should satisfy particular geometric constraints.

Understanding these situations requires a careful analysis of the conditions on ternary cubic forms imposed by the existence of aligned eigenpoints.
To simplify these analyses, we clarify how we can benefit from the action of special orthogonal matrices.

\section{Invariance under the action of orthogonal matrices}
\label{invariance}

In what follows, it will be useful to fix particular coordinates for points and lines related to eigenschemes.
To do that, we employ a property of invariance of eigenschemes with respect to the action of the following group.

\begin{definition}
We define $\SO_3(\C)$ to be the complexification of the group of special orthogonal real matrices, namely
%
\[
  \SO_3(\C) :=
  \bigl\{
    M \in \mathrm{GL}_3(\C) \, \mid \,
    M M^t = I_3 \ \text{and} \ \det(M) = 1
  \bigr\} \,.
\]
%
The group $\SO_3(\C)$ acts on $\C^3$ by matrix multiplication:
%
\[
  \begin{array}{ccc}
    \SO_3(\C) \times \C^3 & \rightarrow & \C^3 \\
    (M, v) & \mapsto & Mv
  \end{array}
\]
%
Since all the elements of~$\SO_3(\C)$ are invertible, the latter action descends to an action on $\p^2$.

Moreover, the group~$\SO_3(\C)$ acts also on ternary forms via
%
\[
  M \cdot f (x,y,z) = f(M^{-1} \cdot \prescript{t} {}( x \ y \ z ) ).
\]
%
\end{definition}

\begin{prop}
\label{two_orbits}
The action of~$\SO_3(\C)$ on $\p^2$ has two orbits:
%
\begin{align*}
  \mathcal{O}_1 &:=
  \bigl\{
    P \in \p^2 \, \mid \,
    P = (a:b:c) \ \text{with} \ a^2 + b^2 + c^2 = 0
  \bigr\} \\
  \mathcal{O}_2 &:= \p^2 \setminus \mathcal{O}_1
\end{align*}
%
A representative for~$\mathcal{O}_1$ is~$(1:\iii:0)$ and a
representative for~$\mathcal{O}_2$ is~ $(1:0:0)$.
The orbit~$\mathcal{O}_1$ is the set of points of the so-called \emph{isotropic conic}, denoted~$\iso$.
\end{prop}
\begin{proof}
Suppose that $P \in \p^2$ and $P = (a:b:c)$ with $a^2 + b^2 + c^2 = 0$.
We produce a matrix $M \in \SO_3(\C)$ such that $M \left(\begin{smallmatrix} 1 \\ \iii \\ 0 \end{smallmatrix}\right)$ and $\left(\begin{smallmatrix} a \\ b \\ c \end{smallmatrix}\right)$ are proportional.
Up to relabeling the coordinates, we can suppose that $a \neq 0$.
Hence, by rescaling the coordinates of~$P$, we have $P = (1: b: c)$ with $b^2 + c^2 = -1$.
One can check that the matrix
%
\[
  M :=
  \begin{pmatrix}
    -1 & 0 & 0 \\
    0 & \iii b & -\iii c \\
    0 & \iii c & \iii b
  \end{pmatrix}
\]
%
satisfies the requirements.

Now, suppose that $P \in \p^2$ and $P = (a:b:c)$ with $a^2 + b^2 + c^2 \neq 0$.
Up to rescaling, we can suppose that $a^2 + b^2 + c^2 = 1$.
Again, we produce a matrix $M \in \SO_3(\C)$ such that $M \left(\begin{smallmatrix} 1 \\ 0 \\ 0 \end{smallmatrix}\right)$ and~$\left(\begin{smallmatrix} a \\ b \\ c \end{smallmatrix}\right)$ are proportional.
First of all, suppose that $b^2 + c^2 \neq 0$ and let $\omega$ be a root of the polynomial $t^2 - (b^2 + c^2)$ in $\C[t]$.
Then, the matrix
%
\[
  M :=
  \begin{pmatrix}
    a & \omega & 0 \\
    b & -\frac{ab}{\omega} & \frac{c}{\omega} \\
    c & -\frac{ac}{\omega} & -\frac{b}{\omega}
  \end{pmatrix}
\]
%
satisfies the requirements.
With the same technique, if $a^2 + c^2 \neq 0$, we can produce a matrix $M \in \SO_3(\C)$ that maps $\left(\begin{smallmatrix} 0 \\ 1 \\ 0 \end{smallmatrix}\right)$ to $\left(\begin{smallmatrix} a \\ b \\ c \end{smallmatrix}\right)$; similarly, when $a^2 + b^2 \neq 0$, we can map $\left(\begin{smallmatrix} 0 \\ 0 \\ 1 \end{smallmatrix}\right)$ to $\left(\begin{smallmatrix} a \\ b \\ c \end{smallmatrix}\right)$.
Since $\left(\begin{smallmatrix} 1 \\ 0 \\ 0 \end{smallmatrix}\right)$, $\left(\begin{smallmatrix} 0 \\ 1 \\ 0 \end{smallmatrix}\right)$, and $\left(\begin{smallmatrix} 0 \\ 0 \\ 1 \end{smallmatrix}\right)$ are all $\mathrm{SO}_3(\C)$-equivalent, the only case to consider is when
%
\[
  b^2 + c^2 = a^2 + c^2 = a^2 + b^2 = 0 \,,
\]
%
which, however, can never occur.
\end{proof}

The following result is well known; we recall it for the sake of completeness.

\begin{prop}
Let $M \in \SO_3(\C)$ and let $f$ be a ternary cubic.
Let $P = (A: B: C)$ be a point in~$\p^2$.
Then we have
%
\[
  P \in \Eig{f} \iff M \cdot \prescript{t} {}(A \ B \ C) \in \Eig{M \cdot f} \,.
\]
%
\end{prop}
\begin{proof}
In this proof, for convenience,
we consider the transpose of the defining matrix of an eigenscheme.
A point~$P = (A: B: C)$ is an eigenpoint for~$f$ if and only if
%
\begin{equation}
\label{eq:def_matrix_M}
  \mathrm{rk}
  \begin{pmatrix}
    A & \de_x f(P) \\
    B & \de_y f(P) \\
    C & \de_z f(P)
  \end{pmatrix}
  = 1 \,
 \quad \text{or, equivalently} \quad
  \mathrm{rk} \ M \cdot 
  \begin{pmatrix}
    A & \de_x f(P) \\
    B & \de_y f(P) \\
    C & \de_z f(P)
  \end{pmatrix}
  = 1 \,.
\end{equation}
%
By setting $\prescript{t} {}(A' \ B' \ C' ) := M \cdot \prescript{t} {}(A \ B \ C) $ and~$Q := (A':B':C')$, we have that \Cref{eq:def_matrix_M}
is equivalent to
%
\begin{equation}
\label{eq:transformed}
  \rk
  \begin{pmatrix}
    A' & \\
    B' & M \cdot \nabla f (P) \\
    C' & \\
  \end{pmatrix}
  = 1 \,.
\end{equation}
%
Now we consider the polynomial $M \cdot f$ and we observe that the chain rule gives
%
\begin{gather*}
  \partial_x (M\cdot f) = \partial_x \bigl( f(M^{-1} \ \prescript{t} {} (x \ y \ z)) \bigr) = \prescript{t} {}(M^{-1})^{(1)}(\nabla f) \bigl( M^{-1}\ \prescript{t} {} (x \ y \ z) \bigr) \,, \\
  \partial_y (M\cdot f) = \partial_y \bigl( f(M^{-1} \ \prescript{t} {} (x \ y \ z)) \bigr) = \prescript{t} {}(M^{-1})^{(2)}(\nabla f) \bigl( M^{-1}\ \prescript{t} {} (x \ y \ z) \bigr) \,, \\
  \partial_z (M\cdot f) = \partial_z \bigl( f(M^{-1} \ \prescript{t} {} (x \ y \ z)) \bigr) = \prescript{t} {}(M^{-1})^{(3)}(\nabla f) \bigl( M^{-1}\ \prescript{t} {} (x \ y \ z) \bigr) \,,
\end{gather*}
%
where $(M^{-1})^{(j)}$ denotes the $j$-th column of the matrix $M^{-1}$. Hence
%
\[
  \nabla (M \cdot f) = \prescript{t} {} M^{-1} \cdot (\nabla f) \bigl(M^{-1}\ \prescript{t} {} (x \ y \ z)\bigr) \,,
\]
%
so we have
%
\[
  \nabla (M \cdot f)(Q)=\nabla (M \cdot f)(M \cdot P)=
  \prescript{t} {} M^{-1} \cdot (\nabla f) (M^{-1}\ M \cdot P)=\prescript{t} {} M^{-1} \cdot (\nabla f)(P) \,.
\]
%
Finally, if we choose $M \in \SO_3(\C)$, we have
$\prescript{t} {} M^{-1}=M$. We deduce that
\Cref{eq:transformed} holds if and only if $Q \in \Eig{M\cdot f}$, so the statement is proved.
\end{proof}


\section{Conditions imposed by aligned eigenpoints}
\label{conditions}

Having a given point as eigenpoint determines linear conditions on the space of ternary cubic forms, 
which can hence be encoded into a matrix, called the \emph{matrix of conditions} (\Cref{definition:matrix_conditions}). 
After analyzing the possible ranks of matrices of conditions relative to three or four points (possibly aligned), we characterize in terms of three polynomial conditions (denoted $\delta_1$, $\bar{\delta}_1$ and~$\delta_2$) the situations with two triples of aligned eigenpoints.

\subsection{The matrix of conditions}

Imposing a cubic ternary form to have one or more aligned triples of eigenschemes implies conditions both on the points and on the cubics.
We begin to explore these conditions by introducing, for each point
$P \in \p^2$,
a $3 \times 10$ matrix encoding the condition that $P$ is an eigenpoint of a ternary cubic.

\begin{definition}
\label{definition:matrix_conditions}
Consider $\p^9 = \p(\C[x,y,z]_3)$, the space of all ternary cubics.
Throughout this paper, we consider the standard monomial basis for $\C[x,y,z]_3$ and we set~$\mathcal{B}$ to be the following vector
%
\begin{equation}
\label{eq:vector_basis}
  \mathcal{B} := (x^3, x^2 y, x y^2, y^3, x^2 z, x y z, y^2 z, x z^2, y z^2, z^3)
  \in \bigl( \C[x,y,z]_3 \bigr)^{\oplus 10} \,.
\end{equation}
For $f \in \C[x,y,z]_3$, denote by~$[f]$ the corresponding point in~$\p^9$; we denote by~$w_f$ the (column) vector of coordinates of~$f$ with respect to the basis above; then $w_f$ is also a vector of projective coordinates of~$[f]$.
For a point $P \in \p^2$ with coordinates $(A: B: C)$, the condition on elements~$[f]$ of~$\p^9$ that $P$ is an eigenpoint of the ternary cubic form~$f$ can be expressed in the form
%
\[
  \Phi(P) \cdot w_f = 0 \,,
\]
%
where $\Phi(P)$ is a $3 \times 10$ matrix with entries depending on $A, B, C$.
The matrix $\Phi(P)$ is called the \emph{matrix of conditions} imposed by~$P$.
We denote by~$\phi_1(P)$, $\phi_2(P)$, and~$\phi_3(P)$ the rows of~$\Phi(P)$.
Written as vectors, they are
%
\begin{equation}
\label{equation:matrix_conditions_rows}
  \begin{aligned}
    \phi_1(P) &=
    \scalemath{0.9}{
    (-3A^2B, A(A^2 - 2B^2), B(2A^2 - B^2), 3AB^2,
     -2ABC, C(A^2 - B^2), 2 ABC,
     -B C^2, A C^2, 0)} \,, \\
    \phi_2(P) &= 
    \scalemath{0.9}{
    (-3A^2 C,
     -2ABC,
     -CB^2,
     0,
     A(A^2-2C^2),
     B(A^2 - C^2),
     AB^2,
     C(2A^2-C^2),
     2ABC,
     3AC^2)} \,,\\
    \phi_3(P) &=
    \scalemath{0.9}{
    (0,
     -A^2C,
     -2ABC,
     -3CB^2,
     A^2 B,
     A(B^2 - C^2),
     B(B^2-2C^2),
     2ABC,
     C(2B^2-C^2),
     3BC^2)} \,.
  \end{aligned}
\end{equation}
%
More generally, if $P_1, \dotsc, P_n$ are points in the plane, we denote by~$\Phi(P_1, \dotsc, P_n)$ the matrix whose rows are
%
\[
  \phi_1(P_1), \phi_2(P_1), \phi_3(P_1),
  \dotsc, 
  \phi_1(P_n), \phi_2(P_n), \phi_3(P_n),
\]
%
and we call it the
\emph{matrix of conditions} imposed by $P_1, \dotsc, P_n$.
\end{definition}

\begin{definition}
Given a matrix $M$ of type $m \times 10$, we denote by~$\Lambda(M)$ the projective linear system of cubics $[a_0 x^3 + \dotsb + a_9 z^3]$ such that $M \, \prescript{t}{}{\left( a_0 \,  \cdots \,  a_9 \right)} = 0$.
\end{definition}

We introduce a new piece of notation that makes the description of~$\Phi(P)$ more compact: let $\de_x \mathcal{B}$ denote the vector whose entries are the derivatives of those of~$\mathcal{B}$, and similarly for~$\de_y \mathcal{B}$ and~$\de_z \mathcal{B}$.
In this way, if $P=(A: B: C)$, we get:
%
\begin{equation}
\label{equation:vector_conditions}
  \begin{aligned}
    \phi_1(P) &= A\cdot \de_y \mathcal{B}(P) - B\cdot \de_x \mathcal{B}(P) \,, \\
    \phi_2(P) &= A\cdot \de_z \mathcal{B}(P) - C\cdot \de_x \mathcal{B}(P) \,, \\
    \phi_3(P) &= B\cdot \de_z \mathcal{B}(P) - C\cdot \de_y \mathcal{B}(P) \,.
  \end{aligned}
\end{equation}
%
\begin{rmk}
By analyzing the entries of \Cref{equation:matrix_conditions_rows}, it is not difficult to check that the rank of~$\Phi(P)$ is never $\leq 1$. \Cref{equation:vector_conditions} gives the relation
%
\begin{equation}
\label{eq:syzygy}
  C \, \phi_1(P) - B \, \phi_2(P) + A \, \phi_3(P) = 0 \,.
\end{equation}
%
Therefore, at most two of the vectors~$\phi_1(P)$, $\phi_2(P)$, and~$\phi_3(P)$ are linearly independent and the matrix~$\Phi(P)$ has rank~$2$.
As a consequence, if $P_1, \dots, P_n$ are $n$ points of the plane, we have:
%
\[
  \rk \,\Phi(P_1, \dots, P_n) \leq \min \left\{2n, 10 \right\} \,.
\]
%
\end{rmk}

\subsection{Possible ranks of the matrix of conditions}

In what follows, we want to study the possible values of the rank of the matrix
$\Phi(P_1, \dots, P_n)$ for several configurations of points $P_1, \dots, P_n$
(and several values of~$n$).
In particular, we study the ideal~$J_k$ of order $k$ minors of the
involved matrices of conditions and we deduce some bounds about the rank from the primary
decomposition of~$J_k$. 
Most of these computations are done with the aid of a computer algebra system.
Nevertheless, in many cases, the result cannot be reached just by brute force, 
but it is necessary to preprocess the ideal~$J_k$. 
In particular, it is often convenient to first saturate the ideal~$J_k$ with respect to
the condition that some points are distinct or that three of them are not aligned 
(when this is the case). 
Another important simplification that we adopt sometimes, makes use
of the action of~$\SO_3(\C)$: thanks to it we can assume that one of
the point is either $(1: 0: 0)$ or $(1: \iii: 0)$; see \Cref{two_orbits}.

We start with the following lemma, which is extremely useful
to speed up the computations.

\begin{lemma}
\label{lemma:minors}
Let $l_1 < \cdots <l_n$ be $n$ indices (where $3 \leq n \leq 10$) and let $P = (A: B: C)$ be a point of the plane.
Construct three $1 \times n$ matrices $w_1$, $w_2$, $w_3$ by extracting the entries of position $l_1, \dotsc, l_n$ from~$\phi_1(P)$, $\phi_2(P)$, and~$\phi_3(P)$, respectively. If $L$ is a $(n-2) \times n$ matrix, set:
%
\[
  L_1 := \left( \begin{array}{c} w_1 \\ w_2 \\ L \end{array} \right), \quad
  L_2 := \left( \begin{array}{c} w_1 \\ w_3 \\ L \end{array} \right), \quad
  L_3 := \left( \begin{array}{c} w_2 \\ w_3 \\ L \end{array} \right)
\]
%
Then
%
\[
  B \det(L_1) = A \det(L_2), \quad
  C \det(L_1) = A \det(L_3), \quad
  C \det(L_2) = B \det(L_3)
\]
%
hence $(A: B: C) = \bigl( \det(L_1): \det(L_2): \det(L_3) \bigr)$.
\end{lemma}
\begin{proof}
The statement easily follows from the equality $C w_1 - B w_2 + A w_3 = 0$, which is a direct consequence of \Cref{eq:syzygy}.
\end{proof}

% \Cref{lemma:minors} justifies the following choice.

% \begin{definition}
% \label{definition:reduced_matrix_conditions}
% For $n$ points $P_1, \dotsc, P_n$ in the plane, the \emph{cut matrix of conditions} of $P_1, \dotsc, P_n$ is the submatrix of~$\Phi(P_1, \dotsc, P_n)$ whose rows are $\phi_1(P_1), \phi_2(P_1), \dotsc, \phi_1(P_n), \phi_2(P_n)$.
% \end{definition}

Next, we point out a property of the lines that are tangent to the absolute conic~$\iso$.
First of all, we fix some notation that is used throughout the paper.

\begin{definition}
\label{definition:sigma}
We define a bihomogeneous polynomial of bidegree~$(2,2)$ on $\p^2 \times \p^2$: for $P_1 = (A_1: B_1: C_1)$ and $P_2 = (A_2: B_2: C_2)$, we set
%
\begin{equation}
\label{formula:sigma}
\begin{aligned}
  \sigma(P_1, P_2) &= \scl{P_1}{P_1} \scl{P_2}{P_2} - \scl{P_1}{P_2}^2 \\
   &= s_{11}s_{22}-s_{12}^2
\end{aligned}
\end{equation}
%
\end{definition}

\begin{rmk}
\label{rmk:sigma_discr}
The form~$\sigma$ is the discriminant of the intersection between the line~$P_1 \vee P_2$ and~$\iso$.
Since the general point of the line~$P_1 \vee P_2$ can be expressed in the form
$t P_1 + sP_2$ where $(t:s) \in \p^1$, the intersection between the line~$P_1 \vee P_2$
and~$\iso$ is determined by the pairs $(\overline{t}: \overline{s}) \in \p^1$ which are solutions of
%
\[
  \scl{t P_1 + sP_2}{t P_1 + sP_2} =0.
\]
%
The bilinearity and symmetry of the form $\scl{}{}$ gives the quadratic polynomial
%
\begin{equation}
\label{eq:intersection_isotropic}
  \scl{t P_1 + sP_2}{t P_1 + sP_2} =
  \scl{P_1}{P_1} t^2 + 2\scl{P_1}{P_2} ts + \scl{P_2}{P_2}s^2,
\end{equation}
%
whose discriminant is $4 ( \scl{P_1}{P_2}^2 - \scl{P_1}{P_1} \scl{P_2}{P_2})$.
Hence the condition $\sigma(P_1, P_2)=0$ implies that the polynomial in \Cref{eq:intersection_isotropic}
has a double root, and this proves the statement.
\end{rmk}

\begin{prop}
\label{proposition:sigma_tangency}
Let $P_1$, $P_2$ be two distinct points in the plane and let $r$ be the line joining them.
Then the following are equivalent:
%
\begin{enumerate}
  \item $\sigma(P_1, P_2) = 0$;
  \item $\sigma(Q_1, Q_2) = 0$ for all pairs of distinct points $Q_1, Q_2 \in r$;
  \item the line~$r$ is tangent to~$\iso$ at some point;
  \item there exists a point $T \in r$ such that $T \in \iso$ and $\scl{T}{Q} = 0$ for all $Q \in r$, $Q \neq T$; in this case we say that $T$ is \emph{orthogonal} to~$r$.
\end{enumerate}
%
\end{prop}
\begin{proof}
By \Cref{rmk:sigma_discr}, we have that $\sigma(P_1, P_2) = 0$ if and only if $r$ is tangent to~$\iso$ in a point~$T$; this shows that the first three items are equivalent.

Now, if $\sigma(P_1, P_2) = 0$, then $r$ is tangent to~$\iso$ at a point~$T$, hence $\scl{T}{T} = 0$.
Moreover, we also have that $\sigma(T, Q) = 0$ for all $Q \in r$ with $Q \neq T$.
Hence, $T$ is orthogonal to~$r$.
The converse is immediate.
\end{proof}

\begin{prop}
\label{proposition:three_aligned_ranks}
Let $P_1, P_2, P_3$ be three distinct aligned points of the plane and let
$r$ be the line passing through them. Then:
%
\begin{itemize}
  \item $5 \leq \rk \,\Phi(P_1, P_2, P_3) \leq 6$;
  \item $\rk \,\Phi(P_1, P_2, P_3) = 5$ if and only if $r$ is tangent to~$\iso$ in one of the three points~$P_1, P_2$, or~$P_3$.
\end{itemize}
%
\end{prop}
\begin{proof} We denote the coordinates of the points as follows:
%
\[
  P_1 = (A_1: B_1: C_1) \,, \quad
  P_2 = (A_2: B_2: C_2) \,, \quad
  P_3 = u_1 \, P_1 + u_2 \, P_2 \,.
\]
%
The characterization of the triples $P_1, P_2, P_3$ of distinct points such that $\rk \,\Phi(P_1, P_2, P_3) < 6$ follows from the computation of a suitable saturation of the ideal generated by the $17640$ order six minors of~$\Phi(P_1, P_2, P_3)$.
We claim that we can greatly simplify the computations.

Indeed, consider the matrix~$M$ constructed with the following six rows:
%
\[
  \phi_{i_1}(P_1), \ \phi_{i_2}(P_1), \
  \phi_{j_1}(P_2),\ \phi_{j_2}(P_2), \
  \phi_{k_1}(P_3), \ \phi_{k_2}(P_3)
\]
%
where $i_1, i_2, j_1, j_2, k_1, k_2 \in \{1, 2, 3\}$ with $i_1 \neq i_2$, $j_1 \neq j_2$, and $k_1 \neq k_2$.
First, assume that $i_1=j_1=k_1=1$ and $i_2=j_2=k_2=2$.
Fix six columns $1\leq l_1 < \cdots < l_6 \leq 10$ of~$M$ and let $N$ be the order~$6$ matrix obtained from~$M$ with these columns.
Since two rows of~$N$ are obtained from
entries of~$\phi_1(P_1)$ and of~$\phi_2(P_1)$,
\Cref{lemma:minors} gives that $A_1$ divides~$\det(N)$. For a similar
reason, also $A_2$ and $u_1A_1+u_2A_2$ divides~$\det(N)$, so
$\det(N) = A_1A_2(u_1A_1+u_2A_2)\cdot D$ for a suitable~$D$.
Again by \Cref{lemma:minors}, we also see that if we take different values of $i_1, i_2, j_1, j_2, k_1, k_2$ in the definition of~$M$ above, then
the corresponding~$\det(N)$ is of the form $X_1X_2X_3\cdot D$ where
$X_1$ is a coordinate of~$P_1$, $X_2$ is a coordinate of~$P_2$, and $X_3$ is
a coordinate of~$P_3$ (and $D$ is the same as above).
Since each of the three points has at least one non-zero coordinate,
if the order six minor constructed with the columns $l_1, \dots, l_6$ is zero, 
then necessarily $D$ must be zero. 
In this way, to force all order~$6$ minors of~$\Phi(P_1, P_2, P_3)$ to be zero,
it is enough to compute the ideal of the order six minors of the matrix~$M$ above
and divide each minor by $A_1A_2(u_1A_1+u_2A_2)$. 
By this procedure, after a saturation w.r.t.\ the condition that $P_1, P_2, P_3$ are pairwise distinct,
we get an ideal that has a very simple primary decomposition, 
given by the following three ideals:
%
\[
  \bigl( \scl{P_i}{P_1}, \scl{P_i}{P_2}, \scl{P_i}{P_3} \bigr) \quad
  \mbox{for } i = 1, 2, 3 \,.
\]
%
As a consequence, the line~$r$ is tangent to~$\iso$
(in $P_1$ or $P_2$ or $P_3$).
It is not difficult to see, via a symbolic computation,
that at least one order~$4$ minor of~$\Phi(P_1, P_2, P_3)$ is non-zero.
\end{proof}

\begin{prop}
\label{manca il riferimento su ancillary  non e': condition_rank_aligned}
%%%  conto si trova su file prop47_5ago.sage
Let $P_1, P_2, P_4$ be three distinct points of the plane. Then:
%
\begin{itemize}
  \item $5 \leq \rk \,\Phi(P_1, P_2, P_4) \leq 6$;
  \item if
  $\rk \, \Phi(P_1, P_2, P_4) = 5$, then $P_1, P_2, P_4$
  are aligned and the line joining them is tangent to~$\iso$
  in one of the three points.
\end{itemize}
%
\end{prop}
\begin{proof}
By \Cref{two_orbits}, we can split the proof into two parts, considering the case in
which $P_4 = (1: 0: 0)$ and the case in which $P_4 = (1: \iii: 0)$.
In both cases, the computation of the ideal of order~$5$ minors of the matrix~$\Phi(P_1, P_2, P_4)$ and its saturation w.r.t.\ the distinct point condition gives the whole ring,
so $\Phi(P_1, P_2, P_4)$ cannot have rank smaller than~$5$.
Moreover, the computation of the ideal of the order~$6$ minors of~$\Phi(P_1, P_2, P_4)$ and its
saturation w.r.t.\ the distinct point condition gives that
$P_1, P_2, P_4$ must be aligned and, according to \Cref{proposition:three_aligned_ranks},
that line is tangent to~$\iso$ in one of the three given points.
\end{proof}

\begin{prop}
\label{prop:condition3+1}
Let $P_1, P_2, P_3, P_4$ be four distinct points of the plane such that
$P_1, P_2, P_3$ are aligned and let $r$ be the line joining them.
If $\rk \,\Phi(P_1, P_2, P_3, P_4) \leq 7$ then $r$ is tangent to~$\iso$ in one of the three points~$P_1$,  $P_2$, and~$P_3$.
\end{prop}
\begin{proof}
Again, we distinguish two cases: $P_1 = (1: 0: 0)$ and
$P_1 = (1: \iii: 0)$. In both cases, the ideal of the order~$8$
minors of $\Phi(P_1, P_2, P_3, P_4)$ can be computed and saturated
w.r.t.\ the distinct point condition and that
$P_1, P_2, P_4$ are not aligned.
The direct inspection of the ideal obtained from these procedures gives the thesis.
\end{proof}

\begin{rmk}
As we clarified in \Cref{thm:general_no_d_collinear}, four aligned eigenpoints are never contained in a $0$-dimensional, eigenscheme.
In \Cref{4ptiSuRetta2}, we will show that if an eigenscheme contains four distinct aligned eigenpoints, then it contains the whole line joining them.
\end{rmk}

\begin{prop}
\label{proposition:four_aligned}
Let $Q_1, \dotsc, Q_4$ be four distinct aligned points of the plane and
let $r$ be the line through them. Then:
%
\begin{itemize}
  \item $6 \leq \rk \,\Phi(Q_1, \dotsc, Q_4) \leq 7$;
  \item $\rk \,\Phi(Q_1, \dotsc, Q_4) = 6$ if and only if $r$ is tangent to~$\iso$.
\end{itemize}
%
The same statements hold when we consider $k \geq 4$ distinct aligned points $Q_1, \dotsc, Q_k$.
\end{prop}
\begin{proof}
A symbolic computation shows that all the maximal minors of~$\Phi(Q_1, \dotsc, Q_4)$ are zero, so we get 
$\rk \, \Phi(Q_1, \dotsc, Q_4) \leq 7$.
Indeed, in view of \Cref{lemma:minors}, we can just check the order~$8$ minors of the cut matrix of conditions of $Q_1, \dotsc, Q_4$;
it turns out that they are all zero.

On the other hand, the rank of $\Phi(Q_1, \dotsc, Q_4)$ cannot be $5$:
if this were the case, for any triple~$T$ extracted from the four points, the matrix of conditions would have rank~$5$, hence by \Cref{proposition:three_aligned_ranks} the line~$r$ would be tangent to~$\iso$ in an element of~$T$; this is a contradiction.
So, the first item of the statement is proven.

We proceed to prove the second item.
We take the cut matrix of conditions of~$Q_1, Q_2, Q_3$, to which we add the row~$\phi_1(Q_4)$.
We call this matrix~$M_1$.
Similarly, we construct the matrices $M_2$ and~$M_3$.
We consider the coordinates of the points as variables and we compute the ideal~$J_1$,
the radical of the ideal of the maximal minors of~$M_1$, namely, the order~$7$ minors;
similarly, we obtain $J_2$ and~$J_3$.
The saturation of the ideal sum $J_1 + J_2 + J_3$ by the distinct point condition is a principal ideal generated by~$\sigma(r)$.
Since the four points $Q_1, \dotsc, Q_4$ play symmetric roles, the statement is proven.
\end{proof}

\subsection{The conditions \texorpdfstring{$\delta_1$}{delta1}, \texorpdfstring{$\bar{\delta}_1$}{deltabar1}, and \texorpdfstring{$\delta_2$}{delta2}}

We now define three quantities depending on a triple or on a $5$-tuple of points in the plane.
These quantities are crucial to describe what happens when we have aligned eigenpoints.
In the following, we denote by~$P \vee Q$ the line through two distinct points~$P$ and~$Q$.

\begin{definition}
\label{definition:delta1}
We define a homogeneous polynomial of degree~$(2,1,1)$ on $\p^2 \times \p^2 \times \p^2$:
if $P_i = (A_i: B_i: C_i)$ for $i \in \{1, 2, 4\}$, we set
%
\[
  \delta_1(P_1, P_2, P_4) :=
  \scl{P_1}{P_1} \scl{P_2}{P_4} - \scl{P_1}{P_2}\scl{P_1}{P_4} =
  \scl{P_1\times P_2}{P_1 \times P_4} \,,
\]
%
where $\times$ denotes the cross product, i.e.,
%
\[
  P_1 \times P_2 = 
  (B_1 C_2 - C_1 B_2, \, C_1 A_2 - A_1 C_2, \, A_1 B_2 - B_1 A_2) \,.
\]
%
\end{definition}

\begin{rmk}
\label{rmk:delta1_meaning}
Geometrically, the condition $\delta_1(P_1, P_2, P_4) = 0$ corresponds to the orthogonality of the vector planes corresponding to the lines~$P_1 \vee P_2$ and~$P_1 \vee P_4$, if the three points are not collinear, while
$\delta_1(P_1, P_2, P_4) = 0$ implies that $\sigma (P_1,P_2)=0$ if they are collinear.
\end{rmk}

\begin{definition}
\label{definition:delta1b}
Let $P_1$, $P_2$ and~$P_3$ be distinct aligned points in the plane.
We define the polynomial
%
\[
  \overline{\delta}_1(P_1, P_2, P_3) :=
  \scl{P_1}{P_1} \scl{P_2}{P_3} + \scl{P_1}{P_2}\scl{P_1}{P_3} \,.
\]
%
\end{definition}

\begin{definition}
\label{definition:Vconf}
Let $P_1, P_2, P_3, P_4, P_5$ be five distinct points of the plane~$\p^2$
such that $P_1, P_2, P_3$ and $P_1, P_4, P_5$ are aligned.
We call such a configuration a \emph{$V$-configuration}.
\end{definition}

\begin{definition}
Let $P_1, \dots, P_5$ be a $V$-configuration.
We define the polynomial
%
\[
  \delta_2(P_1, P_2, P_3, P_4, P_5) :=
  \scl{P_1}{P_2} \scl{P_1}{P_3} \scl{P_4}{P_5} -
  \scl{P_1}{P_4} \scl{P_1}{P_5} \scl{P_2}{P_3} \,.
\]
%
\end{definition}

Recall that $\scl{P_i}{P_j}$ is also denoted by~$s_{ij}$.
Then the expressions~$\delta_1$ and~$\delta_2$ become
%
\[
  \delta_1(P_1, P_2, P_4) = s_{11} s_{24}-s_{12}s_{14} \,, \quad
  \delta_2(P_1, P_2, P_3, P_4, P_5) = s_{12}s_{13}s_{45}-s_{14}s_{15} s_{23} \,.
\]
%
\begin{rmk}
\label{rmk:characteristics_d1_d2}
\textbf{@@Verifiche contenute nel file} \verb+formulae.sage+ \\
It holds
%
\begin{align}
\label{rmk_delta_case1}
  \delta_1(P_1, P_2, P_4) = 0 \mbox{ iff } &\scl{P_4}{s_{11}P_2-s_{12}P_1} = 0\\
  \mbox{iff } &\scl{P_2}{s_{11}P_4-s_{14}P_1} = 0 \,. \nonumber
\end{align}
%
\begin{align}
\label{rmk_delta_case2}
  \overline{\delta}_1(P_1, P_2, P_3) = 0 \mbox{ iff } &
  P_1 \mbox{ is on~$\iso$ and } P_1 \vee P_2 \vee P_3 \mbox{ is tangent to~$\iso$ in $P_1$, or} \\
  & P_3 = (s_{12}^2+s_{11}s_{22}) \, P_1 - 2s_{11}s_{12} \, P_2 \mbox{ or} \nonumber \\
  & P_2 = (s_{13}^2+s_{11}s_{33}) \, P_1 - 2s_{11}s_{13} \, P_3 \nonumber
\end{align}
%
\end{rmk}

Suppose the point~$P_1$ is fixed. In general, in order to have
$\delta_2(P_1, P_2, P_3, P_4, P_5) = 0$, we can fix three of the other points and determine the fourth one, 
by observing that $\delta_2=0$ gives a linear equation in its coordinates.
There are, however, some exceptions. 
From the definition of~$\delta_2$, we see for instance that
if $P_2$ and~$P_4$ satisfy $\{s_{12}=0, s_{14}=0\}$ or $P_2$ and~$P_5$ satisfy $\{s_{12}=0, s_{15}=0\}$ or $P_2$ and~$P_3$ satisfy
$\{s_{12}=0, s_{23}=0\}$, etc., then $\delta_2$ is zero,
despite of the value of the other points.
To be more precise, we have:
%
\begin{prop}
\label{prop:definitionP3}
Let $P_1, \dots, P_5$ be five points in a $V$-configuration. Then it holds:
$\delta_2(P_1, P_2, P_3, P_4, P_5) = 0$ if and only if (up to a permutation
of some of the points $P_2, \dots, P_4$) one of the following conditions
is satisfied:
%
\begin{enumerate}
  \item $s_{12} = 0$ and $s_{14} = 0$;
  \label{defP3_1}
  \item $s_{12} = 0$ and $s_{22} = 0$;
  \label{defP3_2}
  \item $\sigma(P_1, P_2) = 0$ and $\sigma(P_1, P_4) = 0$;
  \label{defP3_3}
  \item $P_3 = (s_{14}s_{15}s_{22}-s_{12}^2s_{45})P_1  +s_{12}(s_{11}s_{45}-s_{14}s_{15})P_2$.
  \label{defP3_4}
\end{enumerate}
%
\end{prop}

\begin{proof}
%% file: whenP3isNotDefined.sage
Condition~(\ref{defP3_4}), when defined (i.e., when the coefficients of~$P_1$ and~$P_2$ are not zero), easily comes from the definition of~$\delta_2$.
From it we see that $P_3$ is not defined by that formula if and only if
$s_{14}s_{15}s_{22}-s_{12}^2s_{45}=0$ and $s_{11}s_{12}s_{45}-s_{12}s_{14}s_{15}=0$.
Hence we study the ideal generated by these two polynomials, plus the
polynomial $s_{12}s_{13}s_{45}-s_{14}s_{15} s_{23}$ which defines
$\delta_2$ in terms of the scalar product of the points. The corresponding
ideal decomposes into several ideals which are, up to a permutation of
the indices: $(s_{12}, s_{14})$, $(s_{12}, s_{22})$ and:
%
\[
  J = (s_{13}s_{22} - s_{12}s_{23}, s_{14}s_{15} - s_{11}s_{45}, s_{12}s_{13} -
  s_{11}s_{23}, s_{12}^2 - s_{11}s_{22}) \,.
\]
%
Then we give generic coordinates to the five
points and substitute them into the ideal $J$. It is possible to see that
$J$ and the ideal $\bigl(\sigma(P_1, P_2), \sigma(P_1, P_4)\bigr)$ are equal (up to
saturations w.r.t.\ $v_2$ and the condition that $P_1, P_2, P_4$ are not
aligned).
\end{proof}
%
\begin{rmk}
In the first case of \Cref{prop:definitionP3}, the corresponding cubic
is studied in \Cref{further_alignments}, configuration $(C_5)$.
In the second case, $\sigma(P_1, P_2) = 0$ and $P_2\in \iso$, so
we have that the line~$P_1 \vee P_2 \vee P_3$ is tangent to~$\iso$ in~$P_2$
and this case was considered in~\Cref{proposition:three_aligned_ranks}
(the matrix $\Phi(P_1, P_2, P_3)$ has rank~$5$, so
$\Phi(P_1, \dots, P_5)$ has rank $9$ (or smaller) regardless of
$P_4$ and~$P_5$).
The third case gives either that all the
points of the two lines of the $V$-configurations are eigenpoints (in case
$P_2 \not\in \iso$ or $P_4 \not\in \iso$) or the matrix $\Phi(P_1, \dots, P_5)$
has rank $8$ (if $P_2, P_4 \in \iso$); see \Cref{rank_8} and \Cref{positive_dim}, respectively.
The last case is the generic one and expresses $P_3$ in terms of the remaining four points.
\end{rmk}
%
\begin{prop}
\label{prop:d1d2}
Let $P_1, \dots, P_5$ be a $V$-configuration. Then
%
\[
  \rk \,\Phi(P_1, \dots, P_5) \leq 9
  \quad \mbox{if and only if} \quad
  \delta_1(P_1, P_2, P_4) \cdot \delta_2(P_1, \dots, P_5) = 0 \,.
\]
%
\end{prop}
\begin{proof}
We denote the coordinate of the points as follows:
%
\[
  P_1 = (A_1: B_1: C_1) \,, \quad
  P_2 = (A_2: B_2: C_2) \,, \quad
  P_4 = (A_4: B_4: C_4) \,,
\]
%
then $P_3 = u_1 \, P_1 + u_2 \, P_2$ and $P_5 = v_1 \, P_1 + v_2 \, P_4$ for some $u_1, u_2, v_1, v_2$.
We proceed as in
the proof of \Cref{proposition:three_aligned_ranks}: we find, via symbolic computation, that the determinant of
the cut matrix of conditions of $P_1, \dotsc, P_5$ is
%
\begin{gather}
\label{delta1delta2}
  A_1A_2A_4(u_1A_1+u_2A_2)(v_1A_1+v_2A_4) \cdot u_1^2u_2^2v_1^2v_2^2 \cdot D^5 \cdot
  \delta_1(P_1,P_2,P_4) \cdot \delta_2(P_1,\dots,P_5) \,,
\end{gather}
%
where $D$ is the determinant of the matrix whose rows are $P_1, P_2, P_4$
and is non-zero, as well as are non-zero $u_1, u_2, v_1, v_2$ (since we
assume that $P_1, P_2, P_4$ are not aligned and the points are distinct).
As a consequence of \Cref{lemma:minors}, we have that when we compute all the order~$10$ minors of the matrix of conditions, we obtain all polynomials of the form
%
\begin{gather*}
  X_1X_2X_4 X_3 X_5 \cdot u_1^2u_2^2v_1^2v_2^2 \cdot D^5 \cdot
  \delta_1(P_1,P_2,P_4) \cdot \delta_2(P_1,\dots,P_5) \,,
\end{gather*}
%
where $X_i$ varies among all components of~$P_i$ for $i \in \{1, \dotsc, 5\}$.
Since each of $P_1, \dotsc, P_5$ has at least one non-zero coordinate, we have that
$\rk \, \Phi(P_1, \dots, P_5) \leq 9$ if and only if
%
\[
  \delta_1(P_1, P_2, P_4) \cdot \delta_2(P_1, \dots, P_5) = 0 \,. \qedhere
\]
%
\end{proof}

%% for the proof of the avobe lemma, see the first part of the file
%% rank_8_twoTangIso.sage"
%%
\begin{lemma}
\label{lemma:special_case_rank_8}
Let $P_1, \dots, P_5$ be a $V$-configuration of points and assume that
%
\[
  \scl{P_1}{P_2}=0 \,, \quad
  \scl{P_2}{P_2}=0 \,, \quad
  \scl{P_1}{P_4}=0 \,, \quad
  \scl{P_4}{P_4}=0 \,.
\]
%
Then the matrix $\Phi(P_1, \dots, P_5)$ has rank~$8$.
\end{lemma}
\begin{proof}
%% first part of the file rank_8_twoTangIso.sage
By \Cref{proposition:sigma_tangency},
the lines~$P_1 \vee P_2$ and~$P_1 \vee P_4$ are tangent to~$\iso$ in~$P_2$ and~$P_4$, respectively. 
The point~$P_1$ cannot be on~$\iso$, hence, using the
action of~$\SO_3(\C)$, we can assume $P_1 = (1: 0: 0)$.
Since every element of~$\SO_3(\C)$ fixes the whole~$\iso$ (and sends a tangent line to it into another tangent line to it), 
when we transform the point~$P_1$ into $(1: 0: 0)$, 
we transform the points~$P_2$ and~$P_4$ into, respectively,
the points $(0: \iii: 1)$ and $(0: -\iii: 1)$ (which are the common points to~$\iso$ and the tangent lines through~$P_1$).
Therefore, it is enough to study the
specific configuration of the points:
%
\begin{gather*}
  P_1 = (1: 0: 0) \,, \quad P_2=(0: \iii: 1) \,, \quad P_3=(u_1, \iii u_2, u_2) \,, \\
  P_4 = (0: -\iii: 1) \,, \quad P_5 = (v_1, -\iii v_2, v_2) \,,
\end{gather*}
%
where $(u_1: u_2), (v_1: v_2) \in \p^1$.
In the $15\times 10$ matrix $\Phi(P_1, \dots, P_5)$ we can erase the
rows: $\phi_2(P_1)$ (which is a zero row), the row~$\phi_1(P_2)$
(since $\phi_1(P_2)=\iii\phi_2(P_2)$) and, for similar reasons, the
rows: $\phi_1(P_3)$, $\phi_1(P_4)$ and~$\phi_1(P_5)$.
Moreover, since the rank of the matrix $\Phi(P_1, P_2, P_3)$ is $5$,
we can also erase the row~$\phi_3(P_3)$ and, for the same reason, the
row~$\phi_3(P_5)$. The remaining matrix~$M$ is a $8 \times 10$ matrix.
It is not possible that all the order~$8$ minors
of~$M$ are zero: the ideal they generate, after saturations with respect to the distinct point condition, is the whole ring.
\end{proof}

\begin{theorem}
\label{theorem:rank_V}
Let $P_1, \dots, P_5$ be a $V$-configuration of
points. Then we have:
%
\begin{enumerate}
  \item $8 \leq \rk \,\Phi(P_1, \dots, P_5) \leq 10$\,;
  \item $\rk \,\Phi(P_1, \dots, P_5) \leq 9$ if and only if
  $\delta_1(P_1, P_2, P_4) \cdot \delta_2(P_1, \dots, P_5) =0$\,;
  \item $\rk \,\Phi(P_1, \dots, P_5) = 8$ if and only if, one of
  the following two conditions is satisfied:
  %
  \begin{itemize}
    \item $\delta_1(P_1, P_2, P_4) = 0$, \
    $\overline{\delta}_1(P_1, P_2, P_3) = 0$, \ 
    $\overline{\delta}_1(P_1, P_4, P_5) = 0$\,;
    \item the line~$P_1 \vee P_2$ is tangent to~$\iso$ in~$P_2$ or~$P_3$
    and the line~$P_1 \vee P_4$ is tangent to~$\iso$ in $P_4$ or $P_5$; 
    moreover, in this case we have $\delta_1(P_1, P_2, P_4) \neq 0$.
  \end{itemize}
  %
  In particular, in both cases $\delta_2(P_1, P_2, P_3, P_4, P_5) = 0$ holds.
\end{enumerate}
%
\end{theorem}
\begin{proof}
%%% si basa sui file:
%%% rank_8_2_1_ii_0.sage e
%%% rank_8_1.sage
If the rank is $\leq 7$, from
\Cref{prop:condition3+1} applied to $P_1, P_2, P_3, P_4$ and $P_1, P_4, P_5, P_2$,
the lines~$P_1 \vee P_2$ and~$P_1 \vee P_4$ are tangent to~$\iso$ (the first in $P_2$ or $P_3$ and the second in $P_4$ or $P_5$).
Then we get a contradiction from \Cref{lemma:special_case_rank_8}.
This shows the first item.

The second item follows from \Cref{prop:d1d2}.

We are left to proving the third item. By \Cref{lemma:special_case_rank_8} and \Cref{prop:d1d2}, the two conditions imply that the rank is~$8$.

For the converse, by \Cref{invariance}, we distinguish two possibilities:
$P_1 = (1:\iii :0)$ and
$P_1 = (1: 0: 0)$.
Let $P_2 = (A_2: B_2: C_2)$, $P_4 = (A_4: B_4: C_4)$ and
$P_3 = u_1 \, P_1 + u_2 \, P_2$, $P_5 = v_1 \, P_1 + v_2 \, P_4$ be the other points.
If $P_1 = (1: \iii: 0)$, we transform the matrix $M = \Phi(P_1, \dots, P_5)$ by elementary row and column operations into $\mathbb{I}_2 \oplus M'$, where $\mathbb{I}_2$ is the order~$2$ identity matrix; thus, $\rk \, M \leq 9$ if and only if $\rk \, M' \leq 7$.
The computations require several other strategies, but at the end we are
able to see that in this case (i.e.\ the case in which $P_1$ is on~$\iso$), the matrix $M'$ cannot have rank less than~$7$, therefore if $P_1$ is on 
$\iso$, $\rk\, M$ cannot be less than~$9$.

If $P_1 = (1: 0: 0)$, then the ideal of the order $9$ minors of
$\Phi(P_1, \dots, P_5)$ (after several manipulations and saturations)
gives that there a few possibilities which can be summarized by the
following conditions: either $\delta_1(P_1, P_2, P_4) = 0$,
\ $\overline{\delta}_1(P_1, P_2, P_3) = 0$,
\ $\overline{\delta}_1(P_1, P_4, P_5) = 0$, or the line~$P_1 \vee P_2$ is tangent to~$\iso$ (in $P_2$ or $P_3$) and the line~$P_1 \vee P_4$ is tangent to~$\iso$ (in $P_4$ or $P_5$). We see that in the second case it is necessarily $\delta_1(P_1,P_2,P_4)
\neq 0$; indeed, in the case $P_2, P_4 \in \iso$, then $\scl{P_2 +P_4}{P_2+P_4}= \scl{P_2}{P_2} + \scl{P_4}{P_4}+2\scl{P_2}{P_4} =2\scl{P_2}{P_4} \neq 0$, as
the point~$P_2 + P_4$ is different from both~$P_2$ and~$P_4$, and the line~$P_2 \vee P_4$ has no other intersection points with~$\iso$. It follows that
$\delta_1 (P_1,P_2,P_4)=s_{11}s_{24} \neq 0$, as $P_1 \not\in \iso$. The cases $P_3 \in \iso$, respectively $P_5\in \iso$, are similar.

Therefore, if $\rk \, \Phi(P_1, \dots, P_5) = 8$, then one of the two conditions above must hold.
\end{proof}


\section{\texorpdfstring{$V$}{V}-configurations of rank~\texorpdfstring{$9$}{9}}
\label{rank_9}

In this section, we consider the following problem:
given five generic points in a $V$-configuration and imposing one
of the two conditions
$\delta_1(P_1, P_2, P_4) = 0$ or $\delta_2(P_1, \dots, P_5) = 0$, assuming the rank of $\Phi(P_1, \dots, P_5)$ is~$9$, what can we say about all the eigenpoints of the corresponding cubics?

Since the rank of $\Phi(P_1, \dots, P_5)$ is~$9$, there exists precisely one cubic curve~$C$ having $P_1, \dots, P_5$ among its eigenpoints. Another immediate consequence of the rank~$9$ condition is that the
matrix $\Phi(P_1, \dots, P_5)$ has $9$ linearly independent rows, so
we can extract from it a $9 \times 10$-submatrix
$\mathcal{H}$ whose order~$9$ minors are not all zero. 
If $\mathcal{H}_i$ is the order~$9$ minor of~$\mathcal{H}$ given by erasing the $i$-th column ($i=1, \dots, 10)$, 
then, from Cramer's rule, a polynomial defining~$C$ is given by
%
\[
  f(X) = \sum_{i=1}^{10}(-1)^i\det(\mathcal{H}_i)\cdot \mathcal{B}_i \,,
\]
%
where $\mathcal{B}$ is as in \Cref{eq:vector_basis}. In other words,
%
\[
  f(X) = \det \left( 
  \begin{array}{c} \mathcal{H} \\ \mathcal{B} \end{array}
  \right) \,.
\]
%
From the matrix $\mathcal{H}$ it is moreover possible to construct
the three minors of \Cref{eq:def_matrix}, indeed we have:
\begin{prop}
\label{proposition:geiser1}
If $g_1, g_2, g_3$ are the three minors of the matrix
%
\[
  \left(
  \begin{array}{ccc}
    x & y & z \\
    \de_x f(X) & \de_y f(X) & \de_z f(X)
  \end{array}
  \right)
\]
%
they can be computed by:
%
\[
  \det \left( 
  \begin{array}{c} \mathcal{H} \\ \phi_1(X) \end{array}
  \right),\quad
  \det \left( 
  \begin{array}{c} \mathcal{H} \\ \phi_2(X) \end{array}
  \right), \quad
  \det \left( 
  \begin{array}{c} \mathcal{H} \\ \phi_3(X) \end{array}
  \right)
\]
%
\end{prop}
\begin{proof} We have:
%
\begin{eqnarray*}
  g_1 & = & x \cdot \de_y f(X)- y \cdot \de_x f(X) \\
  & = &
  x \cdot \de_y \det \left(
  \begin{array}{c} \mathcal{H} \\ \mathcal{B} \end{array} 
  \right) - y \cdot
  \de_x \det \left(
  \begin{array}{c} \mathcal{H} \\ \mathcal{B} \end{array}
  \right) \\
  & = & \det \left(
  \begin{array}{c} \mathcal{H} \\ x \cdot \de_y \mathcal{B} - y \cdot \de_x \mathcal{B} \end{array}
  \right) \\
  & = & \left(
  \begin{array}{c} \mathcal{H} \\ \phi_1(X) \end{array}
  \right)
\end{eqnarray*}
%
and similarly for $g_2 = x \cdot \de_z f(X)- z \cdot \de_x f(X)$
and $g_3 = y \cdot \de_z f(X)- z \cdot \de_y f(X)$.
\end{proof}

\begin{es}
Consider the following five points in a $V$-configuration:
%
\begin{gather*}
  p_1 = (2: -1: 1) \,, \quad p_2 = (-1: 1: 3) \,, \quad p_4 = (3: 6: -1) \,, \\
  p_3 = p_1+p_2 \, \quad p_5 = p_1+p_4 \,.
\end{gather*}
%
They satisfy the condition $\delta_2(p_1, \dots, p_5) = 0$.
We consider the $9\times 10$ matrix $\mathcal{H}$ whose rows
(suitable rescaled) are $\phi_i(p_j)$ for $i = 1, 2$ and $j = 1, 2, 3, 4$, and $\phi_1(p_5)$
%
\[
  \mathcal{H} =
  \left(
  \begin{array}{rrrrrrrrrr}
    12 & 4 & -7 & 6 & 4 & 3 & -4 & 1 & 2 & 0 \\
    -12 & 4 & -1 & 0 & 4 & -3 & 2 & 7 & -4 & 6 \\
    -3 & 1 & 1 & -3 & 6 & 0 & -6 & -9 & -9 & 0 \\
    -9 & 6 & -3 & 0 & 17 & -8 & -1 & -21 & -6 & -27 \\
    0 & 1 & 0 & 0 & 0 & 4 & 0 & 0 & 16 & 0 \\
    -12 & 0 & 0 & 0 & -31 & 0 & 0 & -56 & 0 & 48 \\
    54 & 63 & 36 & -108 & -12 & -9 & 12 & 2 & -1 & 0\\
    27 & 36 & 36 & 0 & 21 & 48 & 108 & -17 & -36 & 9 \\
    336 & 328 & 119 & -588 & -56 & -33 & 56 & 7 & -4 & 0
  \end{array}
  \right)
\]
%
The cubic with $p_1, \dots, p_5$ as eigenpoints is given by
%
\[
  \det \left(
  \begin{array}{c}
    \mathcal{H}\\
    \mathcal{B}
  \end{array}
  \right)
\]
%
whose value is (after a suitable rescaling):
%
\[
  766x^3 - 1176x^2y + 417xy^2 - 128y^3 + 1488x^2z - 1722xyz
  + 597y^2z - 207xz^2 + 504yz^2 + 911z^3
\]
%
The polynomials~$g_1$, $g_2$, and~$g_3$ are:
%
\begin{align*}
  g_1 &= 392x^3 + 488x^2y - 656xy^2 + 139y^3 + 574x^2z + 594xyz - 574y^2z - 168xz^2 - 69yz^2 \\
  g_2 &= 496x^3 - 574x^2y + 199xy^2 - 904x^2z + 1120xyz - 139y^2z - 81xz^2 + 574yz^2 + 69z^3 \\
  g_3 &= 496x^2y - 574xy^2 + 199y^3 + 392x^2z - 416xyz + 464y^2z + 574xz^2 + 513yz^2 - 168z^3
\end{align*}
%
and can be computed either using their definition,
or~\Cref{proposition:geiser1}.
The common zeros of~$g_1$, $g_2$, $g_3$ are the points $p_1, \dots, p_5$
and the points $p_6 = (21: 34: -8)$ and $p_7 = (19: 35: -9)$.
\end{es}

\begin{rmk}
When the matrix $\Phi(P_1, \dots, P_5)$ has rank~$8$, as in \Cref{rank_8}, the family of
cubics with $P_1, \dots, P_5$ as eigenpoints is one dimensional, so the
above construction can be modified, by substituting one of the rows of~$\mathcal{H}$ with a random row of elements of~$\C$.
\end{rmk}

\begin{rmk}
\label{rmk:construction_five_d1}
Recalling \Cref{rmk:characteristics_d1_d2}, it is quite easy to construct five points $P_1, \dots, P_5$ that are in a $V$-configuration
and such that $\delta_1(P_1, P_2, P_4)= 0$: the points~$P_1$
and~$P_2$ can be taken in an arbitrary way, $P_4$ has to be chosen in such
a way that it satisfies \Cref{rmk_delta_case1}
and~$P_3$ and~$P_5$ have to be chosen on the lines~$P_1 \vee P_2$ and~$P_1 \vee P_4$,
respectively. In particular, the corresponding locus of cubic curves
has dimension~$7$.
The construction of a random example
of five points as above, gives a smooth cubic curve whose $7$ eigenpoints
do not have other collinearities (in addition to those of a
$V$-configuration).
\end{rmk}

\begin{rmk}
\label{rmk:construction_five_d2}
If we want a $V$-configuration that satisfies the condition
$\delta_2(P_1, \dots, P_5) = 0$, we choose $P_1$, $P_2$, $P_4$ arbitrarily;
we choose $P_3$ on the line~$P_1 \vee P_2$ and~$P_5$ on the line~$P_1 \vee P_4$ so that $P_3$ satisfies \Cref{prop:definitionP3}, \Cref{defP3_4}.
Also in this case, the locus of cubics is of dimension~$7$.
\end{rmk}

Cubics whose eigenpoints satisfy the condition $\delta_2(P_1, \dotsc, P_5) = 0$ and $\rk \, \Phi(P_1, \dotsc, P_5) = 9$ exhibit further peculiarities. It turns out that the $7$ eigenpoints of such cubics satisfy the further
condition that also $P_6$ and~$P_7$ are aligned with~$P_1$. To explain
why, we proceed as follows.

Consider again a matrix $\mathcal{H}$ of rank $9$ extracted from
the matrix $\Phi(P_1, P_2, P_3, P_4, P_5)$, where now $P_3$ is determined by Equation~\eqref{} \textbf{(SERVE QUESTA CONDIZIONE O E' SUFFICIENTE AVERE RANGO 9 E $\delta_2 = 0$?)}
and the three polynomials $g_1, g_2, g_3$ are constructed
in \Cref{proposition:geiser1}. The condition
$\rk \,(\mathcal{H}) = 9$ ensures that
$g_1, g_2$ and~$g_3$ cannot be zero and are polynomials of degree
$3$ in $x, y, z$.
The computation shows that three polynomials have a common factor,
hence are of the form $\Omega \cdot g_1$, $\Omega \cdot g_2$, $\Omega \cdot g_3$, where
$\Omega$ is a polynomial in the variables
$A_1, B_1, C_1, A_2, B_2, C_2, A_4, B_4, C_4, v_1, v_2$ but
does not contain the variables $x, y, z$.
Hence $g_1, g_2, g_3$ are again
degree three polynomials in $x, y, z$, with coefficients in the variables
$A_1, \dots, C_4, v_1, v_2$. Their common zeros are the eigenpoints of the
cubic~$C$ since by construction they are proportional to the three minors of the matrix
%
\[
  \left(
  \begin{array}{ccc}
    x & y & z \\
    \partial_x f & \partial_x f & \partial_x f
  \end{array}
  \right)
\]
%

\begin{prop}
\label{proposition:G_split}
Assume that $P_1, \dots, P_5 $ satisfy $\delta_2(P_1, \dotsc, P_5) = 0$ and $\rk \, \Phi(P_1, \dotsc, P_5) = 9$. Then the cubic curve $A_1 g_3 - B_1 g_2 + C_1 g_1=0$ contains the lines~$P_1 \vee P_2$ and~$P_1 \vee P_4$.
% The polynomial $C_1G_1-B_1G_2+A_1G_3$ splits into three factors $r_1$, $r_2$, $r_3$, each linear in $x, y, z$.
\end{prop}

\begin{definition}
%   Let $\Eig{f}$ be $0$-dimensional and reduced, and set
%%
%\[
%  M_1 :=y \partial _z f - z \partial _y f, \quad
%  M_2 := z \partial _x f - x \partial _z f, \quad
%  M_3 :=x \partial _y f - y \partial _x f.
%\]
%
The Geiser map associated with a zero dimensional eigenscheme $\Eig{f}$ is the rational map defined by
%
\[
  \gamma_{\Eig{f}} \colon \p ^2 \dasharrow \p^2, \quad
  \gamma_{\Eig{F}} (P) = \bigl( g_1(P):g_2(P):g_3(P) \bigr) \,,
\]
%
where $\p^2 = \p\bigl(I_{\Eig{f}}(3)^{\vee}\bigr)$, namely the projectivization of the dual of the space of degree~$3$ forms in the homogeneous ideal of~$\Eig{f}$.
\end{definition}

Geiser maps are a classical topic and several of their properties are understood.
As an example, the map~$\gamma_{\Eig{f}}$ is generically finite of degree~$2$, see for instance \cite[Section~8.7.2]{Dolgachev}.
%Moreover, the ramification locus of~$\widetilde{\gamma _{E(F)}$ is given by the \emph{Jacobian locus}~$\Sigma$ defined by the determinant of the Jacobian $\mathrm{Jac}(I_Z)$, that is the locus of singular points of the net (see
%    \cite[Book II, Chapter~IX, Theorem~25]{Cool}). When $Z$ is general, the Jacobian locus~$\Sigma$ is a curve of degree~$6$ which is singular at $Z$ as illustrated in \cite[Chapter~IX, Theorem~27]{Cool}. We define $B(Z)$ to be the branch locus of $\widetilde{\gamma_Z}$, that is the direct image of~$\Sigma$. For modern references, see for instance \cite[Section~8.7.2]{Dolgachev} and \cite[Section~7]{OS1}, where it is proven that a general Geiser map is branched along a smooth L\"uroth quartic.

\begin{lemma}
The Geiser map~$\gamma_{\Eig{f}}$ is surjective and its fiber over a point $Q = (a:b:c)$ is given by
%
\begin{equation}
\label{eq:fibers}
  \left\{
  \begin{array}{l}
    a x + by + cz = 0, \\[2pt]
    a \, \partial_x f + b \, \partial_y f + c \, \partial_z f = 0,\\
  \end{array}
  \right.
\end{equation}
%
that is, the intersection between the polar line~$L_Q$ relative to the isotropic conic and~$\mathrm{Pol}_Q f$, the
first polar curve of~$f$ with respect to~$Q$.

In particular, the only possible curves contracted by the Geiser map~$\gamma_{\Eig{f}}$ are lines.
\end{lemma}

\begin{proof}
We observe that for any point $P=(A:B:C) \in \p^2 \setminus \Eig{f}$, the homogeneous coordinates
of the image $\gamma_{\Eig{f}}(P) = \bigl( g_1(P): g_2(P): g_3(P) \bigr)$ are the ones of the unique point in the intersection of the two lines
%
\[
  Ax + By+ Cz = 0, \qquad
  \partial_x f(P) \, x + \partial_y f(P) \, y + \partial_z f(P) \, z = 0.
\]
%
So for any $Q = (a:b:c) \in \p^2$, the fiber~$\gamma_{\Eig{f}}^{-1}(Q)$ consists of the points $P \in \p^2$ such that
%
\begin{equation}
\label{eq:polars}
  Aa + Bb+ Cc = 0, \quad
  \partial_x f(P) \, a + \partial_y f(P) \, b + \partial_z f(P) \, c = 0,
\end{equation}
%
which proves that $\gamma_{\Eig{f}}$ is surjective and that \Cref{eq:fibers} holds.
%To prove that there are no other contracted curves than lines,
%set $Z=E(F)$ and consider the blow-up of the plane $\p^2$ along $Z$. We get a generically finite morphism
%%
%\[
%  \widetilde \gamma_Z \colon \Bl_Z \p^2 \to \p^2.
%\]
%%
%Observe that the fibers of~$\widetilde{\gamma_Z}$ are generically contained in the divisor~$W$ with bihomogenous equation $x_0 y_0 + x_1 y_1 + x_2 y_2=0$.
%Since both~$\Bl_Z \p^2$ and~$W$ are irreducible, and since $\widetilde{\gamma_Z}$ is the restriction of the second projection $p_2 \colon \p^2 \times \p^2 \rightarrow \p^2$, it follows that $S\subseteq W$;
%in particular, every fiber of~$\widetilde{\gamma_Z}$ is contained in a line and by construction the same holds for every fiber of~$\gamma_Z$.
%As a consequence, the map $\gamma_Z$ contracts only lines.
\end{proof}

\begin{prop}
\label{prop:allineati_contrae}
Suppose that $\Eig{f}$ is a zero dimensional eigenscheme of a cubic ternary form.
If $\Eig{f}$ contains a triple of points on a line~$L \subset \p^2$, then the Geiser map~$\gamma_{\Eig{f}}$ contracts~$L$.
\end{prop}

\begin{proof}
%Consider the isomorphism  $\varphi \colon \p (I_{\Eig{F}}(3))^{\vee} \stackrel{\cong}{\longrightarrow} \p (I_{\Eig{F}}(3))$ associating with each point~$(a:b:c)$ the line $ax + by + cz = 0$.
%For any $P \in \p^2 \setminus \Eig{F}$, the point~$\varphi(\gamma_{\Eig{F}} (P))$ corresponds to the line~$L_P$ given by
%
For any $P \in \p^2 \setminus \Eig{f}$, the point~$\gamma_{\Eig{f}} (P)\in \p \bigl(I_{\Eig{f}}(3)^{\vee}\bigr)$ corresponds to the line~$L_P$ given by
\[
  g_1 (P) \, x + g_2(P) \, y + g_3(P) \, z = 0,
\]
%
which in turn determines the following
pencil of cubics $L_P$ in the linear system $\p \bigl(I_{\Eig{f}}(3)\bigr)$:
%
\[
  s \cdot \bigl( g_1(P) \, g_2 - g_2(P) \, g_1 \bigr) + t \cdot \bigl( g_1(P) \, g_3 - g_3(P) \, g_1 \bigr),
\]
%
where $(s: t) \in \p^1$, assuming that $g_1(P) \neq 0$ (the other cases are similar).
It is clear that the base locus of the pencil~$L_P$ contains the eight points of~$\Eig{f} \cup \{P\}$.
So $\gamma_{\Eig{f}} \colon \p^2 \dasharrow \p \bigl( I_{\Eig{f}}(3)^\vee \bigr)$
associates with each point $P \in \p^2 \setminus \Eig{f}$ the pencil of cubics through $\Eig{f} \cup P$.

If $\Eig{f}$ contains a triple of points on a line~$L$, for any $P \in L \setminus \Eig{f}$ such a pencil of cubics has $L$ as a fixed component and the other component varies in a pencil of conics through the remaining $4$ points, so $\gamma_{\Eig{f}}$ is constant on~$L$.
\end{proof}

%We know that $G_1, G_2, G_3$ admit the syzygy $z \, G_1 - y \, G_2 + x \, G_3 = 0$.
%Then, due to the exactness of the Koszul complex of the regular sequence $(x,y,z)$ (see, for instance, \cite[Theorem 7.3.13]{Dolgachev}),
%we know that $G_1, G_2, G_3$ are the $2\times2$ minors of a matrix of the form
%%
%\[
% \left(
% \begin{array}{ccc}
%  x & y & z \\
%  H_1 & H_2 & H_3
% \end{array}
% \right)
%\]
%%
%where $H_1, H_2, H_3$ are quadratic ternary forms.

%Consider now the map
%%
%\[
%  \bar{\gamma} \colon \p^2 \setminus \Eig{f} \rightarrow \p^2, \quad P \mapsto (H_1(P): H_2(P): H_3(P)).
%\]
%%
%This map is the composition of~$\gamma_{\Eig{f}}$ with a linear isomorphism.
%In particular, the map $\bar{\gamma}$ contracts both the lines $P_1 +P_2$ and $P_1 +P_4$.

We now turn back to our situation, when we have five points $P_1, \dots, P_5$ forming a $V$-configuration and satisfying $\delta_2(P_1, \dots, P_5)=0$ and $\rk \, \Phi(P_1, \dotsc, P_5) = 9$. This means that there exists a unique (up to moltiplication by scalars) cubic ternary form~$f$ such that $P_1, \dotsc, P_5 \in \Eig{f}$.

\begin{proof}[Proof of \Cref{proposition:G_split}]
The equation of the line~$P_1 \vee P_2$ can be expressed in the following way
%
\begin{equation}
\label{eq:lineP1P2}
  \left\langle P_1 \times P_2, (x,y,z) \right\rangle = 0 \,;
\end{equation}
%
since $P_1 \vee P_2$ is contracted by~$\gamma_{\Eig{f}}$,
and by the description of the fibers given in \Cref{eq:fibers},
we have that such a line is contained in the conic of equation
%
\[
  \left\langle P_1 \times P_2, (\de_x f, \de_y f, \de_z f) \right\rangle = 0 \,.
\]
%
As a consequence, for any point $\overline{P} = (\bar x: \bar y: \bar z)$ of the line from \Cref{eq:lineP1P2}, the following equations are satisfied:
%
\[
  \left\{
  \begin{array}{l}
    \left\langle P_1 \times P_2, P_1 \right\rangle = 0 \,,\\[2pt]
    \bigl\langle P_1 \times P_2, \overline{P} \bigr\rangle = 0 \,,\\[2pt]
    \bigl\langle P_1 \times P_2, \nabla f (\overline{P}) \bigr\rangle = 0 \,.
  \end{array}
  \right.
\]
%
It follows that the linear system in the variables $X,Y,Z$
%
\[
  \left\{
  \begin{array}{l}
    \bigl\langle P_1, (X,Y,Z) \bigr\rangle = 0 \,,\\[2pt]
    \bigl\langle \overline{P}, (X,Y,Z) \bigr\rangle = 0 \,,\\[2pt]
    \bigl\langle \nabla f (\overline{P}),
    (X,Y,Z) \bigr\rangle = 0 \,.
  \end{array}
  \right.
\]
%
admits as a non-zero solution $P_1 \times P_2$,
hence the determinant of the order~$3$ coefficient matrix
%
\[
  \left(
  \begin{array}{c}
    P_1 \\
    \overline{P} \\
    \nabla f (\overline{P})
  \end{array}
  \right)
\]
%
is zero. This is equivalent to saying that $A_1 g_3 - B_1 g_2 + C_1 g_1$ vanishes at all the points~$\overline{P}$ of the line~$P_1 \vee P_2$.

The argument concerning the line~$P_1 \vee P_4$ can be repeated by replacing~$P_2$ with~$P_4$.
\end{proof}
%
\begin{prop}
\label{proposition:third_alignment}
If we have a $V$-configuration of five points $P_1, \dots, P_5$
such that the rank of the matrix $\Phi(P_1, \dots, P_5)$ is $9$ and
such that $\delta_2(P_1, \dots, P_5) = 0$,
then the unique ternary cubic determined by the condition that $P_1, \dots, P_5$
are eigenpoints has also the two other eigenpoints~$P_6$ and~$P_7$
aligned with~$P_1$.
\end{prop}
\begin{proof}
Let $f$ be the unique form such that $P_1, \dotsc, P_5 \in \Eig{f}$.
According to \Cref{proposition:G_split}, the polynomial
$C_1g_1-B_1g_2+A_1g_3$ splits into three factors $r_1$, $r_2$, $r_3$,
linear in $x, y, z$, which
correspond to the line~$P_1 \vee P_2$, the line~$P_1 \vee P_4$, and the line~$P_6 \vee P_7$.
The factorization of~$f$ gives that the factor~$r_3$ is:
%
\[
  r_3 = 2yv_1A_1B_1^4A_2^2B_2A_4^3-2xv_1B_1^5A_2^2B_2A_4^3+\cdots
  -2 yv_1A_1B_1^4A_2^2B_2A_4^3
\]
%
and is composed by almost $2000$ monomials. When we
evaluate~$r_3$ on the coordinates of~$P_1$, we get zero.
\end{proof}


\section{\texorpdfstring{$V$}{V}-configurations of rank~\texorpdfstring{$8$}{8}}
\label{rank_8}
%% file contiCasoDegenere2.sage and file conf_sigma12_sigma14.sage
%% molti altri file via via indicati nei commenti di latex
%
In this section, we study the possible configurations of
eigenpoints for cubics with a $V$-configuration
satisfying the condition $\rk \, \Phi(P_1, \dots, P_5) = 8$. According
to~\Cref{theorem:rank_V}, we have to distinguish the two cases
%
\begin{gather}
  \delta_1(P_1, P_2, P_4)=\overline{\delta}_1(P_1, P_2, P_3) =
  \overline{\delta}_1(P_1, P_4, P_5) = 0 \,,
  \label{rk8_1} \\
  \sigma(P_1, P_2) = 0, \ \sigma(P_1, P_4) = 0 \ \ \mbox{and} \ \ s_{22} = 0,
  \ s_{44} = 0 \,.
  \label{rk8_2}
\end{gather}
%
(the latter condition means that \emph{the two lines of
the $V$-configuration are tangent to $\iso$ in $P_2$ and~$P_4$}).
%
To start, we characterize when, in a $V$-configuration
$P_1, \dotsc, P_5$, the point~$P_1$ is singular.

\begin{prop}
\label{proposition:P1_sing}
If $P_1, \dots, P_5$ is a $V$-configuration and if
$\mathcal{C}$ is a cubic which
has $P_1, \dots, P_5$ among its eigenpoints, then $P_1$ is
singular for~$\mathcal{C}$ if
and only if $P_1$ is not on $\iso$ and one of the following conditions
is satisfied:
%
\begin{enumerate}
  \item $\delta_1(P_1, P_2, P_4) = 0$ and $\overline{\delta}_1(P_1, P_2, P_3) = 0$;
  \item $\delta_1(P_1, P_2, P_4) = 0$ and $\overline{\delta}_1(P_1, P_4, P_5) = 0$;
  \item $\overline{\delta}_1(P_1, P_2, P_3) = 0$ and
  $\overline{\delta}_1(P_1, P_4, P_5) = 0$.
\end{enumerate}
%
\end{prop}
\begin{proof}
%% file: singularCubic_1ii0.sage e singularCubic_100.sage
As usual, we can split the problem into two cases: $P_1 = (1: \iii: 0)$
and $P_1 = (1: 0: 0)$. In both cases, we construct the matrix
$\Phi(P_1, \dots, P_5)$ (for $P_2$, $P_4$ generic, $P_3$ aligned with
$P_1$ and~$P_2$ and~$P_5$ aligned with $P_1$ and~$P_4$). To this matrix we add
a row which correspond to the condition that $P_1$ is singular, obtaining
in this way a matrix~$M$. After
suitable manipulations of~$M$, we see that $\Lambda(M)$ is empty
in the case $P_1=(1: \iii: 0)$, while if $P_1$ is the point~$(1:0:0)$
and one of the three above conditions is satisfied, then
$\dim \bigl( \Lambda(M) \bigr) \geq 0$.
\end{proof}

Let us start to analyze the two conditions from \Cref{rk8_1} and \Cref{rk8_2}.
The three equalities of~\Cref{rk8_1} give that,
using \Cref{rmk_delta_case2}, either $P_1$ is on~$\iso$ and
$P_1\vee P_2$ is the tangent line to~$\iso$ in~$P_1$ (but this is not compatible
with the equation $\overline{\delta}_1(P_1, P_4, P_5)=0$), or,
from \Cref{rmk_delta_case1}, we can choose the points as follows:
%
\begin{itemize}
  \item $P_1$ and $P_2$ in an arbitrary way;
  \item $P_4$ in the $\p^1$
  space of points orthogonal to $s_{11} \, P_2 - s_{12} \, P_1$;
  \item $P_3 = (s_{12}^2+s_{11}s_{22}) \, P_1 - 2s_{11}s_{12} \, P_2$;
  \item $P_5 = (s_{14}^2+s_{11}s_{44}) \, P_1 - 2s_{11}s_{14} \, P_4$.
\end{itemize}
%

For such a choice
of points, the matrix $M = \Phi(P_1, \dots, P_5)$ has rank $8$
%% verifica nel file matrix_of_order_8.sage
and therefore $\dim \Lambda(M) = 1$ and the dimension of the variety
of the corresponding cubics is $6$.
Moreover, from~\Cref{proposition:P1_sing}, we have that all these cubics
are singular in $P_1$.
If we take a random point of this variety, we see that the corresponding
cubic has $7$ distinct eigenpoints (is as expected, singular in $P_1$) and the
$7$ points do not satisfy other collinearities in addition to those of the
$V$-configuration.

\begin{rmk}
\label{rmk:particular_cases}
If $P_1, \dots, P_5$ satisfy the conditions~(\ref{rk8_1}), as seen,
in general a cubic with these eigenpoints has only $(P_1, P_2, P_3)$ and $(P_1, P_4, P_5)$ aligned, but
does not have the collinearity $(P_1, P_6, P_7)$. Since conditions~(\ref{rk8_1})
implies that $\delta_2(P_1, \dots, P_5) = 0$, we see
that the hypothesis ``rank $9$''
in~\Cref{proposition:third_alignment} is necessary.
\end{rmk}

Concerning possible sub-cases with further collinearities of the points,
the following results hold:
%%%%
%% tutti i conti si trovano nei file:
%% three_deltas_I.sage
%% three_deltas_II.sage
%% three_deltas_III.sage
%% three_deltas_IV.sage
%% c'e' anche il file three_d_alt.sage che puo' essere di aiuto
%% ma non e' essenziale
%%
%% c'e' anche il file three_d_alt.sage che calcola tutti gli autopunti
%% in configurazione (8) nel caso dei 3 delta = 0.
%%
\begin{prop}
\label{three_d_three_alignments}
If the five points $P_1, \dots, P_5$ satisfy \Cref{rk8_1},
then, in $\Lambda \bigl( \Phi(P_1, \dotsc, P_5)\bigr)$ there is
a cubic curve with $7$ eigenpoints with the following three alignments:
$(P_1, P_2, P_3)$, $(P_1, P_4, P_5)$ and $(P_1, P_6, P_7)$. \\
No choices of $P_1, \dots, P_5$ allow one to obtain further alignments of the
$7$ eigenpoints.
\end{prop}
%
\begin{proof}
%% file: three_deltas_I.sage and three_deltas_II.sage
Using the action of~$\SO_3(\C)$ (see~\Cref{two_orbits}),
we can assume $P_1= (1: \iii: 0)$ or $P_1= (1: 0: 0)$. As observed
in the proof of~\Cref{theorem:rank_V}, if $P_1 = (1: \iii: 0)$,
%% Vedi file rank_8_2_1_ii_0.sage+)
the matrix
$\Phi(P_1, \dots, P_5)$ cannot have rank smaller then $9$, so the only
case to consider is given by $P_1 = (1: 0: 0)$. Hence
$P_2 = (A_2: B_2: C_2)$ and~$P_3$, $P_4$ and~$P_5$ can be chosen according 
to the above formulas.\\
Consider the matrix $\Phi(P_1, \dots, P_5)$ and select $8$ linearly independent
rows (since $P_1$ is numeric, it can be seen that for instance the rows
of position $0, 1, 4, 5, 7, 8, 12, 14$ are a good choice and avoid to consider
many exceptions) and let $M_1$ be a matrix obtained in this way.
We fix two numeric $10$-dimensional vectors~$V_a$ and~$V_b$ and we construct
$M_a$ (resp.\ $M_b$) obtained by horizontally stacking vector
$V_a$ (resp.\ $V_b$)
to $M_1$. The rank of~$M_a$ and~$M_b$ is~$9$ (unless an unlucky choice
of the vectors, in which case they can be changed). The matrices~$M_a$ and~$M_b$ 
play the same role of the matrix~$\mathcal{H}$
of~\Cref{rank_9}, hence we can do the same construction explained
there, in order to obtain the line~$r_a$ passing through the two
eigenpoints~$P_6$ and $P_7$ of the cubic obtained from~$M_a$ and the
line~$r_b$ passing through the analogous eigenpoints of the cubic
obtained from~$M_b$.
Hence the line $r = w_ar_a+w_br_b$
(where $w_a, w_b$ are parameters) is the line passing through the eigenpoints
$P_6$ and~$P_7$
of the generic cubic of the one dimensional family given by the line
of~$\p^9$ determined by the cubic obtained from $M_a$ and
the cubic obtained from $M_b$. The condition $P_1 \in r$
determines the values of~$w_a$ and~$w_b$, hence we get the cubic
with the three collinearities $(P_1, P_2, P_3)$,
$(P_1, P_4, P_5)$ and~$(P_1, P_6, P_7)$. If we require that the eigenpoints 
have other alignments, w.l.o.g.\ we can assume that $P_6$ is on the line 
$P_2 \vee P_4$, hence it can be determined as the intersection of~$r$
and~$P_2 \vee P_4$. If we impose that this point is an eigenpoint, we get
a contradiction, thus there are no other sub-cases to consider.   
\end{proof}
%
Moreover, we have:
%
\begin{prop}
\label{prop:d2_6allin}
If the five points $P_1, \dots, P_5$ satisfy~(\ref{rk8_1})
and if we impose the condition that there is an eigenpoint, say $P_6$,
which is aligned with $P_2$ and~$P_4$ then the eigenpoints satisfy all these
alignments:
%
\[
  (P_1, P_2, P_3), (P_1, P_4, P_5),
  (P_2, P_4, P_6), (P_2, P_5, P_7),
  (P_3, P_4, P_7), (P_3, P_5, P_6)
\]
%
Hence the points~$P_6$ and~$P_7$ are determined by $P_1, \dots, P_5$
since
$P_6 = (P_2 \vee P_4) \cap (P_3 \vee P_5)$
and $P_7 = (P_3 \vee P_4) \cap (P_2 \vee P_5)$.
(Of course, a similar result holds if we take $P_3$ in place of~$P_2$ or~$P_5$
in place of~$P_4$).
\end{prop}
%
\begin{proof}
%% file three_deltas_III.sage, three_deltas_IV.sage
We define the points $P_1, \dots, P_5$ as in the previous proposition, then 
we consider a point of the line $P_2 \vee P_4$ and we impose that it is 
an eigenpoint (i.e.\ we impose that $\rk \, \Phi(P_1, \dotsc, P_6) < 10$
This request defines a point~$P_6$ and it turns out that $P_6$ is also aligned 
with~$P_3$ and~$P_5$. We then define $P_7$ as the point 
$(P_2 \vee P_5) \cap (P_3 \vee P_4)$. It turns out that 
$\rk \, \Phi(P_1, \dots, P_7) < 10$, so $P_7$ is an eigenpoint. The seven 
points $P_1, \dotsc, P_7$ satisfy the above collinearities.
\end{proof}
\Cref{three_d_three_alignments} and \Cref{prop:d2_6allin} exhaust all the possible configurations
of collinearities of the eigenpoints in case of condition~(\ref{rk8_1}).

%

Now we consider the \Cref{rk8_2}.
%% La dimostrazione della proposizione prop:rk8_2B si trova nel file
%% rank_8_twoTangIso.sage
%% Il file confV_tg_iso.sage fa il caso generale (cioe' quando P1
%% e' il punto (A1, B1, C1) e non (1, 0, 0)) e serve per provare
%% le formule realtive a P6 e P7 nella configurazione (5).
%% Il file confV2_tg_iso.sage contiene i conti per la configurazione (8)
%% e prova la formula P6 = s15*P3+s13*P5

%
A generic cubic which satisfies the condition~(\ref{rk8_2}) is
a cubic of the one-dimensional linear system
$\Lambda\bigl(\Phi(P_1, \dotsc, P_5)\bigr)$, where:
%
\begin{itemize}
  \item $P_1$ is any point of the plane (not on $\iso$);
  \item $P_2$ and $P_4$ are the two tangency points to $\iso$ given by the lines from~$P_1$;
  \item $P_3$ is any point on the line~$P_1 \vee P_2$ (different from~$P_1$ and~$P_2$)
  and $P_5$ is any point on the line~$P_1 \vee P_4$ (different from~$P_1$ and~$P_4$).
\end{itemize}
%
The variety of all the cubics with a $V$-configuration
that satisfies condition~(\ref{rk8_2}) is therefore five-dimensional.
The five points satisfy the condition $\delta_2(P_1, P_2, P_3, P_4, P_5) = 0$.
%% 2+1+1+1, l'ultimo 1 perche' rango matrice = 8


The reciprocal position of the eigenpoints of the cubics of this family
is described by the following:
%
\begin{prop}
\label{prop:rk8_2B}
The generic cubic of the family of cubics which satisfy condition~(\ref{rk8_2})
has seven eigenpoints with the alignments:
%
\[
  (P_1, P_2, P_3), \ (P_1, P_4, P_5), \ (P_1, P_6, P_7)
\]
%
Among these points we have the relation
$\scl{P_1 \times P_6}{P_3\times P_5}=0$
(i.e.\ the lines~$P_1 \vee P_6$ and~$P_3 \vee P_5$ are orthogonal).
%% ortogonalita': provata nel file rank_8
In the family there is a sub-family of cubics whose eigenpoints have the following alignments:
%
\[
  (P_1, P_2, P_3),\ (P_1, P_4, P_5),\ (P_1, P_6, P_7),\ (P_2, P_4, P_6).
\]
%
In this case the points~$P_6$ and~$P_7$ are given by the formulas:
%
\begin{equation}
\label{formuleP6_P7}
P_6 =  s_{15}s_{34}\, P_2 + s_{13}s_{25}\, P_4, \quad 
P_7 = s_{15}(s_{26}s_{46}+s_{24}s_{66})\, P_1+ s_{11}s_{24}s_{56}\, P_6
\end{equation}
%
and a sub-family whose eigenpoints have the following
alignments:
%
\[
  (P_1, P_2, P_3),\ (P_1, P_4, P_5), \ 
  (P_1, P_6, P_7),\ (P_2, P_5, P_6), \ 
  (P_3, P_4, P_6),\ (P_3, P_5, P_7)
\]
%
In this case the point~$P_6$ (given, for instance, by the formula
$P_6 = s_{15} \, P_3 + s_{13} \, P_5$) is obtained as
the intersection~$(P_2 \vee P_5) \cap (P_3 \vee P_4)$ and consequently
$P_7 = (P_1 \vee P_6) \cap (P_3 \vee P_5)$.\\
No other collinearities among the eigenpoints are possible.
\end{prop}
%
\begin{proof}
%% file rank_8_twoTangIso.sage
By consider again the action of~$\SO_3(\C)$, we can assume:
%
\[
  P_1 = (1: 0: 0), \quad
  P_2 = (0: \iii: 1), \quad
  P_4 = (0: -\iii: 1)
\]
%
(and $P_3 = u_1 \, P_1 + u_2 \, P_2$, $P_5 = v_1 \, P_1 + v_2 \, P_4$). 
In this situation, it
is easy to see that $\Lambda\bigl(\Phi(P_1, \dots, P_5)\bigr)$
is the following family of cubic curves:
%
\[
  \mathcal{C}(l_1, l_2) = l_1 \, \mathcal{C}_1 + l_2 \, \mathcal{C}_2
\]
%
where $l_1, l_2$ are parameters and
%
\begin{align*}
  \mathcal{C}_1 & = x \cdot \left(2x^{2} + 3 y^{2} + 3 z^{2}\right)\\
  \mathcal{C}_2 & = (y + \iii z) \cdot (y - \iii z)
  \cdot \bigl(2 \iii x u_{2} v_{2} + y (u_{2} v_{1}- u_{1} v_{2})
  - \iii z (u_{2} v_{1} + u_{1} v_{2})\bigr)
\end{align*}
%
%% C_1 and C_2 are in the file rank_8_twoTangIso.sage, see ***C1_C2***
%%
The explicit equation of the cubics $\mathcal{C}(l_1, l_2)$ allows us to
construct the ideal of the seven eigenpoints and (after
saturations w.r.t.\ the five ideals of the points $P_1, \dotsc, P_5$),
we get an ideal generated by a line~$r$ and a conic~$c$, whose zeros are
the points~$P_6$ and~$P_7$. Since the line~$r$ contains the point~$P_1$,
we have that $P_1, P_6, P_7$ are collinear and $r$ is orthogonal to~$P_3 \vee P_5$.
In order to see if there are further collinearities among the
eigenpoints, we have to distinguish three cases (up to permutation
of the indices): $P_2, P_4, P_6$ are collinear or $P_2, P_5, P_6$ are
collinear, or $P_3, P_5, P_6$ are collinear. In the first case, $P_6$
is the point $r \cap (P_2\vee P_4)$. If it is an eigenpoint, it must be
on $c$, hence $l_1$ and~$l_2$ have a specific value which gives a sub-family
of~$\mathcal{C}(l_1, l_2)$ and we can compute the explicit coordinates of
all the seven eigenpoints of the cubics of this family. The other two
cases can be studied in a similar way.
%% file rank_8_twoTangIso.sage, file confV_tg_iso.sage, file
%% confV2_tg_iso.sage.
\end{proof}
\begin{rmk}
\label{rmk:three_orthog} For further references, it is perhaps worth noting that the explicit 
construction of the 
points in~\Cref{prop:d2_6allin} and~\Cref{prop:rk8_2B} allows one to verify
several relations among them. In particular, the seven eigenpoints
of~\Cref{prop:d2_6allin} are such that
\[
\scl{P_1\times P_2}{P_1 \times P_4}=0, \quad 
\scl{P_2\times P_4}{P_3 \times P_5}=0, \quad 
\scl{P_3\times P_4}{P_2 \times P_5}=0;
\]
and it holds:
\[
P_4 = (P_2\times P_3)s_{25}s_{35}-s_{23}(P_2\times P_5)s_{35}+ s_{23}s_{25}(P_3\times P_5).
\]
In the first sub-family described in~\Cref{prop:rk8_2B} we have 
$P_1 = P_2 \times P_4$, hence 
\[
\scl{P_1\times P_2}{P_2 \times P_4}=0, \quad 
\scl{P_1\times P_4}{P_2 \times P_4}=0, \quad 
\scl{P_1\times P_6}{P_2 \times P_4}=0
\]
and in the second family:
\[
\scl{P_1\times P_4}{P_3 \times P_4}=0, \quad 
\scl{P_1\times P_2}{P_2 \times P_5}=0, \quad 
\scl{P_3\times P_5}{P_1 \times P_6}=0.
\]
hence it holds:
\[
P_3 = (P_1 \times P_5)s_{16}s_{56}-s_{15}(P_1\times P_6)s_{56}+s_{15}s_{16}(P_5 \times P_6)
\]
\end{rmk}
%

% \begin{rmk}
% \label{rmk:delta1_and_delta2}
% It is easy to verify that in
% a $V$-configuration which satisfies \Cref{C8:cnd2}, we have
% $\delta_1(P_1, P_2, P_4)\neq 0$ and
% $\delta_2(P_1, P_2, P_3, P_4, P_5) = 0$,
% while in \Cref{C8:cnd1} both $\delta_1(P_1, P_2, P_4)$ and
% $\delta_2(P_1, P_2, P_3, P_4, P_5)$ are zero.
% \end{rmk}


\section{The locus of cubic ternary forms with an aligned triple of eigenpoints}
\label{locus_one_alignment}

The results of the previous sections allow one to determine the dimension and the degree of the locus of cubics having
at least one aligned triple of eigenpoints. We first recall a result by Abo concerning the locus of cubics with a non reduced or positive-dimensional eigenscheme.

\begin{definition}
An order~$d$ tensor $T \in (\C^n)^{\otimes d}$ is called \emph{regular} if $\Eig{T}\subset \p^{n-1}$ is reduced of dimension~$0$.

We set
%
\[
 \Delta_{n,d} := \{[T]\in \p \bigl( (\C^n)^{\otimes d} \bigr) \ | \ \Eig{T} \textrm{\ is \ not \ regular} \} \subset \p \bigl( (\C^n)^{\otimes d} \bigr) \,.
\]
%
The locus~$\Delta_{n,d}$ is called the \emph{eigendiscriminant} (see \cite[Definition 5.5]{Abo}).
\end{definition}

\begin{theorem} (\cite[Corollary 5.8]{Abo})
\label{theorem:eigendiscriminant}
The eigendiscriminant $\Delta_{n,d}\subseteq \p \bigl( (\C^n)^{\otimes d} \bigr)$ is an irreducible hypersurface
of degree~$n(n-1)(d-1)^{n-1}$.
\end{theorem}

The statement of \Cref{theorem:eigendiscriminant} is still valid when restricting to symmetric tensors, as they form a linear subspace of~$\p \bigl( (\C^n)^{\otimes d} \bigr)$ and the generic symmetric tensor has a regular eigenscheme.

In the particular case $n=d=3$, we have that
the eigendiscriminant hypersurface has degree
%
\[
  \deg \Delta_{3,3} = 24 \,.
\]
%

\begin{definition}
\label{def:locus_L}
Let $\sU \subset \p^9$ be the following locus:
%
\[
  \sU:= \{[f]\in \p^9 \setminus \Delta_{3,3} \ | \ \Eig{f} \ \textrm{contains \ an \ aligned \ triple}\}, \,
\]
%
and define $\sL \subseteq \p^9$ as the closure of~$\sU$:
%
\[
  \sL := \overline \sU \subset \p^9 \,.
\]
%
\end{definition}

\begin{theorem}
\label{theorem:irreducible}
The variety~$\sL$ is an irreducible hypersurface.
\end{theorem}

\begin{proof}
We first prove that $\dim \sL = 8$. As there exist
polynomials $f$ such that $\Eig{f}$ has no aligned triple, we have $\dim \sL \le 8$.

Consider the symmetric product $(\p^2)^{(3)}$ and set $\mathcal{AL} \subset (\p^2)^{(3)}$ to be the locus of unordered triples of distinct aligned points. Observe that $\mathcal{AL}$ is irreducible of dimension $5$ as its closure is the hypersurface given by the vanishing of the determinant of the order~$3$ matrix of the coordinates of three general points. Alternatively, the closure of the locus $\mathcal {AL}$ can be seen as the symmetric quotient of the projective line bundle $\sF \subset \p^2 \times \p^2 \times \p^2$ parametrized by the triples $(P_1, P_2, u_1 P_1 +u_2P_2)$, where $(u_1:u_2) \in \p^1$.

If we set
%
\[
  \sU':= \{[f]\in \p^9 \setminus \Delta_{3,3} \ | \ \Eig{f} \ \textrm {contains \ exactly \ $1$ \ aligned \ triple}\} \,,
\]
%
we have a surjective morphism
%
\[
  \alpha \colon \mathcal{U}' \to \mathcal{AL} \,,
\]
%
assigning to each $[f] \in \mathcal{U}'$ the unique triple of eigenpoints.
The fibers of~$\alpha$ are projective linear systems of dimension~$3$ by \Cref{manca il riferimento su ancillary  non e': condition_rank_aligned} over the irreducible open subset
%
\[
  \mathcal{W} := \mathcal{AL}
  \setminus \{(P_1,P_2,P_3) \in \mathcal{AL}
  \ | \ \sigma(P_1,P_2)=0, s_{11} s_{22} s_{33}=0\}
\]
%
of aligned triples not lying
on a tangent line to the isotropic conic~$\iso$ and with tangency point one of the~$P_i$'s. 
By the Fiber Dimension Theorem, this implies that the open subset $\alpha^{-1} (\mathcal{W}) \subset \sU$ is irreducible of dimension
%
\[
  \dim \alpha ^{-1} (\mathcal{W}) = 8 \,,
\]
%
so $\sL$ is a hypersurface.

To prove the irreducibility, we observe that since the eigenpoint condition corresponds to two linear conditions on the coefficients of cubic polynomials, by linear algebra methods we can determine a local parametrization of~$\sL$.
Specifically, given a general aligned triple $(P_1, P_2, u_1 P_1 +u_2P_2)$, the matrix of conditions $\Phi(P_1, P_2, u_1 P_1 +u_2P_2)$ has rank $6$, so the associated linear system
$\Lambda \bigl( \Phi(P_1, P_2, u_1 \, P_1 + u_2 \, P_2) \bigr)$ has dimension~$3$. We can then express $6$ coefficients of the general polynomial
$[f]\in \Lambda \bigl( \Phi(P_1, P_2, u_1 \, P_1 + u_2 \, P_2) \bigr)$ as rational function of $P_1,P_2,u_1,u_2$ and of the remaining $4$ coefficients. The free coefficients depend on the position of a non zero minor of order six, which, in principle, could involve any choice of columns. In fact this is not the case, as it turns out, that the last $3$ columns of the matrix
$M = \Phi(P_1, P_2, u_1 \, P_1 + u_2 \, P_2)$ are always in the linear span of the first $7$ columns.

If the three components of~$P_1 \times P_2$ are called $\alpha, \beta, \gamma$, let:
%
\[
  N_1 = \left(
  \begin{array}{ccc}
    \alpha & 0 & 0 \\
    0 & \beta & 0\\
    0 & 0 & \gamma
  \end{array}
  \right), \quad
  N_2 = \left(
  \begin{array}{ccc}
    0 & \alpha & 0 \\
    \gamma & 0 & 0\\
    0 & 0 & \beta
  \end{array}
  \right).
\]
%
Moreover, let $c_0, \dots, c_9$ be the $10$ columns of~$M$. Then the six
columns $c_0, c_1, c_2, c_4, c_5, c_7$ of~$M$ are linearly dependent. More
precisely, if $L_1$ is the $9\times 3$ matrix whose columns are
$c_0, c_2, c_7$ and $L_2$ is given by the columns $c_1, c_4, c_5$,
then we have:
%
\begin{equation}
  (L_1 N_1 + L_2N_2) (P_1 \times P_2) = 0
  \label{combLin}
\end{equation}
%
(the linear combination of $c_0, c_1, c_2, c_4, c_5, c_7$ which is zero
is obtained expanding this expression).
Similarly, the columns $c_1, c_2, c_3, c_5, c_6, c_8$ are linearly dependent
and~(\ref{combLin}) holds if $L_1$ in this case is $[c_1, c_3, c_8]$ and
$L_2$ is $[c_2, c_5, c_6]$. Finally, the columns
$c_4, c_5, c_6, c_7, c_8, c_9$ are linearly dependent and~(\ref{combLin})
holds if we take $L_1 = [c_4, c_6, c_9]$ and $L_2 = [c_5, c_7, c_8]$.

As a consequence, local parametrizations of~$\sL$ can be given by considering the following open subsets:
for any multiindex
%
\[
  I = \{i_1, \dots, i_6\} \subset \{0, 1, \dots, 6\} \,,
\]
%
we set
%
\begin{multline*}
  \sV_I :=
  \bigl\{
    [(P_1, P_2, u_1 P_1 +u_2P_2)] \in \mathcal{AL} \ | \ \text{the\ columns\ of\ } \Phi
    (P_1, P_2, u_1 \, P_1 + u_2 \, P_2) \\
    \text{\ relative \ to} \ I\ \text{are\ independent}
  \bigr\} \,.
\end{multline*}
%
Then for any $[(P_1, P_2, u_1 \, P_1 + u_2 \, P_2)] \in \sV_I$, the coefficients of any element
$[f] =[\mathcal{B} \cdot w_f]$ of the linear system
$\Lambda \bigl( \Phi (P_1, P_2, u_1 P_1 + u_2P_2) \bigr)$ can be expressed in the form
%
\begin{equation}
\label{eq: parametrizzazione}
  w_f = (b_0,
  b_1,
  b_2,
  b_3,
  b_4,
  b_5, b_6,b_7,b_8,b_9) = w_f(P_1,P_2,u_1,u_2,b_j,b_7,b_8,b_9) \,,
\end{equation}
%
where $b_i = b_i(P_1,P_2,u_1,u_2,b_j,b_7,b_8,b_9)$ for $i\in I$ and $j \not \in I$, $1\le j \le 7$, are suitable rational functions.
On such an open subset, we have then the rational map
%
\[
  \alpha_I \colon \sV_I \times \p^3 \to \sL, \quad
  \alpha_I (P_1,P_2,u_1,u_2,b_j,b_7,b_8,b_9)=
  [\mathcal{B} \cdot w_f(P_1,P_2,u_1,u_2,b_j,b_7,b_8,b_9)] \,.
\]
%
Next we observe that for any $I \subset \{1,\dots, 7\}$, the image
%
\[
  W_I := \alpha_I (\sV_I \times \p^3)
\]
%
is irreducible, being image of an irreducible variety, and it
contains an open subset. Indeed, assume for simplicity (the other cases are similar) that
$I=\{1,2,3,4,5,6\}$; an element $[f]\in W_I$ is represented by the determinant of the
matrix given by the rows $0,1,3,4,6,7$ of $\Phi (P_1, P_2, u_1 \, P_1 + u_2 \, P_2)$ and the additional $4$ rows
%
\[
  \left(
  \begin{array}{cccccc}
    0 & \cdots & 1&0&0&-b_6 \\
    0 & \cdots & 0&1&0&-b_7 \\
    0 & \cdots & 0&0&1&-b_8 \\
    & & & \mathcal{B} & & \\
  \end{array}
  \right) \,.
\]
%
In particular, the coefficient $b_9$ of~$z^3$ is equal to the determinant of the order~$6$ minor of the matrix $\Phi (P_1, P_2, u_1 \, P_1 + u_2 \, P_2)$ relative to the first $6$ columns. It follows that $W_I \supset \sL \cap \{ a_9 \neq 0\}$. 

Finally, we claim that the irreducible subvarieties $W_I$ for $I\subset \{1,\dots, 7\}$ have a common non-empty intersection
including internal points. Indeed, as one can verify with an explicit symbolic computation, an example of a polynomial
class~$[g]$ with an aligned triple of eigenpoints and satisfying
%
\[
  [g] \in \bigcap_I W_I \cap \Delta_{3,3}
\]
%
is given by a random choice of~$P_1$, $P_2$ and $(u_1:u_2)$ and $b_7,b_8,b_9$.

Hence $\sL$ is covered by irreducible open subsets, each intersecting every other, so it is irreducible.
\end{proof}

\begin{prop}
The locus $\sV \subset \sL \subset \p^9$ of cubic forms with two or more aligned triples of eigenpoints has dimension
%
\[
  \dim \sV = 7.
\]
%
\end{prop}
\begin{proof}
By \Cref{theorem:rank_V}, five points forming a $V$-configuration are eigenpoints of some cubic form of and only if $\delta_1 (P_1,P_2,P_4) \cdot
\delta_2 (P_1,P_2,P_3,P_4,P_5)=0$.
Moreover, there is an open subset of the space of $V$-configurations, where the rank of the condition matrix is~$9$. 
As such a space has dimension~$8$ and the two equations define two divisors, such a locus is of dimension~$7$ and it is birational to~$\sV$.
\end{proof}

As a consequence, we have the following result.

\begin{corollary}
\label{lemma:pencil_one_aligned}
If $[f],[g] \in \p^9 \setminus \Delta_{3,3}$ are general cubics,
then the general cubic in the pencil
$\lambda f + \mu g$ for $(\lambda: \mu) \in \p^1$ has no aligned triples of eigenpoints, and there is a finite number of cubics having exactly one aligned triple.
\end{corollary}

%\begin{definition}
% We define $\Delta \subset \sL$ to be the closure of the locus of cubics with at least two aligned triples of eigenpoints.
%\end{definition}
%
%\begin{prop}
%  The variety~$\Delta$ has dimension~$7$ and it is the union of two irreducible components~$\Delta_1$ and~$\Delta_2$.
%\end{prop}

In what follows, we use the following notation: if $g_1, g_2, g_3$ are the three minors of \Cref{eq:def_matrix} relative
to a cubic form~$f$, we denote by
%
\[
  \Sigma_f := \p \bigl( \left\langle g_1, g_2, g_3 \right\rangle \bigr),
\]
%
the net of cubics, whose base locus is the eigenscheme~$\Eig{f}$.

\begin{lemma}
\label{lemma:scroll}
If $f$ and $g$ are general cubics, then
%
\[
  \mathcal{N} := \bigcup_{(\lambda : \mu) \in \p^1} \Lambda_{\lambda f + \mu g} \subset \p^9
\]
%
is an embedding of a rational projective bundle and has degree~$3$.
\end{lemma}
\begin{proof}
Consider the projective bundle given by the family of planes
%
\[
  \mathcal{P} := \{ \Lambda_{\lambda f + \mu g} \, : \, (\lambda: \mu) \in \p^1 \} \subset \p^1 \times \p^9
\]
%
Then $\mathcal{N}$ is the projection of~$\mathcal{P}$ on the second factor.
However, the map $\mathcal{P} \to \mathcal{N}$ contracts no subvariety of any plane of~$\mathcal{P}$, so either it is an embedding or it contracts some horizontal curve. In the latter case, all the planes of the family should intersect in at least one point. In particular, the two nets $\Lambda_f$ and~$\Lambda_g$ should have non-empty intersection.
If we denote by~$g_1$, $g_2$ and~$g_3$ the $2 \times 2$ minors relative to~$f$, and by~$h_1$, $h_2$ and~$h_3$ the ones relative to~$g$, the vectorial dimension of the linear span $\left\langle g_1, g_2, g_3, h_1, h_2, h_3 \right\rangle$ should be strictly less than $6$. This can be avoided, since such a condition corresponds to a proper closed subscheme of~$\p^9 \times \p^9$.

It follows that if $f$ and~$g$ are general enough, then $\mathcal{N}$ is a $3$-dimensional rational normal scroll in
%
\[
  \mathcal{N} \subset \p(\left\langle g_1, g_2, g_3, h_1, h_2, h_3 \right\rangle) \cong \p^5.
\]
%
Being a variety of minimal degree, its degree is $5+1-3 = 3$.
\end{proof}

\begin{theorem}
The degree of~$\sL$ is equal to
%
\[
  \deg \ \sL = 15 \,.
\]
%
\end{theorem}

\begin{proof}
We start by observing that a reduced $0$-dimensional eigenscheme contains an aligned triple if and only if the net of cubics
$\Lambda_f = \left\langle g_1, g_2, g_3 \right\rangle$ contains a cubic which splits in three lines, a so called \emph{triangle}. Moreover, if $f$ is general enough, we have exactly one aligned triple and the other $4$ points are in general position; in this case, the net~$\Lambda_f$ contains exactly three triangles, namely the unions of the line passing through the aligned triple and the reducible conics through the $4$ points in general position.

To determine the degree of~$\sL$ we consider a general pencil of cubic forms $\lambda f + \mu g$, and we compute the number of elements with associated net $\Lambda_{\lambda f + \mu g}$ containing a triangle.

To this aim, denote by $\mathcal{T} \subset \p^9$ the variety of triangles; it is a classical result that its dimension is~$6$ and its degree is~$15$,
see for instance \cite[Section~2.2.2]{3264}. We now consider the variety~$\mathcal{N}$ from \Cref{lemma:scroll}.
Note that, since each net containing a triangle, actually contains exactly $3$ of them, the number of nets of~$\mathcal{N}$ containing some triangle is given by
%
\[
  \frac{\mathcal{T} \cdot \mathcal{N}}{3} = \frac{{15} \cdot {3}}{3} = 15 \,.
\]
%
This implies that $\deg \sL = 15$.
\end{proof}

\begin{es}
The following pencil of cubic forms admits exactly $15$ cubics with an aligned triple of eigenpoints ..
\end{es}


\section{Eigenschemes of positive dimension}
\label{positive_dim}

In this section, we consider positive dimensional eigenschemes.
By \cite{BGV}, an eigenscheme cannot be of pure dimension~$1$, and the possible $1$-dimensional components are of degree~$1$ or~$2$.
In what follows, we shall determine the degree of the $0$-dimensional residual scheme in both cases.

\begin{prop}
\label{p2}
Let $C = V(f) \subset \p^2$ be a cubic curve.
Assume that $\dim \Eig{f} = 1$.
%
\begin{enumerate}
  \item If the $1$-dimensional component of~$\Eig{f}$ is a line~$l$,
  then the residual subscheme $Z := \mathrm{Res}_l \bigl( \Eig{f} \bigr)$ in~$\Eig{f}$ with respect to~$l$ has degree~$3$. Moreover, the scheme $Z$ is not contained in a line.
  \item If the $1$-dimensional component of~$\Eig{f}$ is a conic~$q$,
  then the residual subscheme $Z := \mathrm{Res}_q \bigl( \Eig{f} \bigr)$ in~$\Eig{f}$ with respect to~$q$ has degree~$1$.
\end{enumerate}
%
\end{prop}

The proof relies on the exactness of the Koszul complex associated with the regular sequence~$x$, $y$, and~$z$, specifically we use the following result (see \cite[Theorem~7.3.13]{Dolgachev} or \cite[Lemma~3.9]{BGV}).

\begin{lemma}
\label{lem:Koszul}
Let $h_1,h_2,h_3\in\C[x,y,z]_d$ with $d \ge 1$. Then
%
\begin{equation}
\label{eq:linear_relazion}
  xh_1-yh_2+zh_3 = 0
\end{equation}
%
if and only if there exist $m_1,m_2,m_3\in\C[x,y,z]_{d-1}$ such that
%
\begin{equation}
\label{eq:minors_lemma}
  h_1 = ym_3-zm_2 \,, \qquad
  h_2 = zm_1-xm_3 \,, \qquad
  h_3 = xm_2-ym_1 \,.
\end{equation}
%
\end{lemma}

\begin{proof}
If $h_1,h_2,h_3$ satisfy \Cref{eq:minors_lemma}, then it is immediate to check that they satisfy \Cref{eq:linear_relazion} as well. 
Conversely, assume that \Cref{eq:linear_relazion} holds and let $R = \C[x,y,z]$.
The Koszul complex in the ring~$R$ is an exact sequence of $R$-modules
%
\[
  0 \to R\xrightarrow{\alpha} R^{\oplus 3} \xrightarrow{\beta} R^{\oplus 3} \xrightarrow{\gamma} R \to R/(x,y,z) \to 0 \,,
\]
%
where the maps are $\alpha(p) = (p x, p y, p z)$, $\gamma (w_1,w_2,w_3) = w_1 x + w_2 y + w_3 z$ and $\beta$ is defined by the matrix
%
\[
  \left(
  \begin{array}{ccc}
    0 & -z & y\\
    z & 0 & -x\\
    -y & x & 0 \\
  \end{array}
  \right) \,.
\]
%
The syzygy $xh_1-yh_2+zh_3=0$ implies that $(h_1, -h_2, h_3)$ is in the kernel of~$\gamma$,
and since the Koszul complex is exact, the triple $(h_1,-h_2, h_3)$ lies in the image of~$\beta$.
It follows that there exist $m_1,m_2,m_3 \in \C[x,y,z]_{d-1}$ such that \Cref{eq:minors_lemma} holds.
\end{proof}

Now we are in the position to prove \Cref{p2}.

\begin{proof}[Proof of \Cref{p2}]
Let $g_1$, $g_2$ and~$g_3$ be the order~$2$ minors determining the eigenscheme of~$f$, and let $g$ be the greatest common factor.
By writing
%
\[
  g_i = g \, h_i, \quad i=1,2,3
\]
%
we have that the residual scheme is defined by the ideal
$(h_1,h_2,h_3)$. Moreover, the
linear
syzygy between the generators~$g_i$ gives rise to the syzygy:
%
\[
  z\, h_1 - y\, h_2 + x\, h_3 = 0 \,.
\]
%
By \Cref{lem:Koszul}, the triple $(h_1,h_2,h_3)$ is the triple of order two minors of a matrix
%
\[
  \begin{pmatrix}
    x & y & z \\
    m_1 & m_2 & m_3
  \end{pmatrix} \,.
\]
%
for suitable forms $m_i \in \C[x,y,z]_r$, where $r =2 - \deg g \ge 0$.
By the assumption that $g$ is the greatest common factor of the three minors $g_1$, $g_2$ and~$g_3$,
we have that the zero scheme $Z$ of $(h_1,h_2,h_3)$ is $0$-dimensional, and
by \Cref{thm:nonempty} its degree is~$3$ if $r=1$. Observe that $Z$, being intersection of three conics with no common component, is not contained in a line. 

Finally, if $r=0$, the triple $(h_1,h_2,h_3)$ corresponds to three linear forms belonging to a pencil, hence the zero locus is a point.
\end{proof}


\begin{es}
Consider the form
%
\[
  f(x, y, z) = x^2 (y - z) \,.
\]
%
In the language of \Cref{p2} and its proof, we have $l=x$,
%
\[
 h_1 = x^2-2y^2+2y z \,, \quad 
 h_2 = 2z^2-x^2-2y z \,, \quad 
 h_3 = 2yz-2z^2-xy \,.
\]
%
The two syzygies in degree~$3$ are:
%
\[
  z \, h_1 - y \, h_2 + x \, h_3 = 0, \quad x \, h_2 + 2(y+z) \, h_3 + x \, h_1 = 0.
\]
%
Finally, $Z = \{ (0:1:1),(2:1:-1),(-2:1:-1) \}$.
Observe that one point is on the singular line $x=0$.
\end{es}

\subsection{Eigenschemes containing a line}

In this section, we consider polynomials having a whole line in the eigenscheme. We shall prove that residually we then have, in general, other three eigenpoints, and that the condition of having a line of eigenpoints is equivalent to having four collinear eigenpoints.

Finally, we shall prove that such polynomials belong all to the locus $\sL \subset \p^9$ (see \Cref{def:locus_L}).

\begin{lemma}
\label{4ptiSuRetta2}
Suppose $P_1, P_2, P_3, P_4$ are four distinct points belonging to a line~$t$
and let $C$ be a cubic such that has $P_1, \dots, P_4$ among its eigenpoints.
Then all the points of~$t$ are eigenpoints of~$C$.
Moreover,
%
\begin{equation*}
  6 \leq \rk \,\Phi(P_1, P_2, P_3, P_4) \leq 7
\end{equation*}
%
and the rank is~$6$ if and only if $\sigma(P_1, P_2) = 0$, i.e.\ if
and only if $t$ is tangent to the isotropic conic. If the rank of
the above matrix is~$7$, then all the cubics which have the line~$t$
of eigenpoints are given by~$t^2 \ell$ where $\ell$ is any line of the plane.
\end{lemma}
\begin{proof}
%%% +file: fourCollinearPoints.sage
It is not restrictive to assume that one of the points (say~$P_1$) is
the point~$(1: 0: 0)$.
With this simplification, the computation of the order~$6$ and order~$7$
minors of $\phi(P_1, \dots, P_4)$ is not hard and
the first part of the thesis follows. If the rank of $\Phi(P_1, \dots, P_4)$
is $7$, then the cubics which have $P_1, \dots, P_4$ as eigenpoints
give a two-dimensional linear subspace~$W$ of~$\p^9$, but also all
the cubics defined by~$t^2 \ell$ have $P_1, \dots, P_4$ as eigenpoints and are
a two-dimensional linear space of~$\p^9$ contained in~$W$. Hence the two linear
spaces coincide.
\end{proof}

\begin{lemma}
\label{lemma: twoTangentsCiso} Let $r$ be a line of the plane of
equation $ax+by+cz=0$ and suppose $r$ intersects $\iso$ in two distinct points~$P_1$ and~$P_2$. Consider the cubic:
\begin{equation}
\label{2_lines_of_eigenpoints}
  C(r) = \bigl( r^2-3\left(a^2+b^2+c^2\right)\iso \bigr) \, r \,
\end{equation}
then, the two tangents to $\iso$ in $P_1$ and~$P_2$ are eigenpoints of 
$C(r)$.
\end{lemma}
\begin{proof}
%% file: lemma_twoTangentsCiso.sage 
A generic point on $\iso$ is of the form $(\lambda^2 + \mu^2, 
\iii\lambda^2 -\iii\mu^2, 2\iii\lambda \mu)$ for $(\lambda: \mu) \in \p^1$.
By using this description, we can write two points~$P_1$ and~$P_2$ on~$\iso$,
compute the line $r$ joining them and define~$C(r)$.
It is then easy to see that the two tangents to~$\iso$ in~$P_1$ and~$P_2$ are eigenpoints
for~$C(r)$.
\end{proof}

\begin{prop}
\label{prop:eigenline_non_tangent}
Let $t$ be a line in the plane.
%
\begin{itemize}
   \item 
   If $t$ is not tangent to the isotropic conic, a cubic~$C$ has $t$ among its eigenpoints if and 
only if it is of the form $C = t^2\ell$, where $\ell$ is any line of the plane.
  \item 
  If $t$ is tangent to the isotropic conic~$\iso$ (in a point~$P$),
  a cubic~$C$ has $t$ among its eigenpoints if and only if it is of the form
  %
  \begin{equation}
  \label{cubiche_con_retta_autop_tg}
    C = t^2 \ell+\lambda C(r_0),
  \end{equation}
  %
  where $\ell$ is any line of the plane, $\lambda \in \C$, 
  $r_0$ is any fixed line passing through $P$ and~$C(r_0)$ is
  defined by \Cref{2_lines_of_eigenpoints}.
\end{itemize}
%
\end{prop}
\begin{proof}
%% file: una_retta_tg_autop.sage
The first part of the proposition is contained in~\Cref{4ptiSuRetta2}.
Suppose $t$ is tangent to $\iso$. We assume that $t$ is the line
$x+\iii y$, hence tangent to $\iso$ in $P = (1: \iii: 0)$.
We impose to the points
of~$t$ to be eigenpoints of a generic cubic $a_0x^3+a_1x^2y+\cdots+a_9z^3$
and we get a linear system in $a_0, \dots, a_9$ whose solution gives that
the cubics can be expressed as a linear combinations of the following
four cubics:
%
\[
  h_1 = t^2x, \quad
  h_2 = t^2y, \quad
  h_3 = t^2z, \quad
  h_4 = C(r),
\]
%
where $r$ is a fixed line passing through~$P$ (if, for instance, 
we choose the line $z=0$ for $r$, 
then $h_4$ is $z(x^2 + y^2 + 2/3z^2)$. Hence the cubics are of the
form of \Cref{cubiche_con_retta_autop_tg}. Conversely, it is easy to verify
that all the cubics of~(\ref{cubiche_con_retta_autop_tg}) are
linear combinations of $h_1, \dots, h_4$, hence it is enough
to see that every linear combination of $h_1, \dots, h_4$ gives a
cubic which has the line~$t$ among its eigenpoints, and the this result follows from an explicit computation.
\end{proof}

\begin{prop}
Let $t$ be a line tangent to the isotropic conic in a point~$P$.
Then all the cubics of the family from \Cref{cubiche_con_retta_autop_tg} are singular in~$P$ and~$t$ is one of the two tangents to the cubic in~$P$.
Notice that, in this case, when $C$ has a node at~$P$, the other tangent to~$C$ at~$P$ is orthogonal to~$t$.
\end{prop}
\begin{proof}
%% file: una_retta_tg_autop.sage
As seen in \Cref{prop:eigenline_non_tangent}, up to the action of~$\SO_3(\C)$, all the
cubics are a linear combination of $H_1, \dots, H_4$ and a direct
computation gives the thesis.
The last statement follows from the fact that all lines passing through a point on~$\iso$ are orthogonal to the tangent to~$\iso$ at that point.
\end{proof}

NOTA: @@Le cubiche con P1 singolare e tangente in P1 data da tg1 sono
un sistema lineare di dimensione 5, le cubiche
di~(\ref{cubiche_con_retta_autop_tg}) sono di dimensione
4, Che altre proprieta' deve avere una cubica per rientrare
in~(\ref{cubiche_con_retta_autop_tg})?

%
\begin{prop}
\label{prop:limitCubics}
Any cubic that has a line~$t$ in the eigenscheme the limit of a family of cubics whose general member has a $0$-dimensional eigenscheme with an aligned triple.
\end{prop}
\begin{proof}
%% file cubic_limit.sage
If the line~$t$ is not tangent to $\iso$, we can assume it is the line $z=0$. We fix on it the three points
%
\[
  P_1= (1: 0: 0), \ P_2 = (0: 1: 0), \ P_3 = (1: 1: 0)
\]
%
and we consider the three dimensional linear system of 
cubics $\Lambda \bigl( \Phi(P_1, P_2, P_3) \bigr)$, whose elements are of the form
%
\[
  C = (ax + by + cz)z^2 + d(x^3+y^3), \quad a, b, c, d \in \C \,.
\]
%
If $d=0$, the cubic $C$ is of the form $t^2 \ell$, hence is the 
generic cubic that has the line $t$ in the eigenscheme.
If $d \neq 0$, the eigenpoints of~$C$ different from $P_1, P_2, P_3$ are
the common zeros of the following three polynomials:
%
\begin{align*}
  g_1 & = 3y^2d - 2xya - 2y^2b + z^2b - 3yzc \,,\\
  g_2 & = 3x^2d - 2x^2a + z^2a - 2xyb - 3xzc \,,\\
  g_3 & = 2x^3ya - 2x^2y^2a + y^2z^2a + 2x^2y^2b -
          2xy^3b - x^2z^2b + 3x^2yzc - 3xy^2zc \,.
\end{align*}
%
In general, the ideal $(g_1, g_2, g_3)$ has $4$ points in 
general position. \\
In case the line $t$ is tangent to the isotropic conic, we can 
assume it is the line $x+\iii y =0$. Here we fix the three
points
\[
P_1 = (1: \iii: 0), \ P_2 = (0: 0: 1), \ P_3 = (1: \iii: 1) \,.
\]
In this case the linear system of cubics 
$\Lambda \bigl(\Phi(P_1, P_2, P_3)\bigr)$ is four dimensional and 
is given by:
\[
C = (x+\iii y)^2(ax + by+cz)+
 d(x^2 + y^2 + 2/3z^2)z+e (x^3 -\iii y^3 + z^3), 
 \quad a, b, c, d, e \in \C
\]
If $e=0$ the cubic $C$ is of the form described by \Cref{cubiche_con_retta_autop_tg}, hence is the generic cubic which contains the line $t$ in the eigenscheme. If $e \not= 0$, as above, 
the remaining eigenpoints of~$C$ are given by an ideal generated 
by the polynomials:
%
\begin{align*}
h_1 & = xza + \iii yza + \iii xzb - yzb - x^2c -2\iii xyc + y^2c - x^2d - y^2d + 3xze -3\iii yze - 3z^2e \,, \\
h_2 & = 2x^2a + 3\iii xya - y^2a -\iii x^2b + 3xyb + 2\iii y^2b + 2xzc + 2\iii yzc +3\iii xye \,, \\
h_3 & = 3x^2zb + 6\iii xyzb - 3y^2zb + 2\iii x^3c - 5x^2yc -4\iii xy^2c + y^3c + 2\iii xz^2c - 2yz^2c \\
&\phantom{=} + 2\iii x^3d - x^2yd + 2\iii xy^2d - y^3d -6\iii x^2ze - 6xyze -3\iii y^2ze + 6\iii xz^2e - 3yz^2e \,, \\
h_4 & = 3y^2za -\iii yz^2a + 6xyzb + 9\iii y^2zb -3\iii xz^2b + 4yz^2b + 4\iii x^2yc - 9xy^2c -5\iii y^3c \\
&\phantom{=} + 2x^2zc + 4\iii xyzc - 2y^2zc + 4\iii yz^2c + 2z^3c + 4\iii x^2yd - xy^2d + 5\iii y^3d + 2x^2zd \\
&\phantom{=} + 2y^2zd -12\iii xyze - 15y^2ze - 6xz^2e + 21\iii yz^2e + 6z^3e \,. \qedhere
\end{align*}
%
\end{proof}

\subsection{Eigenschemes containing a conic}
\Cref{lemma: twoTangentsCiso} can be extended, indeed we have:
%
\begin{prop}
\label{prop:twoLines}
A cubic $C$ of the plane has two distinct lines $t_1$ and~$t_2$ in its eigenscheme if and only if $t_1$ and~$t_2$ are tangent to the isotropic conic (in two points~$P_1$ and~$P_2$) and $C= C(r)$ as in \Cref{2_lines_of_eigenpoints}, where $r = P_1 \vee P_2$.
\end{prop}
%
\begin{proof}
Suppose $t_1$ and~$t_2$ are two lines which are contained in the eigenschemes of~$C$. From \Cref{prop:eigenline_non_tangent} we have that $t_1$ and~$t_2$ are tangent to $\iso$ and \Cref{lemma: twoTangentsCiso} gives that $C$ is given by \Cref{2_lines_of_eigenpoints}. If, conversely, $C$ is given by \Cref{2_lines_of_eigenpoints}, where $r$ is a line, the result again follows from \Cref{lemma: twoTangentsCiso}.
\end{proof}

\begin{lemma}
\label{lemma:fourOnIso}
Suppose $P_1, P_2, P_3, P_4$ are four distinct points on $\iso$ and let
$C$ be a cubic that has $P_1, \dots, P_4$ among its eigenpoints.
Then all the points of~$\iso$ are eigenpoints of~$C$ 
and~$C$ splits into the union of~$\iso$ and a line~$r$.
In particular, if four distinct points of~$\iso$ are eigenpoints of a cubic,
then all the points of~$\iso$ are eigenpoints of the cubic.
\end{lemma}
\begin{proof}
%% file eigenpointsConic.sage
We can assume that $P_1 = (1: \iii: 0)$. Hence all the other points
of~$\iso$ are given by
%
\[
  P(w) = \bigl(\iii(w^2 + 1): 1 - w^2: -2w\bigr) \text{ with } w \in \C \,,
\]
%
so $P_2 = P(w_1)$, $P_3 = P(w_2)$, and $P_4 = P(w_3)$, where the $w_i$'s are
all distinct. The cubics which have $P_1, \dots, P_4$ among the eigenpoints
are the linear system given by~$\Lambda (M)$, where $M = \Phi(P_1, P_2, P_3, P_4)$.
It can be verified that the rank of~$M$ is $7$, hence $\Lambda(M)$ is a linear variety of~$\p^9$ of dimension~$2$, but all the cubics which split into~$\iso$ and a line
of the plane do have all the points of~$\iso$ among the eigenpoints and
these cubics give a linear variety of dimension~$2$ in $\p^9$
which therefore coincides with~$\Lambda(M)$.
\end{proof}

\begin{lemma}
\label{lemma3ptiSuCiso}
It is not possible to have a cubic $C$ which has, among its eigenpoints,
an irreducible conic $\Gamma$ tangent in a point~$P_1$ to the isotropic
conic and intersecting $\iso$ in two other distinct points.
\end{lemma}
\begin{proof}
%% file caso3ptiIso_2.sage+\\
As in the proof~\Cref{lemma:fourOnIso}, we can assume that
$P_1 = (1: \iii: 0)$
and~$P_2$ and~$P_3$ are $P(w_1)$ and~$P(w_2)$. Then we construct
the pencil of conics passing through $P_1$, $P_2$ and~$P_3$ and
tangent to~$\iso$ in~$P_1$ and we construct two distinct points~$Q_1$ and~$Q_2$ on a conic of this family. If there is a conic~$\Gamma$
tangent to~$\iso$ in~$P_1$ and passing through~$P_2$ and~$P_3$,
then, for suitable values of the parameters defining the points
$P_1, P_2, P_3$ and~$Q_1$ and~$Q_2$, the matrix
$M = \Phi(P_1, P_2, P_3, Q_1, Q_2)$
must have rank~$9$ or smaller. Again, 
some computations give that $M$ cannot 
have rank~$8$ or smaller and if $M$ has rank $9$, the corresponding cubic 
does not have a conic among its eigenpoints.
\end{proof}

\begin{lemma}
\label{lemma:bitangentToCiso}
If a cubic~$C$ has, among its eigenpoints, an irreducible conic $\Gamma$
which is tangent to $\iso$ in two distinct points~$P_1$ and~$P_2$, then
the equation of~$C$ is of the form $(\lambda \iso + \mu r^2)r$ where
$r$ is the equation of the line~$P_1\vee P_2$ and $\lambda, \mu$ are parameters
in~$\C$.
\end{lemma}
\begin{proof}
%% file: caso2ptiIso_partI.sage, 
%% caso2ptiIso_partII.sage.
The proof is quite similar to the one of~\Cref{lemma3ptiSuCiso}. Also
here we can assume that $P_1 = (1: \iii: 0)$ and we construct the pencil of
bitangent conics in $P_1$ and~$P_2$ to $\iso$. Then we construct three
points~$Q_1$, $Q_2$, and~$Q_3$ on a conic of the pencil
and we study the matrix $M = \Phi(P_1, P_2, Q_1, Q_2, Q_3)$. We can see
that the rank of this matrix is $9$ (for each admissible value of the
involved parameters), hence $M$ identifies a cubic~$C$. The computation
of~$C$ shows that its equation is of the above form.
\end{proof}

\begin{prop}
All the cubics $C$ defined by the polynomial
%
\begin{equation}
\label{cub_conica_eig}
  (\lambda \iso + \mu r^2) \, r \quad \mbox{$\lambda, \mu \in \C$}
\end{equation}
%
where $r=ax+by+cz$ is a line of the plane,
have the eigenpoints given by the conic
$\lambda \iso+3\mu r^2$ (bitangent to~$\iso$ in the points~$\iso \cap r$)
and the point~$(a:b:c)$ (the pole of~$r$
w.r.t.~$\iso$). Conversely, if
$C$ is a cubic of the plane which has a conic among its eigenpoints,
then there exists a line~$r$ in the plane and $\lambda, \mu \in \C$
such that the equation of~$C$ is given by \Cref{cub_conica_eig}.
\end{prop}
\begin{proof}
%% file: caso2ptiIso_partI.sage+ (parte finale del file)
If the equation of~$C$ is of the form described above, a direct computation
shows that the eigenpoints of~$C$ are the pole of~$r$ with respect to $\iso$
and the conic $\lambda \iso+3\mu r^2$. Suppose, conversely, that $C$ has
a conic~$\Gamma$ among its eigenpoints. 
First, we assume that $\Gamma$ is irreducible.
If $\Gamma$ is the isotropic conic,
then \Cref{lemma:fourOnIso} gives that $C$ splits into~$\iso$ and a line.
If $\Gamma$ is not the isotropic conic, as a consequence
of~\Cref{lemma:fourOnIso} and~\Cref{lemma3ptiSuCiso} it must be bitangent,
in two points~$P_1$ and~$P_2$, to~$\iso$. 
Then the statement is proven from~\Cref{lemma:bitangentToCiso}. 
If $\Gamma$ is reducible,
from~\Cref{prop:twoLines} we have that $\Gamma$ splits
into two tangent lines to~$\iso$.
\end{proof}

\begin{rmk}
If a line $r=ax+by+cz$ is tangent to $\iso$ in a point~$P$, which hence must be~$(a:b:c)$, 
the conic $\lambda \iso +3 \mu r^2$ is tangent to $\iso$ in $P$.
\end{rmk}
Finally, with similar techniques, we can prove the following results:
\begin{lemma}
    It is not possible to have a cubic $C$ which has, among its eigenpoints, an irreducible conic $\Gamma$ osculating in a point $P_1$ the isotropic conic and intersecting $\iso$ in another point.
    %% file: OsculatConic.sage
\end{lemma}
\begin{lemma}
    If a cubic $C$ has, among its eigenpoints, an irreducible conic $\Gamma$ which is iperosculating $\iso$ in a point $P_1$, then the equation of $C$ is of the form 
    $r(\iso - r^2)$, where $r$ is the tangent line to $\iso$ in $P_1$ and the equation of $\Gamma$ is $\iso-3r^2$.
    %% file: IperOsc.sage
\end{lemma}
Moreover:
\begin{prop}
    All the cubics $C$ defined by the polynomial
\begin{equation}
\label{cub_osc}
  r(\iso -r^2)
\end{equation}
where $r$ is the tangent line to $\iso$ in a point $P_1$ have the eigenpoints given by the conic $\iso-3r^2$ which is iperosculating $\iso$ in $P_1$ and the point $P_1^2$ (embedded component). Conversely, if $C$ is a cubic of the plane which has, among its eigenpoints, a conic iperosculating $\iso$ in $P_1$, then $C$ is given by $r(\iso-r^2)$, where $r$ is the tangent line to $\iso$ in $P_1$.
%
\end{prop}

Summing up, we obtain:

\begin{corollary}
      The eigenscheme of a cubic~$C$ contains a conic~$\Gamma$ if and only if either $\Gamma = \iso$ or $\Gamma$ is bitangent to~$\iso$ or $\Gamma$ is iperosculating $\iso$.
\end{corollary}


\section{Possible configurations of the seven eigenpoints}
\label{further_alignments}

\begin{table}[ht]
\caption{All possible configurations of~$7$ points with alignments that can appear as eigenschemes of a ternary cubic form, provided that the eigenscheme is $0$-dimensional and reduced.}
\centering
\begin{tabular}{|clc|}\hline
  n. lines & collinear vertices & config.\\ \hline
 1& (1, 2, 3) &  $(C_1)$\\
 2& (1, 2, 3), (1, 4, 5) &  $(C_2)$\\
 3& (1, 2, 3), (1, 4, 5), (1, 6, 7) & $(C_3)$\\
 3& (1, 2, 3), (1, 4, 5), (2, 4, 6) & $(C_4)$\\
 4& (1, 2, 3), (1, 4, 5), (1, 6, 7), (2, 4, 6) & $(C_5)$\\
 4& (1, 2, 3), (1, 4, 5), (2, 4, 6), (3, 5, 6) & $(C_6)$\\
 5& (1, 2, 3), (1, 4, 5), (1, 6, 7), (2, 4, 6), (2, 5, 7)& $(C_7)$\\
 6& (1, 2, 3), (1, 4, 5), (1, 6, 7), (2, 4, 6), (2, 5, 7), (3, 4, 7)& $(C_8)$\\
 7& (1, 2, 3), (1, 4, 5), (1, 6, 7), (2, 4, 6), (2, 5, 7), (3, 4, 7), (3, 5, 6) &  $(C_9)$\\ \hline
\end{tabular}
\label{table:all_alignments}
\end{table}
%
\begin{figure}[ht]
  \centering
  \begin{tikzpicture}
    \begin{scope}[scale=0.8]
    \begin{scope}
      \node[point, label={[label distance = -4pt, below]$P_1$}] (P1) at (0,0) {};
      \node[point, label={[label distance = -4pt, below]$P_3$}] (P3) at (3.6,0) {};
      \node[point, label={[label distance = -4pt, below]$P_2$}] (P2) at ($(P1)!0.5!(P3)$) {};
      \node[point, label={[label distance = -4pt, left]$P_4$}] (P4) at (-0.5, 0.9) {};
      \node[point, label={[label distance = -4pt, left]$P_5$}] (P5) at (0.7, 2.3) {};
      \node[point, label={[label distance = -4pt, right]$P_6$}] (P6) at (1.8, 2.1) {};
      \node[point, label={[label distance = -4pt, right]$P_7$}] (P7) at (2.9, 1.2) {};
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P2); 
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P2) -- (P3);
    \end{scope}
    \begin{scope}[xshift=6cm]
      \node[point, label={[label distance = 0pt]210:$P_1$}] (P1) at (0,0) {};
      \node[point, label={[label distance = -4pt, below]$P_3$}] (P3) at (3.6,0) {};
      \node[point, label={[label distance = -4pt, below]$P_2$}] (P2) at ($(P1)!0.5!(P3)$) {};
      \node[point, label={[label distance = 0pt, left]$P_5$}] (P5) at (1.5, 2.4) {};
      \node[point, label={[label distance = 0pt, left]$P_4$}] (P4) at ($(P1)!0.3!(P5)$) {};
      \node[point, label={[label distance = 0pt, right]$P_6$}] (P6) at (2.2, 2.1) {};
      \node[point, label={[label distance = 0pt, right]$P_7$}] (P7) at (3.2, 1.2) {};
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P2);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P2) -- (P3);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P4);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P5);
    \end{scope}
    \begin{scope}[xshift=12cm]
      \node[point, label={[label distance = 0pt]210:$P_1$}] (P1) at (0,0) {};
      \node[point, label={[label distance = -4pt, below]$P_3$}] (P3) at (3.6,0) {};
      \node[point, label={[label distance = -4pt, below]$P_2$}] (P2) at ($(P1)!0.5!(P3)$) {};
      \node[point, label={[label distance = 0pt, left]$P_5$}] (P5) at (1.5, 2.4) {};
      \node[point, label={[label distance = 0pt, left]$P_4$}] (P4) at ($(P1)!0.3!(P5)$) {};
      \node[point, label={[label distance = 0pt, above]$P_6$}] (P6) at (1.6, 1) {};
      \node[point, label={[label distance = 0pt, above]$P_7$}] (P7) at (3.2, 2) {};
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P2);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P2) -- (P3);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P4);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P6);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P6) -- (P7);
    \end{scope}
    \begin{scope}[yshift=-6cm]
      \node[point, label={[label distance = 0pt]210:$P_1$}] (P1) at (0,0) {};
      \node[point, label={[label distance = -4pt, below]$P_3$}] (P3) at (3.6,0) {};
      \node[point, label={[label distance = -4pt, below]$P_2$}] (P2) at ($(P1)!0.5!(P3)$) {};
      \node[point, label={[label distance = 4pt]90:$P_5$}] (P5) at (1.5, 2.4) {};
      \node[point, label={[label distance = 0pt, left]$P_4$}] (P4) at ($(P1)!0.3!(P5)$) {};
      \node[point, label={[label distance = 0pt, right]$P_6$}] (P6) at ($(P5)!0.25!(P3)$) {};
      \node[point, label={[label distance = 0pt, left]$P_7$}] (P7) at (2, 0.9) {};
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P2);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P2) -- (P3);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P4);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P3) -- (P6);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P6) -- (P5);
    \end{scope}
    \begin{scope}[xshift=6cm, yshift=-6cm]
      \node[point, label={[label distance = 0pt]210:$P_1$}] (P1) at (0,0) {};
      \node[point, label={[label distance = -4pt, below]$P_3$}] (P3) at (3.6,0) {};
      \node[point, label={[label distance = -4pt, below]$P_2$}] (P2) at ($(P1)!0.5!(P3)$) {};
      \node[point, label={[label distance = 4pt]90:$P_5$}] (P5) at (1.5, 2.4) {};
      \node[point, label={[label distance = 2pt]100:$P_4$}] (P4) at ($(P1)!0.3!(P5)$) {};
      \node[point, label={[label distance = 0pt, right]$P_6$}] (P6) at ($(P5)!0.25!(P3)$) {};
      \node[point, label={[label distance = 0pt, above]$P_7$}] (P7) at ($(P4)!0.6!(P3)$) {};
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P2);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P2) -- (P3);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P4);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P3) -- (P6);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P6) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P3) -- (P7);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P7) -- (P4);
    \end{scope}
    \begin{scope}[xshift=12cm, yshift=-6cm]
      \node[point, label={[label distance = 0pt]210:$P_1$}] (P1) at (0,0) {};
      \node[point, label={[label distance = 0pt, above]$P_3$}] (P3) at (3.6,0) {};
      \node[point, label={[label distance = -4pt, below]$P_2$}] (P2) at (2.2,0) {};
      \node[point, label={[label distance = 4pt]90:$P_5$}] (P5) at (1.5, 2.4) {};
      \node[point, label={[label distance = 2pt]100:$P_4$}] (P4) at ($(P1)!0.3!(P5)$) {};
      \tkzInterLL(P3,P5)(P2,P4) \tkzGetPoint{P6}
      \node[point, label={[label distance = 0pt, right]$P_6$}] at (P6) {};
      \node[point, label={[label distance = 0pt, above]$P_7$}] (P7) at (3.2, 1.8) {};
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P2);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P2) -- (P3);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P4);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P3) -- (P6);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P6) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P2);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P2) -- (P6);
    \end{scope}
    \begin{scope}[yshift=-12cm]
      \node[point, label={[label distance = 0pt]210:$P_1$}] (P1) at (0,0) {};
      \node[point, label={[label distance = -4pt, below]$P_3$}] (P3) at (3,0) {};
      \node[point, label={[label distance = 0pt]195:$P_2$}] (P2) at ($(P1)!0.5!(P3)$) {};
      \node[point, label={[label distance = 4pt]90:$P_5$}] (P5) at (1.5, 2.6) {};
      \node[point, label={[label distance = 2pt]100:$P_4$}] (P4) at ($(P1)!0.5!(P5)$) {};
      \node[point, label={[label distance = 0pt, right]$P_6$}] (P6) at ($(P3)!0.25!(P5)$) {};
      \tkzInterLL(P3,P4)(P2,P5) \tkzGetPoint{P7}
      \node[point, label={[label distance = 0pt, right]$P_7$}] at (P7) {};
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P2);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P2) -- (P3);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P4);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P3) -- (P6);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P6) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P7);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P7) -- (P3);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P5) -- (P7);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P7) -- (P2);
    \end{scope}
    \begin{scope}[xshift=6cm, yshift=-12cm]
      \node[point, label={[label distance = 0pt]210:$P_1$}] (P1) at (0,0) {};
      \node[point, label={[label distance = -4pt, below]$P_3$}] (P3) at (3,0) {};
      \node[point, label={[label distance = 0pt]195:$P_2$}] (P2) at ($(P1)!0.5!(P3)$) {};
      \node[point, label={[label distance = 4pt]90:$P_5$}] (P5) at (1.5, 2.6) {};
      \node[point, label={[label distance = 2pt]100:$P_4$}] (P4) at ($(P1)!0.5!(P5)$) {};
      \node[point, label={[label distance = 2pt]80:$P_6$}] (P6) at ($(P3)!0.5!(P5)$) {};
      \tkzInterLL(P3,P4)(P2,P5) \tkzGetPoint{P7}
      \node[point, label={[label distance = 0pt]0:$P_7$}] at (P7) {};
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P2);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P2) -- (P3);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P4);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P3) -- (P6);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P6) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P7);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P7) -- (P3);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P5) -- (P7);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P7) -- (P2);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P7);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P7) -- (P6);
    \end{scope}
    \begin{scope}[xshift=12cm, yshift=-12cm]
      \node[point, label={[label distance = 0pt]210:$P_1$}] (P1) at (0,0) {};
      \node[point, label={[label distance = -4pt, below]$P_3$}] (P3) at (3,0) {};
      \node[point, label={[label distance = 0pt]195:$P_2$}] (P2) at ($(P1)!0.5!(P3)$) {};
      \node[point, label={[label distance = 4pt]90:$P_5$}] (P5) at (1.5, 2.6) {};
      \node[point, label={[label distance = 2pt]100:$P_4$}] (P4) at ($(P1)!0.5!(P5)$) {};
      \node[point, label={[label distance = 2pt]80:$P_6$}] (P6) at ($(P3)!0.5!(P5)$) {};
      \tkzInterLL(P3,P4)(P2,P5) \tkzGetPoint{P7}
      \node[point, label={[label distance = 0pt]0:$P_7$}] at (P7) {};
      \node[draw, line] at (P7) [circle through={(P2)}] {};
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P2);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P2) -- (P3);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P4);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P3) -- (P6);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P6) -- (P5);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P4) -- (P7);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P7) -- (P3);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P5) -- (P7);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P7) -- (P2);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P1) -- (P7);
      \draw[line, shorten <= -0.5cm, shorten >= -0.5cm] (P7) -- (P6);
    \end{scope}
    \end{scope}
  \end{tikzpicture}
  \caption{Graphical representations of the $9$ cases of possible alignments described in \Cref{table:all_alignments}.}
  \label{figure:all_alignments}
\end{figure}

\Cref{table:all_alignments} lists all the possible configuration
of 7 distinct points of the plane, according to their collinearities. Here
we want to see which of those nine configurations can be realized by
the seven eigenpoints of a cubic and, in case the configuration is
realizable, which are the cubics with that configuration of eigenpoints.
We shall say that the eigenpoints of a cubic curve are in a $(C_i)$ 
configuration if they are aligned according to~$(C_i)$; we shall say 
that the eigenpoints are in a \emph{strict} $(C_i)$ configuration if, 
in addition, there are no other alignments among the points.

First of all, it is well known that configuration~$(C_9)$ cannot be realized
by seven points of the plane over a field of zero
characteristic (see \cite{Whitney1935}), therefore we do not consider
it in our analysis.

\underline{Configuration $(C_1)$} can be realized. \Cref{proposition:three_aligned_ranks} and \Cref{locus_one_alignment}
give a description of the cubics with such a configuration of points:
we fix two points~$P_1$ and~$P_2$ in~$\p^2$, we take a generic~$P_3$
on the line~$P_1 \vee P_2$, all the cubics with
$P_1, P_2, P_3$ eigenpoints is given by $\Lambda \bigl( \Phi(P_1, P_2, P_3) \bigr)$, the three
dimensional linear subspace of~$\p^9$ (four dimensional, if the
line~$P_1 \vee P_2$ is tangent in~$P_1$, $P_2$, or~$P_3$ to~$\iso$). For generic choices of the points, the cubics have the eigenpoints in a strict 
$(C_1)$ configuration.

\underline{Configuration $(C_2)$} means that the points
$P_1, \dots, P_5$ are in a
$V$-configuration, hence $\rk \, \Phi(P_1, \dotsc, P_5)$
must be~$9$ or~$8$. If the rank is~$9$, then $\delta_1(P_1, P_2, P_4) = 0$
or $\delta_2(P_1, \dotsc, P_5) = 0$, see \Cref{theorem:rank_V}.
From \Cref{proposition:third_alignment}, the only case
to have a strict $(C_2)$ configuration 
is $\delta_1(P_1, P_2, P_4) = 0$. Hence, if we fix two points~$P_1$ and~$P_2$
in the plane in an arbitrary way, we choose $P_4$ such that
$\scl{P_4}{s_{11}\, P_2 - s_{12} \, P_1}=0$, see \Cref{rmk_delta_case1}; then we chose any~$P_3$
on the line~$P_1 \vee P_2$ and any~$P_5$ on the line~$P_1 \vee P_4$, we have a cubic with a configuration of type~$(C_2)$ (strict, in the general case). If the rank is~$8$, configuration~$(C_2)$ can only be obtained from \Cref{rk8_1} (and hence are sub-cases of the case $\delta_1(P_1, P_2, P_4)=0$), (see~\Cref{rmk:particular_cases}).

\underline{Configuration $(C_3)$}. Suppose we have $7$ eigenpoints such that
$(P_1, P_2, P_3)$, $(P_1, P_4, P_5)$ and $(P_1, P_6, P_7)$ are
aligned. Hence
$V_1 = (P_1, P_2, P_3, P_4, P_5)$, $V_2 = (P_1, P_2, P_3, P_6, P_7)$ and
$V_3 = (P_1, P_4, P_5, P_6, P_7)$, are three $V$-configurations. It holds:

\begin{lemma}
\label{no_delta1_delta1} Suppose we have seven eigenpoints $P_1, \dots, P_7$
of a cubic in configuration $(C_3)$. Then among the $7$ points there is a
$V$-configuration that satisfies a $\delta_2$ condition.
\end{lemma}
\begin{proof}
%% la dim si trova in config3.sage
The points $P_1, P_2, P_3, P_4, P_5$ are in a $V$-configuration.
If $\rk \, \Phi(P_1, \dots, P_5) = 8$, the result follows from~\Cref{rank_8}.
Therefore, assume that the matrix $\Phi(P_1, \dots, P_5)$ has rank~$9$.
If $\delta_2(P_1, \dots, P_5) = 0$, the statement follows;
otherwise, $\delta_1(P_1, P_2, P_4) = 0$.
Then consider the $V$-configuration $P_1$, $P_2$, $P_3$, $P_6$, $P_7$.
As above, we suppose $\delta_1(P_1, P_2, P_6) = 0$.
The two equations obtained from the two $\delta_1$ give two
linear equations in the coordinates of~$P_2$.
Assuming the matrix of the system of linear equations have
maximal rank, the solution gives a point~$P_2$ that coincide, as
a projective point, to $P_1$, which is impossible.
Hence we are forced to consider the
case in which the matrix of the linear system does not have maximal rank.
This condition implies that $\scl{P_1}{P_1} = 0$, hence $P_1$ is on the
isotropic conic. W.l.o.g.\ we can assume $P_1 = (1: \iii: 0)$ and again
we can determine~$P_2$ such $\delta_1(P_1, P_2, P_4)$ and
$\delta_1(P_1, P_2, P_6)$ are zero. The matrix $\Phi(P_1, P_4, P_5, P_6, P_7)$
must have rank~$9$ or smaller, then either
$\delta_2(P_1, P_4, P_5, P_6, P_7)=0$ or $\delta_1(P_1, P_4, P_6) = 0$. In
the first case, we have a $\delta_2$ condition among the points, hence
we assume $\delta_1(P_1, P_4, P_6) = 0$. Solving this equation and
considering the corresponding points, we get that either
$\delta_2(P_1, P_2, P_3, P_6, P_7) = 0$ or
$\delta_2(P_1, P_2, P_3, P_4, P_5) = 0$.
\end{proof}

As a consequence of \Cref{no_delta1_delta1}, we have that configuration $(C_3)$ can
be obtained only if we have $5$ points such that
$\delta_2(P_1, \dotsc, P_5) = 0$. \Cref{prop:definitionP3} describes how to
obtain five points which give a configuration $(C_3)$ of eigenpoints (which, 
in general, is strict).

\underline{Configuration $(C_4)$}. It holds:
\begin{prop}
\label{conf4no} If the eigenpoints of a cubic are in a $(C_4)$ configuration, then they actually are in a $(C_8)$ configuration.
\end{prop}
\begin{proof}
%% il file config4.sage contiene la dimostrazione
From the results of \Cref{rank_8}, we know that $(C_4)$
cannot be obtained from a rank 8 $V$-configuration,
hence it remains to consider the case $\delta_1(P_1, P_2, P_4) = 0$,
$\delta_1(P_2, P_1, P_4) = 0$ and $\delta_1(P_4, P_1, P_2) = 0$,
which implies
%
\[
  \delta_2(P_1, P_2, P_3, P_4, P_5) = 0 \,, \quad
  \delta_2(P_2, P_1, P_3, P_4, P_6) = 0 \,, \quad
  \delta_2(P_4, P_1, P_5, P_2, P_6) = 0 \,. \qedhere
\]
%
\end{proof}

\underline{Configuration $(C_5)$}
%% file: config5.sage
\begin{prop}
\label{proposition:condition_5}
%\textbf{(AGGIUNGERE IPOTESI SU RANGO NON 8)}
Suppose we have $7$ points which are eigenpoints of a cubic and are in a strict~$(C_5)$ configuration.
Then it holds
%
\[
  P_1 = P_2 \times P_4 
\]
%
and among the points $P_1, \dots, P_6$, where $P_6$ is collinear with~$P_2$ and~$P_4$, we have the relation
%
\begin{equation}
\label{cndC5}
  s_{26}(s_{45}s_{13}-s_{34}s_{15})+s_{46}(s_{25}s_{13}-s_{23}s_{15}) = 0 \,.
\end{equation}
%
Moreover, the points~$P_6$ and~$P_7$ are determined as follows:
%
\begin{equation}
\label{p6formula}
\begin{multlined}
  P_6 = (s_{15}s_{24}s_{34}+s_{15}s_{23}s_{44} -s_{13}s_{25}s_{44} -s_{13}s_{24}s_{45}) \, P_2 \\ + (s_{13}s_{24}s_{25}-2s_{15}s_{22}s_{34}+s_{13}s_{22}s_{45}) \, P_4 \,, 
\end{multlined}
\end{equation}
%
\begin{equation}
\label{p7formula}
\begin{multlined}
  P_7 = (s_{16}(s_{26}s_{45}+s_{24}s_{56})-s_{26}s_{15}s_{46}-s_{24}s_{15}s_{66}) \, P_1 \\
  - (s_{11}(s_{26}s_{45}+s_{24}s_{56})-s_{26}s_{15}s_{14}-s_{24}s_{15}s_{16}) \, P_6 \,.
\end{multlined}
\end{equation}
%
\end{prop}
\begin{proof}
We can define the points as follows: $P_1$, $P_2$ and~$P_4$ with generic
coordinates, meanwhile $P_3 = u_1 \, P_1 + u_2 \, P_2$, $P_5 = v_1 \, P_1 + v_2 \, P_4$ and
$P_6 = w_1 \, P_2 + w_2 \, P_4$. The points $(P_2, P_1, P_3, P_4, P_6)$ are in a
$V$-configuration and we may assume that $\rk \, \Phi(P_2, P_1, P_3, P_4, P_6) = 9$ (the case in which the rank is~$8$ has been discussed in~\Cref{rank_8} and none of the 
configurations obtained there is compatible with this configuration), hence necessarily $\delta_1(P_2, P_1, P_4)=0$ because we suppose that $P_2$, $P_5$, and~$P_7$ are not aligned.
Analogously, $\delta_1(P_4, P_1, P_2) = 0$ and $\delta_1(P_6, P_1, P_2) = 0$. Moreover, by \Cref{no_delta1_delta1} and by possibly relabelling the points, we may assume that $\delta_2(P_1, P_2, P_3, P_4, P_5)=0$. The saturation of the ideal generated by the four polynomials
above by the condition that $P_1$, $P_2$, and~$P_4$ are not aligned equals the ideal generated by
$s_{12}, s_{14}$, which also contains~$s_{16}$. This shows that
$s_{12} = s_{14}=s_{16}=0$. In particular, we see that $P_1 = P_2 \times P_4$.

To prove \Cref{cndC5}, we take $P_2$ and~$P_4$ generic,
$P_1 = P_2 \times P_4$ and~$P_3$, $P_5$, $P_6$ generic points on the lines
$P_1 \vee P_2$, $P_1 \vee P_4$, $P_2 \vee P_4$, respectively. 
The matrix $M = \Phi(P_1, \dots, P_6)$ must have rank $9$ (or smaller), 
hence, in order to see when $P_6$ is an eigenpoints,
we have to see when all the order $10$-minors are zero
Their computation takes too much time, so we need another strategy:
we select only one non zero order~$10$ minor of~$M$; 
it is a polynomial which factorizes into several factors 
(many of them consequence of the specific minor chosen) 
and a factor which involves the variables used in the definition
of the points (call it $f$). 
We claim that the factor~$f$ is zero if and only if the rank of~$M$ is smaller than $10$.
In order to see this, we specialize the point~$P_1$
into the point~$(1:0:0)$. In this situation the order $10$ minors of~$M$
(specialized on this point~$P_1$) are easy to compute and we can
verify they are zero if and only if the polynomial $f$, specialized
on the point~$(1:0:0)$, is zero. 
To complete the proof of the claim, 
we should also take for $P_1$ the point~$(1:\iii:0)$;
however, this is not necessary, since the configuration of the points implies that $\scl{P_1}{P_1}=0$ and
this is not possible. 
A manipulation of the polynomial~$f$ shows that it is equal to
%
\[
  \bigl((s_{26}(s_{45}s_{13}-s_{34}s_{15})+s_{46}(s_{25}s_{13}-s_{23}s_{15})\bigr)/s_{11} \,,
\]
%
hence the points $P_1, \dots, P_6$ are eigenpoints if and only if \Cref{cndC5} is satisfied.

If in \Cref{cndC5} in place of~$P_6$ we substitute $w_1 \, P_2 + w_2 \, P_4$,
we find~$w_1$ and~$w_2$, which give \Cref{p6formula}.

Finally, to prove \Cref{p7formula}, we change~$P_3$ with~$P_7$ in \Cref{cndC5} and we take into account that $P_7$ is collinear with~$P_1$ and~$P_6$.
\end{proof}
%
The converse of the above proposition is also true, more precisely:
\begin{prop}
   Suppose $P_1, P_2, P_3, P_4, P_5$ are in a $V$-configuration such that
%
\[
  P_1 = P_2 \times P_4 \,,
\]
%
%%%%and the condition matrix satisfies \rk \, \Phi (P_1, \dots, P_5) = 9$, 
and suppose there is a cubic that has $7$ eigenpoints, five of which are $P_1, \dotsc, P_5$. Then the points~$P_6$ and~$P_7$, given by \Cref{p6formula} and \Cref{p7formula}, are the remaining eigenpoints of the cubic, hence $(P_2, P_4, P_6)$ are aligned, so the points are in a $(C_5)$ configuration. Furthermore, in the general case, the configuration is 
strict but, in the case of $\rk \, \Phi(P_1, \dotsc, P_5) = 9$, there are sub-cases in which $(P_2, P_5, P_7)$ or $(P_3, P_4, P_7)$ or $(P_3, P_5, P_7)$ are aligned. In all these cases the points are in a $(C_8)$ configuration.
\end{prop}
%
\begin{proof}
First we suppose that $\rk \, \Phi(P_1, \dotsc, P_5) = 9$.
We fix generic coordinates for the points.
Since $P_1 = P_2 \times P_4$ implies $s_{12}=s_{14}=0$,
we have that $\delta_2 (P_1,P_2,P_3,P_4,P_5)=0$. 
It follows from \Cref{proposition:third_alignment} that the corresponding unique cubic having such $5$ points as eigenpoints has $P_6$ and~$P_7$ aligned with~$P_1$.
Suppose now that $P_6$ is given by \Cref{p6formula}.
In order to see if it is an eigenpoint, it is enough to select nine linearly independent rows from $\Phi(P_1, \dotsc, P_5)$
and to compute the determinant of the three order $10$ matrices whose rows are the nine rows above plus $\phi_1(P_6)$, $\phi_2(P_6)$, and~$\phi_3(P_6)$ respectively.
Since we obtain that these three determinants are zero, we get that $\rk \, \Phi(P_1, \dotsc, P_6) = 9$.
A similar argument shows that $P_7$ is given by \Cref{p7formula}.
A random example shows that, in general, there are no further collinearities among the points.
If we impose that $(P_2, P_5, P_7)$ are aligned, we get that this condition splits into two cases:
one imposes the collinearity of $(P_3, P_4, P_7)$, the other imposes that the collinearity of $(P_3, P_5, P_6)$,
hence we get two $(C_8)$ configurations. Similarly for the other two cases. 

If $\rk \, \Phi(P_1, \dotsc, P_5) = 8$ and $P_1 = P_2 \times P_4$, we have that $s_{12} = s_{14} = 0$, so the $V$-configuration cannot satisfies \Cref{rk8_1}, since we would have \Cref{rmk_delta_case2} and by $s_{12} = 0$, it would be $P_1 = P_3$. Hence,
from the results of \Cref{rank_8} we have that $P_1 \vee P_2$ and $P_1 \vee P_4$ are tangent to the isotropic conic in~$P_2$ and~$P_4$. Hence $s_{22}, s_{44}, s_{23}, s_{45}$ are all zero and \Cref{p6formula} and \Cref{p7formula} specialize to \Cref{formuleP6_P7}; finally, \Cref{prop:rk8_2B} shows that the points are in a $(C_5)$ configuration.
\end{proof}
%

\begin{rmk}
Since in a $(C_5)$ configuration we have 
$P_1 = P_2 \times P_4$ and $(P_2, P_4, P_6)$ aligned, it follows that $s_{12} = 0$,  $s_{14} = 0$ and $s_{16} = 0$.
Moreover, it is
immediate to verify that the line~$P_2 \vee P_4$ is orthogonal to the
lines~$P_1 \vee P_2$, $P_1 \vee P_4$, and~$P_1 \vee P_6$, that is
%
\[
  \scl{P_1 \times P_2}{P_2 \times P_4} = 0, \quad
  \scl{P_1 \times P_4}{P_2 \times P_4} = 0, \quad
  \scl{P_1 \times P_6}{P_2 \times P_4} = 0,
\]
hence \Cref{rmk:three_orthog} gives that the $(C_5)$ configuration obtained in \Cref{prop:rk8_2B} is a sub-case of the $(C_5)$ configuration described here.
%
\end{rmk}
%
%\begin{rmk}
%The family of all the cubics with the eigenpoints in configuration $(C_5)$
%are therefore obtained taking $P_2, P_4$ generic, $P_1 = P_2 \times P_4$,
%$P_5$ generic on the line~$P_1 \vee P_4$ and $P_3$ given by the \Cref{p3formula}.
%It is easy to see that the cubics with eigenpoints in configuration $(C_5)$
%described in~\Cref{prop:rk8_2B} are a sub-family of the cubics here described,
%obtained by the conditions $s_{22} = s_{44} = 0$.
%\end{rmk}

%% conti nel file config5.sage.

The strict configurations $(C_6)$ and~$(C_7)$ cannot be realized by eigenpoints.
Indeed, if we consider suitable ideals given by the conditions
$\delta_1$ and~$\delta_2$ that must be satisfied by the points of the two configurations,
we get that there are no compatible solutions.

\underline{Configuration (8)} is realizable.  
%% file config8N2.sage+
The $(C_8)$ configuration we consider in this paragraph is the following: 
%
\begin{equation}
\label{config:c8}
(P_1, P_2, P_3), \ (P_1, P_4, P_5), \ (P_1, P_6, P_7), \ (P_2, P_4, P_6), \ 
(P_2, P_5, P_7), \ (P_3, P_4, P_7)    
\end{equation}
%
Hence the points $P_1, P_2, P_4, P_7$ lie on three lines, while the points
$P_3, P_5, P_6$ lie on two lines.
The following lemma is easy to verify:

\begin{lemma}
%% v. file configN82.sage
\label{lemma:6ortog} If we have four distinct points of the projective
plane, such that no triplets of them are collinear, then the scalar 
product of at least two of them is not zero.
\end{lemma}

Furthermore:

\begin{lemma}
\label{lemma:three_s_zero}
Suppose that $P_1, \dotsc, P_7$ are eigenpoints in a $(C_8)$ configuration.
We also assume that the matrix of conditions of any $V$-configuration contained in $P_1, \dotsc, P_7$ has rank~$9$. 
Let $U, V, W$ be three points in the set $\{P_1, P_2, P_4, P_7\}$
such that $\scl{U}{V} = 0$ and $\scl{U}{W} = 0$.
Then $\scl{V}{W} = 0$.
\end{lemma}
\begin{proof}
%% vedi file config8N2.sage
From the symmetry of the points, we can assume that $U=P_1$, $V=P_2$, 
$W=P_7$.
We give to the points $P_1, P_2, P_4, P_7$ arbitrary coordinates 
in such a way that $s_{12}= s_{17}=0$ and 
we compute $P_3$ as the intersection of the lines $P_1 \vee P_2$ and 
$P_4 \vee P_7$. The matrix of conditions of the $V$-configuration 
$(P_3, P_1, P_2, P_4, P_7)$ 
is of rank~$9$, hence $\delta_1(P_3, P_1, P_4) = 0$ 
(see~\Cref{theorem:rank_V} and~\Cref{proposition:third_alignment}). 
Similarly, we determine $P_5$ and~$P_6$ and it must hold:
$\delta_1(P_5, P_1, P_2) = 0$ and $\delta_1(P_6, P_1, P_2) = 0$. To 
check  that $s_{27} = 0$, we could 
therefore consider the ideal generated by these three polynomials. 
In order to speed up the computations, it is better to first construct
the ideal
%
\[
  J_1 = 
  \bigl(
    \delta_1(P_3, P_1, P_4), \delta_1(P_3, P_1, P_7),
    \delta_1(P_3, P_2, P_4), \delta_1(P_3, P_2, P_7)
  \bigr)
\]
%
and to saturate it w.r.t.\ the condition that $P_2, P_4, P_7$ are not 
aligned; similarly, construct the ideals~$J_2$ and~$J_3$ obtained 
from the $V$-configurations $(P_5, P_1, P_4, P_2, P_7)$ and 
$(P_6, P_1, P_7, P_2, P_4)$. Then the radical of the ideal $J_1+J_2+J_3$
turns out to be $s_{27}$. This shows the lemma.
\end{proof}

\begin{prop}
\label{prop:conf8_partA}
If $P_1, \dotsc, P_7$ are eigenpoints in a $(C_8)$ configuration, then, up to a permutation of 
the labels of the points, the quantity~$s_{12}$ is not zero and it holds:
\begin{equation}
\label{formula:ortocentro}
P_7 = (P_1 \times P_2)s_{14}s_{24} -
  s_{12}(P_1 \times P_4)s_{24} + s_{12}s_{14}(P_2 \times P_4)
\end{equation}
and the lines $P_1 \vee P_2$, $P_1 \vee P_4$ and~$P_1 \vee P_6$
are orthogonal to, respectively, $P_3 \vee P_4$, $P_2 \vee P_5$ and 
$P_2 \vee P_4$, hence:
%
\[
  \scl{P_1 \times P_2}{P_3 \times P_4} = 0, \quad 
  \scl{P_1 \times P_4}{P_2 \times P_5} = 0, \quad 
  \scl{P_1 \times P_6}{P_2 \times P_4} = 0 \,.
\]
%
\end{prop}
\begin{proof}
%% file config8N2.sage
If some of the $V$-configurations contained in $(C_8)$ are such that their matrices of conditions have rank$8$, 
the result follows from \Cref{rank_8} (see \Cref{rmk:three_orthog}).
We fix generic coordinates for the points $P_1, P_2, P_4, P_7$.
As a consequence of the previous two lemmas, we have that it is not possible
to have $s_{12}=0, s_{14}=0, s_{17}=0$, hence we can assume $s_{12} \not=0$.
Hence, again from \Cref{lemma:three_s_zero}, it is not possible to 
have $s_{14}=s_{24}=0$, hence at least one of the three coefficients~$s_{14}s_{24}$, $s_{12}s_{24}$, or~$s_{12}s_{14}$ is not zero. Since the 
three vectors $P_1\times P_2$, $P_1\times P_4$, $P_2\times P_4$ are linearly independent, we see that $P_7$ is a non zero vector. Also in this general situation, if the points are eigenpoints, it must hold: 
$\delta_1(P_3, P_1, P_4) = 0$, $\delta_1(P_5, P_1, P_2) = 0$ and 
$\delta_1(P_6, P_1, P_2)=0$. These three quantities, up to non zero factors, are 
%
\begin{align*}
 e_1 &= s_{12}s_{47}-s_{17}s_{24} \,,\\
 e_2 &= s_{12}s_{47}-s_{14}s_{27} \,,\\ 
 e_3 &= s_{17}s_{24}-s_{14}s_{27} \,,
\end{align*}
%
hence $e_1-e_2+e_3 = 0$ and the system
$e_1=0, e_2 = 0$ is a linear system in the coordinates of the points. 
Its solution (w.r.t.\ the coordinates of~$P_7$) gives new coordinates for the point~$P_7$ and it is easy to see that \Cref{formula:ortocentro} holds.
\end{proof}

We have also the converse, i.e.\
\begin{prop}
Suppose $P_1, P_2, P_4$ are three points of the plane. Suppose that \Cref{formula:ortocentro} defines a point in~$\p^2$ and that
%
\[
P_3 = (P_1 \vee P_2) \cap(P_4 \vee P_7), \quad 
P_5 = (P_1 \vee P_4) \cap (P_2 \vee P_7), \quad
P_6 = (P_1 \vee P_7) \cap (P_2 \vee P_4).
\]
%
also define points in~$\p^2$.
Then the points $P_1, \dotsc, P_7$ are in a $(C_8)$ configuration
and are eigenpoints of a unique cubic 
of the plane. In particular:
\[
  \scl{P_1 \times P_2}{P_3 \times P_4} = 0 \,, \quad 
  \scl{P_1 \times P_4}{P_2 \times P_5} = 0 \,, \quad 
  \scl{P_1 \times P_6}{P_2 \times P_4} = 0 \,.
\]
\end{prop}
\begin{proof}
%% file connfig8N2.sage
It is enough to show that the rank of the matrix $\Phi(P_1, \dots, P_7)$
is $9$ and this computation can be done, for instance, assuming that 
$P_1$ is the point~$(1: 0: 0)$ or the point~$(1: \iii: 0)$.
\end{proof}

\begin{corollary}
The locus of ternary cubic forms~$F$ with a $(C_8)$ configuration has dimension~$6$.
\end{corollary}


\subsection*{Comparison with ODECO tensors}
%% file odeco.sage
As mentioned in the introduction, a class of
symmetric tensors fitting in the framework of collinearities in the eigenscheme is represented by ODECO tensors, which were introduced by~\cite{Rob} and studied by \cite{BDHE, Koiran2021, Biaggi2022}.

Possibly after an $\SO_3(\C)$ transformation, such forms are of the type
%
\[
  f = \lambda_1 x^3 +\lambda_2 y^3 + \lambda_3 z^3
  \quad \text{for some } \lambda_1, \lambda_2, \lambda_3 \in \C \,,
\]
%
and, using the language of \cite{Rob}, they admit the following three pairwise orthogonal eigenvectors $Q_1, Q_2, Q_3 \in \C^3$
%
\[
  Q_1 = \left( \frac{1}{\lambda_1},0,0 \right) \,, \quad
  Q_2 = \left( 0,\frac{1}{\lambda_2},0 \right) \,, \quad
  Q_3 = \left( 0,0,\frac{1}{\lambda_3} \right) \,.
\]
%
The remaining $4$ eigenvectors are uniquely determined, precisely:
%
\[
  Q_4 = Q_1+Q_2+Q_3\,, \ Q_5 = Q_1+Q_2,\, \ Q_6 = Q_1+Q_3 \,, \ Q_7 = Q_2+Q_3\,.
\]
%
In particular, we see that such eigenschemes are in a $(C_8)$ configuration, with the additional condition
%
\[
  \left\langle Q_i,Q_j \right\rangle = 0
  \quad \text{ for } (i, j) \in \{(1, 2), (1, 3), (2, 3), (1, 7), (2, 6), (3, 5)\} \,.
\]
%
The seven vectors fit into the description of \Cref{prop:conf8_partA}, since $Q_3 $ is proportional to $Q_1 \times Q_2$ and this is coherent with \Cref{formula:ortocentro}
(because $\scl{Q_1}{Q_2} = 0$)
and
%
\[
  (Q_1 \vee Q_2) \times (Q_3 \vee Q_5) = 0 \,, \quad
  (Q_1 \vee Q_3) \times (Q_2 \vee Q_6) = 0 \,, \quad
  (Q_2 \vee Q_3) \times (Q_1 \vee Q_7) = 0 \,.
\]
%
In particular, the locus of ODECO forms is a $5$-dimensional sublocus of the $(C_8)$ locus.
It is a challenging question to investigate the irreducible components and the degrees of such loci.


\bibliographystyle{alphaurl}
\bibliography{biblio}

\end{document}